---
title: "Term paper"
subtitle: ""
author: "Rose Hörsting & Gina Reinhard"
date: "2025-03-15"
engine: knitr
bibliography: references.bib
format: 
  html:
    df-print: default
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
    number-sections: true
    embed-resources: true
link-citations: true
csl: "apa_7th.csl"
number-sections: true
---

```{=html}
<style>
p {
  text-align: justify;
}
</style>
```

```{r load-libraries, include=FALSE}

library(here)
#install.packages("glossr")
library(glossr)
library(tidyverse)

```

```{r styling, include=FALSE}

use_glossr(styling = list(
  source = "b",
  first = "",
  font_family = "Cambria"
))

```

```{r import-data, include=FALSE}

sefaire_sample <- read.csv(file = here("data", "sefaire_sample.csv"))
etre_sample_1 <- read.csv(file = here("data", "etre_sample_1.csv"))
etre_sample_2 <- read.csv(file = here("data", "etre_sample_2.csv"))
etre_sample_3 <- read.csv(file = here("data", "etre_sample_3.csv"))
etre_sample_4 <- read.csv(file = here("data", "etre_sample_4.csv"))
```

```{r etre-sample, include=FALSE}

etre_sample <- rbind(etre_sample_1, etre_sample_2, etre_sample_3, etre_sample_4)

```

```{r randomisation, include=FALSE}

set.seed(2142637384) #lottery numbers too long, deleted last digit

etre_sample_randomised <- etre_sample |>
  slice_sample(n = 700)

write.csv(etre_sample_randomised,"./data/etre_sample_randomised.csv")

```

```{r downsampling, include=FALSE}
etre_sample_clean <- read.csv(file = here("data","etre_sample_clean.csv"))
set.seed(2142637384)

etre_sample_532 <- etre_sample_clean |> 
  slice_sample(n= 532)
```

```{r downsampling-false-positive, include=FALSE}
#one false positive was found after sampling (ID=332) but sampling completely new not possible
set.seed(2142637384)

etre_sample_533 <- etre_sample_clean |> 
  slice_sample(n= 533)

etre_sample_1 <- etre_sample_533 |> 
  filter(!etre_sample_533$ID %in% etre_sample_532$ID)
```

```{r downsampling-correction, include=FALSE}
# ID217 is "next in line" given the seed
# correction: remove 332 and add 217 to sample
etre_sample_532 <- etre_sample_532 |> 
  filter(ID != "332") |> 
  bind_rows(etre_sample_1)

write.csv(etre_sample_532,"./data/etre_sample_532.csv")

```

```{r etre-difference, include=FALSE}
difference_list <- etre_sample_clean |> 
  filter(!etre_sample_clean$ID %in% etre_sample_532$ID)

difference_ID <- difference_list["ID"]

```

```{r cql-lemmas, include=FALSE}
unique_lemma <- unique(sefaire_sample$lemma)
lemma_list <- sprintf('lemma="%s"', unique_lemma)
cat(paste(lemma_list, collapse = " | "))

```

# Introduction

-   The construction se faire + infinitive has developed through analogy with the passive voice. While it originally had a causative meaning (the subject has an active influence), in certain cases it can take on a purely passive reading, i.e. it behaves passively as an alternative to être.
-   quantitative, multivariational, probabilistic approach / method of statistical modelling of corpus data

# The French passive alternation of *être* and *se faire*

## Canonical passive with *être*

### Structure and meaning

The canonical passive with *être* in French inverts the syntactic functions of the arguments within a clause. The object of the active sentence becomes the subject in the passive voice - a phenomenon known as diathesis. In French, the subject usually occurs at the first position of the sentence, and the passive voice uses this syntactic priority to shift the focus and emphasise the undergoer over the agent.

Interpreting the passive **without** an agent: The passive allows the subject to be omitted, which generally has the effect of presenting the situation without specifying the cause or the individual who is responsible. The passif court the action or the result of the action. (GGF: 219)

-   Unlike e.g. the causative construction, the passive takes only an affected entity as its subject (according to Labelle 2002) (se faire)

The French passive construction consists of the *participe passif* (passive participle), which has the same form as the *participe passé* (past participle) and in most cases *être* as an **attributive** verb. In the passive construction, the subject of the active sentence becomes an adjunct phrase (*complément d'agent*, introduced by *par* or *de*), which can be omitted, while the active direct object takes the place of the subject (GGF 210-211). -\> examples?

We refer to a *passif long* if there is a complément d'agent (1), and a *passif court* in the other cases:

\(1\) *\[...\]* *ils sont élus par nos concitoyens au suffrage universel.* (example from ParlaMint corpus)

\(2\) *Les députés sont élus pour contrôler votre action.* (example from ParlaMint corpus)

An impersonal passive (*passif impersonnel*) is also possible in French for some verbs without a complément direct (*Il a été ri de cette histoire.*). With impersonal passives, the direct object can remain as a complement, as in *Il sera acheté un livre* (‘A book is bought’) (GGF 210).

Some grammars distinguish between a *passif verbal*, with an event interpretation, and an *passif adjectival*, with a resultative interpretation. For (GGF: 219), this is about semantic differences, concerning the aspect of the situation. With a state verb, active and passive sentences are generally semantically equivalent (Tout le monde aime Marie. / Marie est aimée de tout de monde.). With an event verb, expressing a change of state or place, the passive describes a change in the initial situation or the result of this change. In the passé composé, the sentences *On a servi la biere* and *La bière a été servi* appear to be equivalent. In the present tense, on the other hand, or in the imperfect tense, the active phrase (On sert la bière) may describe a situation in progress, and the passive phrase a resulting state (La bière est servi). In the passive, the presence of an agent complement can favour the interpretation of an event, with a present tense of habit (La bière est servi par le patron). (GGF: 219)

The passive voice is relatively rare in French, accounting for only 5% of verbs in the French Treebank (FTB) (\@da_cunha_lalternance_2020), compared to 9% in English for the same text type (\@roland2007frequency). Overall, the passive voice is more frequently used in written French than in spoken French (\@da2022alternance). Approximately a quarter of passive constructions in the FTB are *passif long* (\@da_cunha_lalternance_2020), while in spoken French, this proportion drops to about 10% (source: Hamma). -\> frequency in our corpus?

## *se faire* passive

```{=html}
<!-- ### Historical background

@bellec_grammaticalisation_2015 and @bellec_quand_2014 date the se faire + INF passive construction (PSF) to the middle of the 20th century and call it a "modern" construction that has since experienced a rapid increase in popularity. She describes a process of grammaticalisation and reanalysis for the PSF during/through which “se” has lost its reflexive meaning and “faire” has become an auxiliary. In merging, they form a “semi-auxiliary” and obtain a passive interpretation, while the causative meaning is lost (désémantisation). However, @bellec_construction_2014 claims that the causative reading did not disappear but instead coexists with the passive reading (debated, see below).

-   In contrast, @raineri_analyse_2010 [280-305] based on corpus data: sees origin much earlier, exists since origin of Middle French (14-17th century); NOT derived from causative reading, increased productivity from 19th century onward -->
```

### Structure and meaning

The *se faire* passive construction provides an alternative way to express passive meaning in French: When constructing a passive sentence, the French speaker may choose to use *se faire* instead of the canonical *être*: The two constructions form a passive alternation. Various scholars have examined the *se faire* passive, such as @bellec_construction_2014, @labelle_anticausativizing_2013, @koenig_semantic_1992, @kupferman_construction_1995, @raineri_analyse_2010, @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987, to name a few.

The *se faire* passive construction consists of the pronoun *se* (which shows agreement with the subject), a form of the verb *faire ("to do")*, and a verb in the infinitive.

\(3\) *Des policiers se font agresser.* (ParlaMint corpus)\
"Police officers were attacked/ have been attacked."

Similar to the être passive, the direct object of the active sentence becomes the subject of the passive sentence. This subject is the affected experiencer of the action described (compare to example above). Unlike the être passive, however, the diathesis is not restricted to the direct object: The PSF has "greater syntactic flexibility" (@raineri_analyse_2010 [93-95]) and can also occur with ditransitive verbs, where the indirect object becomes the subject:

\(4\) *Les intéressés peuvent se faire expliquer tout ce qui touche aux retraites.* (ParlaMint corpus)*\
*"Interested parties can have everything related to pensions explained to them."

reconstructed active equivalent: "One can explain everything to do with pensions to interested parties." (find one with "par")

The *se faire* passive also allows possessor constructions, where the subject of the passive corresponds to part of the object in the active sentence:

\(5\) *Le coiffeur coupe les cheveux de Pierre. - Pierre s'est fait couper les cheveux (par le coiffeur)\
*"The hairdresser cuts Pierre's hair - Pierre had his hair cut (by the hairdresser)."

However, whether constructions like (5) count as true passives is controversial. Some scholars argue that the se faire construction is fundamentally causative rather than passive, since the subject both initiates and undergoes the action (e.g. "Pierre made the hairdresser cut his hair"). More on that in @sec-debate.

The semantics of the components, mainly the fundamentally causative nature of the verb "faire" and the passive reading, create a semantic discrepancy/ tension: @labelle_anticausativizing_2013 \[235\] classifies the construction as “formally causative, but semantically passive”. This apparent contradiction has led scholars to question the nature of the construction, as @koenig_semantic_1992 [145] put it: "how can a passive-like construction arise from a combination of a causative and a reflexive marker, where no marking of the passive function is apparent?".

Various answers to this question have been proposed in the literature: @kupferman_construction_1995 sees the causative and passive readings as homonyms with fundamentally different semantics and therefore assigns an inherent passive status to the se faire construction. This is in line with @bellec_construction_2014 and @bellec_quand_2014, who takes a diachronic approach: She describes a process of grammaticalisation and reanalysis for the PSF through which “se” lost its reflexive meaning and “faire” became an auxiliary. According to her, the se faire passive has been grammaticalised into a semi-auxiliary with non-transparent components. In this process, it has lost its causative meaning, which, however, has not disappeared but now coexists with the passive reading. Note, however, that @raineri_analyse_2010 [280-305]'s corpus study suggests that the se faire passive did not derive from the causative reading but evolved independently.

A syntactic approach is taken by @labelle_anticausativizing_2013 [23], who suggests that the passive meaning is obtained by a syntactic mechanism that involves binding the empty NP-object position of the embedded verb to the subject via the reflexive particle "se" [^1]. However, the distinction between causativity and passivity with regard to *se faire* remains controversial, as will be described in the following @sec-debate.

[^1]: In the se faire passive, the embedded verb (i.e., the infinitive after faire) has an empty NP-object position since only the subject is realised. The reflexive clitic "se" binds the empty object position of the embedded verb. Therefore, "se" connects the (missing) object-NP and the surface subject (coindexation). In contrast to the être passive, the object does not become the new subject but is connected to the subject by "se". The subject is interpreted as affected because it is linked to the missing object of the embedded verb (which would be affected in the active sentence but is absent on the surface in se faire). For this to work, the embedded verb needs an agent as a subject (because otherwise, the missing object would not be affected).

### Passive vs. causative - Debate of the construction's passive-status {#sec-debate}

The above-described stance taken by @bellec_construction_2014, @labelle_anticausativizing_2013, @koenig_semantic_1992 and @kupferman_construction_1995 that the PSF is a full-fledged passive construction is by no means undisputed. The opposing view - that the construction is a type of causative - is also widely represented in the literature.

It is most prominently presented by @tasmowski-de_ryck_causatif_1987 and @gaatone_desagreable_1983, and more recently by @sveberg_se_2012. In their view, passive meaning emerges from verb semantics and pragmatics, not from the PSF construction itself. They centre their argumentation that the se faire passive is a weak reading of the causative around the notion that the subject of the se faire construction "is always, even unconsciously, responsible for the occurrence of the embedded event." As @raineri_analyse_2010 [89] describes, there is also the possibility of a multitude of readings (@kokutani_sur_2005) and of a continuum between causative and passive (Hobak Haff 2000), eventually adapted by @raineri_analyse_2010 [154-166].

Sentences including the *se faire* construction can have both/either a passive and/or a causative sense (@koenig_semantic_1992\[143\]):

\(6\) *Ce qui signifie que si nous ne le votons pas, nous **nous ferons taper** des deux côtés.\
*This means that if we don't vote for it, **we”ll be slammed** from both sides.\
This means that if we don't vote for it, **we'll make ourselves get slammed** from both sides.

\(7\) *il y a \[…\] des parlementaires qui **se sont fait élire** sur la promesse de ne pas supprimer 120 000 postes.\
"*there are \[…\] Members of Parliament **who** **were elected** on the promise not to cut 120,000 jobs.”\
"there are \[…\] Members of Parliament **who** **made themselves be elected** on the promise not to cut 120,000 jobs.”

As @washio_when_1993 \[45\] describes, the linguistic proximity of causativity and passivity is not exclusive for French but quite typical in cross-linguistic contexts, stating that "passive and causative senses are shown to share a basic conceptual structure" across the world's languages. Crucially, passive se faire and causative se faire are structurally identical, differing only in their semantics. According to @labelle_anticausativizing_2013 [236] the ambiguity is resolved by "the lexical items involved", that is, by the semantic nature of the three main components: subject referent, "se" and "faire".

Referring to these three components, @bellec_construction_2014 [207] defines the following criteria for a causative reading, which distinguish the causative from the passive interpretation:

-   In the causative, the subject referent is agentive and has the role of both causer and patient. The subject actively initiates what happens to them. In the passive, the subject referent has only the role of the patient and is affected by the event.
-   In the causative reading, the particle “se” has a reflexive meaning whereas in the passive reading, “se” is not reflexive.
-   In the causative reading, “faire” has transparent causative meaning. This causative meaning is lost in the passive reading, where “faire” has no inherent meaning. <!-- add: how will we distinguish them / have we distinguished them in our study -->

Thus, causative and passive readings of the se faire construction are distinguished mainly by the transparency of the lexical items, whether they are understood literally or as constructions without transparent meaning. Sometimes both readings are possible and the ambiguity cannot be resolved (@koenig_semantic_1992\[144\]): (@bellec_construction_2014\[206, 207\] suggests that this is especially the case when cognition and perception verbs are involved.) Therefore, despite its productivity as a passive construction, @raineri_analyse_2010 \[356-357\] concludes that the PSF is problematic because it neither fits into two homonyms nor functions as a single weak causative.

<!-- In the causative reading, "se" is understood as reflexive and "faire" as causative. In contrast, in the causative reading, if interpreted as a passive, the meanings of "se" and "faire" are not transparent. Thus, the subject referent does not necessarily cause the event and is not necessarily responsible (@koenig_semantic_1992\[143\]: "any notion of causation is absent"). -->

<!-- As @labelle_anticausativizing_2013 [15] explains: -->

This debate has been central to the investigation of the se faire constructions and cannot be left aside. However, even if the status of the PSF has been the subject of disagreement and discussion, the literature clearly shows that there is a passive reading associated with this construction. This is independent of whether the passive sense emerges from within the construction or from the components it occurs with, or whether it has an underlying causative meaning. By comparing *se faire* and *être* constructions, and examining the factors contributing to the choice of one or the other, we may be able to shed light on this debate from a corpuslinguistic perspective. <!-- (-> exclude causative readings?). Can our findings contribute to the debate of whether se faire deserves passive-status? If not: no need to spell out the debate in detail. -->

## *être* and *se faire*: Competition or complementation? - Defining the variable context

Both se faire + Vinf and être + Vé are available for the active diathesis [@novakova_construction_2009 [108]] and assign the role of an undergoer to the subject NP. However, as @novakova_construction_2009 [108] notes, the constructions differ in semantic nuances: "Elles ont un sens proche mais pas identique." Thus, they are not semantically identical but rather potentially carry different meanings that may be more or less pronounced depending on the context: Previous research suggests that se faire passive is more likely to have a causative undertone while the être passive is more neutral - possibility of a continuum/ passive gradient. However, this is what makes the analysis interesting and our research worthwhile - what are the exact factors that influence the choice between the the two variants?

According to @raineri_analyse_2010 [87], the se faire passive is "une tournure marquée, sémantiquement contrainte et associée à des types de discours oraux et/ou informels ("marked, semantically constrained phrase associated with oral and/or informal discourse").

The two passive variants may also be used in the same sentence:

*Je remarque d'' ailleurs que ce sont souvent ceux qui filment qui se font éborgner ou sont blessés par des tirs de LBD.*

"Incidentally, I've noticed that it's often those filming **who get their eyes poked** out or **are injured** by LBD fire."

This introduces certain difficulties: The passive/ causative continuum discussed in @sec-debate makes the definition of the variable context particularly challenging as some constructions may be more passive than others - where to draw the line?

@bohmann_large-scale_2023 [35] face similar challenges when defining the variable context of the be and get passive: "passive gradient" exist where both GET and BE may be used, but where the choice between the two entails a semantic difference", "definition of the variable context is not just a problem of automation, but of linguistic interpretation" \[36\]. Follow

Since our aim is to investigate the passive alternation of être and se faire, it is important to ensure that we only compare the être passive with real passive se faire constructions, and not with causative constructions (@bohmann_large-scale_2023 [35]: "all extracted forms should be genuine instances of the variable under investigation").

We follow @raineri_analyse_2010 \[165\] in including only se faire + Inf constructions that have a natural paraphrase with the être passive in order to include only constructions that directly contrast with être. Following @bohmann_large-scale_2023 [35] who claim that "only cases where both variants can be used interchangeably should be considered", we aim to exclude constructions that do not allow for a paraphrase with the other.

Passive constructions of se faire and être are not always interchangeable: As described above, the canonical passive with être requires a direct object in the active equivalent which becomes the new subject, @koenig_semantic_1992 \[144\], @washio_when_1993, @labelle_anticausativizing_2013 [16]. Thus, for reflexive and ditransitive verbs, the être passive is not available and does not compete with se faire. These constructions will be excluded and only transitive verbs will be studied.

-   se faire passive has evolved to be used in situations without a causer (passive reading) @bellec_construction_2014
-   figurative, "besoin d’expressivité and extravagance" instead of "used up" canonical passive @bellec_construction_2014 \[206\]; part of lexicalised expressions @novakova_construction_2009 [117]/ idioms @raineri_analyse_2010 [165]
-   while some scholars speak of semantic differences between the canonical être passive and the se faire passive, @kupferman_construction_1995 sees them as semantically synonymous -\> therefore: competition
-   how much is it used today (according to authors/ background literature) and in which settings
    -   se faire less formal, more figurative and emotional, used to build the passive with indirect object as the agent & to express passivity in dynamic/ ongoing processes, more spoken
    -   etre passive more “neutral”, more formal, more written
-   être: exclude adjectives and some participles that do not form the passive (similar to @bohmann_large-scale_2023 \[36\])/ "instances that allow for gradation" @hundt_be-_2024 [92]

### Maybe: Comparison to similar alternations in other languages

-   English @bohmann_large-scale_2023, East Asian languages @washio_when_1993 and @guan_multifactorial_2024 , Swedish @cuypere_corpus-based_2014
-   se faire passive "typologically marked", comparison to other Romance languages @koenig_semantic_1992 \[145\]; similar constructions exist in other Romance languages but passive reading only allowed in French @labelle_anticausativizing_2013 [236]

# Data and Methods

## Data Collection

### Corpus

In our investigation of spoken French language, we aimed to select a corpus that represents authentic French spoken language. After thorough consideration and exploration of our options, we decided for the French ParlaMint corpus [@11356/1431]. ParlaMint consists of French parliamentary debates from 2017 to 2020 and contains 38,702,395 tokens (32,214,147 words). We chose this corpus because it is well-documented, POS-tagged and contains rich meta data.

We recognize that the genre of parliamentary debates is quite specialised and our results therefore cannot be generalized to all French spoken language. However, the language of parliamentary debates is a fascinating genre in itself (studied before: references) and is expected to reflect a wide range of topics, thus providing a cross-section of language varieties as they occur in French speech.

We extracted the concordances from the French ParlaMint corpus [@11356/1431] using a CQL expression in SketchEngine. To extract the data we chose a semi-automatic approach that included manual filtering.

-   We selected an equal number of occurrences of *être* passive and *se faire* passive.
    -   We made a list of the lemmas in the se faire constructions → used these lemmas to filter the être constructions

    -   Exclusion of causative reading of se faire constructions: the same verbs have been used with être passive in the same corpus (find/extract the constructions) → depends on how many constructions we have left
-   We have not taken into account impersonal passives (*example from corpus*)
-   être passive: In their active/passive alternation study @da_cunha_lalternance_2020 only chose passive sentences in the compound past tense "NP a été VP" in order to favour verbal passives over adjectival passives (*La porte est ouverte*). However, we chose to include the present tense/all tenses. This decision was based on previous literature suggesting that the present tense favors *se faire* passive over *être* passive, making tense an important factor in our study.
-   In line with @da_cunha_lalternance_2020 we chose complément d'agents introduced by the preposition *par*, avoiding additional variation caused by *de*.

```{=html}
<!-- problem: data are extremely unbalanced (ratio in ParlaMint être vs. se faire: ) -\> making the ratio 50: 50 would distort the reality significantly but if the ratio is kept the model will almost always predict être
strategy possibilities:
    -   build a corpus where être is presented more than se faire but not as extreme (approximation, e.g. 40% se faire - 60% etre) ("over/undersampling within limits")
    -   assign more weight to se faire constructions by "weights=" variable of glm function
    -   @bohmann_large-scale_2023 kept the imbalance (ca. 3 Mio be passive vs. 25 000 get passive)
    -   evaluate with precision & recall, not just accuracy (F1-score) -->
```

```{r lemma-list, include=FALSE}

lemmas <- sefaire_sample |> 
  count(lemma)

```

### Semi-automated approach

we opted for a balanced sample containing as many se faire as être passives (numbers and characteristics)

**se faire passive**

We automatically extracted a sample of 1,254 *se faire* + infinitive constructions from the ParlaMint 2.0 corpus on Sketch Engine \[source?\] using a CQL expression. After extraction, we manually filtered the data to exclude false-positive constructions (i.e. constructions that were not *se faire* + infinitive constructions) and those that did not fulfil our predefined criteria for *se faire* passives (see below). The final sample comprised 535 \[count with inline code\] *se faire* passives. We then annotated the lemmas of these constructions and relevant linguistic features, such as verb form, subject animacy and grammatical gender, in order to prepare the data for statistical modelling.

**être passive**

-   extract être passive constructions with CQL expression
-   In Sketch Engine: Import Word List → lemma list extracted from se faire constructions (for comparability of the constructions, making sure that differences between être and se faire are not due to individual differences of the verb) → Filter out only verb lemmas of se faire from concordance lines → download

**determine sample by setting a seed**

-   seed determined by future lottery numbers (12.3.2025, <https://www.lotto.de/lotto-6aus49/lottozahlen>) → function set.seed( ) in R
-   take a random sample from the concordance lines that is equal in size to the manually filtered amount of se faire constructions
-   final sample: same amount of constructions for être and se faire, same lemmas, seed set
-   Our sample therefore comprises XXX sentences/constructions: XX se faire and XX être passive, ... (short summary)

### Borderline cases/ Manual filtering exclusion criteria

As some linguists (e.g., Martin, Novakova / ... as above?) argue, *se faire* + infinitive can sometimes have a 'passive effect' even without a direct 1:1 alternation with *être*. However, for our study of *se faire* + Infinitive and *être* + past participle as a passive alternation, our main criterion was to include only those constructions that could be paraphrased with the *être* passive. Consequently, we excluded expressions that could not be rephrased with the *être* passive, either because the *être* construction was ungrammatical or the meaning was fundamentally altered.

A typical example of a construcion that met our criterion is:

*Il s’est fait insulter par son collègue.* (“He was insulted by his colleague.”) \[eigenes Beispiel ersetzen\]

*Il a été insulté par son collègue.* (“He was insulted by his colleague.”) \[eigenes Beispiel ersetzen\]

By contrast, expressions such as *se faire connaître* were excluded. For instance:

*Il s’est fait connaître grâce à ses romans.* (“He became known through his novels.”) \[eigenes Beispiel ersetzen\]

is acceptable, but

*\*Il a été connu grâce à ses romans.* \[eigenes Beispiel ersetzen\]

is unusual, as *être connu* does not carry the same meaning.

However, our somewhat "fuzzy" definition of what constitutes a passive alternation [@bohmann_large-scale_2023] posed some challenges: We excluded purely causative constructions, but the distinction between causative and passive was sometimes unclear. When is a construction semantically similar enough, and when is its meaning too causative? As expected, given the extensive debate on PSF's passive status (@sec-debate) and the challenges reported by @bohmann_large-scale_2023 in their investigation of the English *be*/*get* passive alternation, we came across a significant number of "unclear" cases during the manual filtering of *se faire* constructions.

(This approach is also supported by linguistic analyses that suggest a continuum between causative and passive meanings in *se faire* constructions. that is, *se faire* constructions can switch between passive and causative interpretations depending on the context. The flexibility of this construction suggests that a strict binary classification is inappropriate and reinforces the idea of a causal/passive continuum - at one pole a reduced agentivity of the subject's referent - rather than a rigid distinction. @raineri_analyse_2010 pp. 154ff)

(If the action of the subject is reduced to mere initiation (\[+ déclenchement\]), the se faire construction can be completely paraphrased with an être passive without changing the content of the sentence. For example, *Il se fait remarquer par une patrouille de police* (“He is noticed by a police patrol”) and Il est remarqué par une patrouille de police (“He is noticed by a police patrol”) show an extensive equivalence between se faire and être. Similarly, the other elements in the co-text of *élire* examples weakens the causative strength and strengthens the passive effect. -\> agentivité **variable** / passive-causative continuum with *élire* depending on context @raineri_analyse_2010 p. 160)

We have found many examples of se faire constructions cases where a clear binary distinction between passive and causative was not possible, for example because when the subject is both the undergoer and the causer of the action.

In general, we opted for a broad approach, including as many constructions as possible. This means that even when the causative reading is more natural (because the subject triggers the event), we included the construction as long as a passive understanding was also plausible (since the subject is also undergoing an action). In doing so, we deviate from the first criterion established by @bellec_construction_2014. Examples of such constructions include *se faire vacciner* "to (go) be/get vaccinated" and *se faire élire* "to be/get elected" or "to stand for election".

Some verbs can have a passive and a causative meaning, depending on the context. For example, *se faire remarquer* can mean both "to get noticed" and "to draw attention to oneself". To ensure consistency and practicality, we decided to include or exclude verbs on the basis of their lemma, rather than on individual occurrences. Below, we outline below some high-frequency ambiguous cases and provide our reasoning for the decidions to include or exclude them.

We excluded:

-   Fixed idiomatic expressions with a figurative meaning: "se faire passer pour" *to impersonate sb.*, "se faire avoir" *to be fooled*, "se faire mousser" *to boast/ to brag, "se faire taper sur les doigts" lit. to get a slap on the fingers, to be scolded* ; example: "Dites donc, vous savez vous faire mousser!" *My, you know how to make a fuss!/ You know how to blow your own trumpet!*

    -   \+ *se faire **rouler** dans la farine + se faire **mener** par le bout du nez*

-   Expressions that cannot be converted into a correct active sentence (i.e., the direct object of the active sentence is not the subject of the passive sentence). (ditransitive, intransitive verbs ?), example: "mais se faire expliquer des elements de justice n'est pas la même chose que d'acheter en ligne des produits bio ou une paire de chaussures." *but getting an explanation of legal matters is not the same as buying organic products or a pair of shoes online*

-   Expressions containing verbs classified as "idiomes lexicalisés" (lexicalised idioms) according to Raineri (2010: 163). These verbs neither carry a true causative nor passive meaning [^2]: entendre, sentir, attendre, désirer (Raineri 2010: 164); example: "Partout le mécontentement populaire se faire entendre sans qu'' on ait besoin de souffler sur les braises." *Everywhere, popular discontent is making itself heard without the need to fan the flames.*

-   (Expressions with a transparent causative meaning - maybe not necessary if covered by the other criteria)

-   (Expressions where an être passive has a very different meaning)

-   impersonal être passive constructions (passif impersonnel)

[^2]: These verbs mainly occur with inanimate subjects (effects, consequences, discontent). @raineri_analyse_2010 [163-164] proposes that the causative meaning is not available for these verbs since the subject is not causing the event. There is also no passive meaning since they cannot cannot occur with a complément d’agent and do not have a paraphrase with être.

-\> summary of the sample?

### Annotation Guidelines

**Gender (masculin/ feminin/ mixed/ unknown)**

-   Assign masculin or féminin if the gender is clearly marked grammatically:

    -   Through a pronoun (il(s), elle(s))

    -   Through an adjective or participle that marks gender (heureux/heureuse, content/contente)

-   Use mixed only if:

    -   The context clearly refers to a mixed-gender group (un groupe d’hommes et de femmes)

    -   The grammatical form is not gendered, but the semantic context indicates mixed gender (nous referring explicitly to men and women)

-   Assign unknown if:

    -   No grammatical gender marking is present

    -   For on, vous, je, and nous (if no grammatical marking appears)

    -   For expletive subjects (il faut)

**Progressivity (progressive/ non-progressive)**

-   Assign "progressive" if:

    -   The "se faire" construction is directly in the imparfait or future (except for futur antérieur) → indicates an ongoing or habitual action

    -   The "se faire" construction is linked to an ongoing action, indicated by: A participe présent → Example: Ils acceptent de se faire endoctriner. → progressive/ gerondif + participe présent (en courant - while running)

    -   periphrases e.g., être en train de (present continuous), aller + inf (futur proche), continuer à + Infinitif

    -   adverbs/ Adverbial Phrases Indicating Ongoing Action: Temporal adverbs that imply duration or repetition: En ce moment (at the moment), Actuellement (currently), Toujours (always), Depuis (since), Encore (still)

-   Assign "non-progressive" if:

    -   passé composé, plus-que-parfait, présent, futur antérieur (...) - progressiveness is hinted at by another verb in the sentence (e.g., falloir or être) and not from the "se faire" construction itself

### Inter-rater agreement

As the analysis required detailed (and time consuming) manual annotation, we decided to split the annotation work between us. However, to ensure consistent annotation, a representative sample of one third of all instances (320; 33,25%) were double annotated, i.e. by both of us. For these instances, we calculated Cohen's ϰ (Kappa) as a measure of inter-rater agreement (following @levshina_how_2015 [201] and the documentation in the psy package documentation (<https://cran.r-project.org/web/packages/psy/psy.pdf>).)

```{r read-annotated-data}
sefaire_sample_annotated <- read.csv(file = here("data", "sefaire_sample_annotated.csv"))
etre_sample_annotated <- read.csv(file = here("data", "etre_sample_annotated.csv"))
```

**Data preprocessing**

```{r data-cleaning-colnames}
colnames(etre_sample_annotated)
colnames(sefaire_sample_annotated)
```

Some differences between the two datasets, mainly regarding columns speaker and reference: se faire sample has a "speaker" column which includes the speaker id and the name and a "reference" column that includes the speaker ID and the date; etre sample has only a "reference" column that includes speaker ID, date, name

Goal: two columns consistent across both datasets, one "speaker" (name), one "reference" (ID), drop date

```{r data-cleaning-colnames-2}
# 1. se faire
# separate "speaker" column into reference and speaker
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  mutate(reference = sapply(str_split(speaker, ",", n = 2), `[`, 1),
         speaker = sapply(str_split(speaker, ",", n = 2), `[`, 2))
# remove now redundant "Reference", remove "Kommentar" column, put new "reference" column in first position
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  select(-Reference, -Kommentar) |> 
  select(reference, everything())

# 2. être
# separate "Reference" column into reference and speaker
etre_sample_annotated <- etre_sample_annotated |> 
  mutate(reference = sapply(str_split(Reference, ",", n = 3), `[`, 1),
         speaker = sapply(str_split(Reference, ",", n = 3), `[`, 3))

# remove now redundant "Reference" column, remove now redundant "Reference", remove "Kommentar" column, put new "reference" und "speaker" column in first position
etre_sample_annotated <- etre_sample_annotated |> 
  select(-Reference, -Kommentar) |> 
  select(reference, speaker, everything())

#still some differences left: être has an "ID" column which se faire does not; être has two agent_pp columns because it was annotated by A and B while se faire only has one, since agent_pp of se faire constructions was annotated there one person (annotator A)

#rename agent_pp for clarity
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  rename(agent_pp_A = agent_pp)
```

Combine the data

```{r combine-data}
passive_sample_combined <- bind_rows(etre_sample_annotated, sefaire_sample_annotated) 
```

check for typos/ inconsistencies in tagging in combined data; convert to factor; add "NA" as value for all cases that are only annotated by one person

```{r data-cleaning-2}
#check data type
class(passive_sample_combined$subj_animacy_A)
#"character", convert to factor and fill "empty" values with NA
passive_sample_combined <- passive_sample_combined |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  mutate(across(where(is.character), factor)) |> 
  mutate(across(where(is.factor), droplevels)) # to not include NA as a level
# summary
passive_sample_combined |> 
  select(subj_animacy_A, subj_animacy_B, subj_gender_A, subj_gender_B, agent_pp_A, agent_pp_B, modal_A, modal_B, tam_A, tam_B, cluster_A, cluster_B, adversativity_A, adversativity_B) |> 
  summary()
#looks good
```

```{r cohens-kappa}
# install.packages("psy")
library(psy)
kappa_animacy <- ckappa(passive_sample_combined[c("subj_animacy_A", "subj_animacy_B")])
kappa_animacy #0.9096219

kappa_gender <- ckappa(passive_sample_combined[c("subj_gender_A", "subj_gender_B")])
kappa_gender #0.8746109

kappa_agent_pp <- ckappa(passive_sample_combined[c("agent_pp_A", "agent_pp_B")])
kappa_agent_pp #0.9118457

kappa_modal <- ckappa(passive_sample_combined[c("modal_A", "modal_B")])
kappa_modal #0.9469328

kappa_tam <- ckappa(passive_sample_combined[c("tam_A", "tam_B")])
kappa_tam  #0.7702391

kappa_cluster <- ckappa(passive_sample_combined[c("cluster_A", "cluster_B")])
kappa_cluster #0.8326557

kappa_adversativity <- ckappa(passive_sample_combined[c("adversativity_A", "adversativity_B")])
kappa_adversativity #0.6700877

kappa_values <- c(kappa_animacy$kappa, kappa_agent_pp$kappa, kappa_modal$kappa, kappa_tam$kappa, kappa_cluster$kappa, kappa_adversativity$kappa)

mean_kappa <- mean(kappa_values)
mean_kappa #0.8402305, very high agreement :)

```

The inter-rater agreement is high at 0.84. Find table to compare

## Logistic Regression

Several corpus studies on voice choice in French have shown that the choice between active and passive constructions (*être* passive) is not purely syntactic, but that semantic and pragmatic factors also play a crucial role in determining whether a passive construction is favoured.

-   Subjects in passive sentences are preferably animate, definite, and pronominal. In the study by @da_cunha_lalternance_2020, animate, definite, and pronominal arguments that function as the objects in active sentences were more frequently found in the passive (and vice versa). -\> prominence / informational structure as significant factor. (And: constituent length, shorter constituents are preferred in subject position and favour the passive voice)

### Subject responsibility

As indicated above, the question whether the referent of the subject in the se faire construction holds any responsibility is central to the debate of the construction's passive-status (see above; @sveberg_se_2012, @bellec_construction_2014 [212-213]).

<!-- whether the passive reading of se faire is a type of the causative reading or its own instance -->

```{=html}
<!--  However, there is general consensus that
the PSF construction has a preference to mark the subject as an active participant (although according to @raineri_analyse_2010 [98] is is not "universally accepted" as some authors also name counterexamples @kupferman_construction_1995 [59].)
In @raineri_analyse_2010 [95]'s view, the idea that the PSF expresses subject responsibility is the "most common hypothesis" ("l'hypothèse la plus répandue") in the literature. -->
```

In line with their argument that the passive reading of se faire + INF is a kind of causative, @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987 state that the subject referent of the se faire passive is always a causer and therefore always responsible to some degree. This is contrasted to the être passive's subject referent, which is see as more neutral. @gaatone_desagreable_1983 [164] distinguishes between an "active patient" in the se faire construction and a "passive patient" in the être passive (@raineri_analyse_2010 [96]).

In contrast, @kupferman_construction_1995 [59] gives PSF examples where the subject cannot be interpreted as responsible, one of which is repeated here:

```{r gloss}
#| echo: false
gloss_non_resp <- as_gloss(
  "elle sanglotait parce que son tout jeune bébé venait de se faire enlever les amygdales.",
translation = "She sobbed because her little baby got operated on for tonsilitis.",
source = "Kupferman (1995: 59)"
)
gloss_non_resp
```

According to him, subject responsibility emerges from the context and not from the PSF, which is inherently passive [98, 99]. Further support for this idea comes from @labelle_anticausativizing_2013 and @washio_when_1993. For @martin_construction_2002, subject dynamism, but not subject intentionality, is a necessary condition in the PSF (@raineri_analyse_2010 [90]).

Thus, although the requirement of subject responsibility in the PSF is not "universally accepted" (@raineri_analyse_2010 [98], "une catégorie hétérogène complexe" (@raineri_analyse_2010 [249]), the analyses show at least some consensus in that the PSF construction prefers to mark the subject as connected to the event (@raineri_analyse_2010), either through affectedness, responsibility or intentionality. This tendency does not exist for the être passive (source).

**Animacy**

Corpus studies have shown that animacy significantly influences voice choice in French. The animacy of the passive subject (not agent) has been shown to increase the likelihood of using the passive voice (*être* passive) over the active voice. @da_cunha_lalternance_2020 (+ @da2022alternance) -\> more explanation as to why

As implied in the debate on subject responsibility, subjects of se faire passive constructions are predominantly active participants in the described action. Since inanimate subjects are rarely affected and even less often responsible, PSF is common with animate subjects. While some scholars claim that only animate subjects are possible with the PSF, others provide counterexamples (@labelle_anticausativizing_2013, @kokutani_sur_2005, @novakova_construction_2009 \[4\], @sveberg_se_2012) (@raineri_analyse_2010 [92]).

However, these examples appear to be exceptions as there is both corpus (@novakova_construction_2009 [111]) and experimental (@koenig_semantic_1992 \[151, 153\]) evidence that subject animacy and PSF are positively correlated: In @novakova_construction_2009 [111]'s corpus study, less than 3% of the se faire sentences had an inanimate subjects. In a survey by @koenig_semantic_1992, French native speakers rated the acceptability of PSF with animate subjects significantly higher than with inanimate subjects. Inanimate subjecta were only allowed by some speakers when the subject "affects a relevant animate entity" (@koenig_semantic_1992 [151, 153]. The preponderance of animate subjects for the se faire passive seems to be one of the few aspects on which scholars agree (@raineri_analyse_2010 [92]).

We tag animacy as binary "inanimate"/ "animate" although we recognize that animacy has been considered as a hierarchical scale in the literature, most prominently by @zaenen_animacy_2004.

**Hypothesis**: Both *se faire* and *être* passive favour animate subjects over inanimate subjects but subject animacy is a stronger predictor of the PSF than the être passive.

**Interaction**: @raineri_analyse_2010 [94] suspects that the PSF's animacy (almost-) constraint restricts verb selection and provides an explanation for the construction's incompatibility with certain verbs (i.e. not all verbs can occur with animate subjects).

<!-- The preference of the se faire passive of animate over inanimate subjects is parallel to the English get passive (@hundt_be-_2024), although @raineri_analyse_2010 [92] finds that the get passive allows more inanimate subjects than the se faire passive. According to @bohmann_large-scale_2023 [43], the get passive has become more likely with inanimate subjects over time which may signify a lessening of subject responsibility. @raineri_analyse_2010 brings up the possibility that a similar process of grammaticalisation may be happening in French. (not sure if relevant)-->

**Gender**

Since previous research suggests that grammatical gender plays a role in voice choice, it is worth investigating whether this effect extends to the alternation between *être* and *se faire* passives.

@da_cunha_lalternance_2020 (+2022) found that in written French, a feminine agent increased the likelihood of *être* passives (in the passif court model, i.e. when the agent is omitted) over the active voice, with no interaction with animacy (whether the argument was human or non-human) \[or definiteness\]. This suggests that gender itself may directly influence voice choice. However, given the study’s small sample size, it is also possible that an interaction effect exists but was not statistically significant \[check 2022 study\].

Since we do not infer the agent from context in this study, our analysis focuses solely on the subject’s gender. If *se faire* passives favour subjects with high responsibility, salience or affectedness, the grammatical gender of the subject may influence the alternation between *se faire* and *être* passive. Specifically, if feminine referents are perceived as less active or agentive in certain contexts, a feminine subject may favour *être* passives over *se faire* passives.

To test this, we categorise gender into four levels: unknown, mixed, masculine, and feminine. Our hypothesis is that a feminine subject will favour the *être* passive and disfavor the *se faire* passive. We will test the potential interaction between gender and animacy, as the effect of grammatical gender may depend on whether the subject is animate or inanimate, potentially linking to the social gender of animate subject referents. Additionally, we will examine how unknown or mixed-gender subjects behave in terms of this alternation.

**Complément d'agent (Agent PP)**

(The par-phrase introduces an external agent into the passive construction: It denotes the causer/ instigator of the action and is often derived through diathesis from the subject of the active equivalent.)

We expect the presence or absence of the par-phrase to influence passive choice. The par-phrase introduces an external agent who is likely the driving force behind the action described: @kupferman_construction_1995 postulates an "agentivity requirement" for agents denoted by the par-phrase, attributing a strong responsibility to the component. @hundt_be-_2024 and @bohmann_large-scale_2023 found that in English, the Agent PP reduces subject responsibility by introducing an agentive/ responsible entity.

We hypothesize that the complément d'agent influences the choice between *se faire* and *être* passives. The presence of a complément d'agent should favor *être* passives, as it introduces a responsible external argument and shifts responsibility away from the subject. In contrast, the absence of a complément d'agent is expected to favor *se faire* passives, which imply some degree of subject responsibility.

We expect phrases with *de* to have an even stronger effect than phrases with *par*, as they typically refer to less agentive or abstract sources (instead of agents), contradicting the dynamic nature of the *se faire* passive, which typically occurs with adversative or strongly agentive actions. To examine this, we will test the interaction between the complément d'agent and the verb cluster??, as we expect the *se faire* passive to be favoured by dynamic and adversative verbs (see seq-XXX).

**Modal verbs** (auxiliary)

such as "devoir", "savoir", "pouvoir", "falloir", "vouloir" very frequent with se faire passive @novakova_construction_2009 \[9\] (in @raineri_analyse_2010 [97], reason: they denote "control" to the subject

-   modal category limited to/ only comprises modal verbs for the scope of our study, not other similar constructions denoting subject responsibility (e.g. avoir besoin de, arriver à, réussir à, envier de, avoir des difficultés pour, avoir la possibilité de, avoir envie de, ...)

### Aspectuality

Regarding the factor of aspectuality, @keenan_passive_2007 [340] cites one of Greenberg's universals:

"If a language has two or more basic passives they are likely to differ semantically with respect to the aspect ranges they cover." \[Greenberg source with page number\]

If we assume that the PSF is a "basic passive", this suggests that the *être* and *se faire* passives have different aspectual preferences. Aspectuality is expressed through the verb, either via inherent lexical semantics ("Aktionsart") or grammatical marking, such as tense, aspect, and mood [^3] (source missing!).

[^3]: The general consensus in the literature is that aspect marks progressivity grammatically and Aktionsart semantically/ lexically. However, this clear distinction between aspect and Aktionart has been doubted, and some theories see Aktionsart und Aspect as two instantiations of Aspectuality (source missing!).

We do not examine mood in this study.

-   **Verb semantics (Aktionsart), Clusters** (@bohmann_large-scale_2023 [51-52]:

    Multiple hypotheses are found in the literature concerning the compatibility of certain verbs with the PSF. For example, @kupferman_construction_1995 \[67\] and @raineri_analyse_2010 [90] state that *se faire* passive works well with dynamic verbs but not with stative verbs, although exceptions are provided by @novakova_construction_2009); @martin_construction_2002 and @raineri_analyse_2010 [100-103].

    -\> The *se faire* construction is generally considered incompatible with stative verbs and instead tends to occur with dynamic processes, focusing on cases where the source role is agentive and the target role is affected, which is the typical case of semantic transitivity. (@raineri_analyse_2010: 90, 238).

    -\> @kupferman_construction_1995 argues that the *se faire* construction excludes stative verbs and verbs of state change with an external ‘cause’ argument, which explains differences in sentence acceptability (e.g. *Le pauvre s'est fait étrangler par le tueur* vs. \**Le pauvre s'est fait étrangler par cette cordelette*).

    -\> However, @martin_construction_2002 disagrees and explains that the source role can cause a change of state without being intentional, calling it an ‘effector’ rather than an ‘agent’. She cites examples such as *Isadora Duncan s'est fait étrangler par son écharpe* to support her view. She also claims that *se faire* constructions with psychological verbs (e.g. *Elle s'est fait séduire par Pierre*) are not always causative. @novakova_construction_2009 also notes that ‘*se faire* constructions with stative verbs can, although rarely, be interpreted passively, as in *Il peine encore à se faire aimer par tous les siens.*

-   volitional verbs fit well with PSF because they involve an action that leads to a mental state change while non-volitional verbs do not

-   Telicity seems to be another important distinction: According to @bellec_construction_2014 \[209\], the *être* passive is likely to describe a result or mark an endpoint of an action, while the *se faire* passive is used for actions in progress and verbs that are in their nature ongoing (e.g. perception verbs, @bellec_construction_2014 [209]). This could be reflected in the telicity of the verb, with telic verbs preferring the *être* passive and atelic verbs favouring the *se faire* passive.

-   Semantic class of the verb (lemma), using the dictionary *Les Verbes Français - LVF* (Dubois&Dubois-Charlier 1997) (in line with @da2022alternance)

-   @da2022alternance showed that the semantic class of the verb plays a crucial role in the use of the passive voice. In their study, 85% of *frapper* class verbs occurred in passive constructions, while only 15% were used in the active voice. This is consistent with findings from journalistic English, where verbs describing violent actions are often used in the passive voice (source). According to @da2022alternance, this preference can be explained by two factors: (1) the topical character of the affected patient, which makes it more likely to be promoted to the subject position, and (2) the possibility of the omission of an unknown or less salient agent.

    We expect this effect to hold for the *se faire* passive as well. Even though both the *se faire* and *être* passives topicalise the patient, the *se faire* passive typically implies some degree of subject involvement, so highly dynamic (and adversative) verbs such as *frapper* might increase the likelihood that it is favoured over the *être* passive, in which the subject is a more passive patient. However, this effect could interact with other factors ??

    @da2022alternance: while the preference for active constructions was not statistically significant for location-related verbs, the probability of communication verbs occurring in the active was significantly higher. Some classes showed no clear effect on passive frequency, with approximately equal distributions of active and passive forms (e.g. verbs of perception, cognition, psychological states, behaviour, movement and transfer). Locative and communicative verbs, on the other hand, appear more frequently in the active voice. One possible explanation for this tendency is verbal disposition? (Stallings, MacDonald & O'Seaghdha 1998). According to this hypothesis, verbs show a preference for certain syntactic constructions that speakers store in their mental lexicon. Communicative verbs often occur with object clauses, which rarely function as subjects. Since the active voice allows these verbs to maintain a direct object, they naturally favour active constructions over passive ones. A similar argument was made by Thuilier (2012a) in relation to communicative verbs and postverbal complements. Moreover, communicative verbs often involve an animate or human argument, which tends to be encoded as the subject in active constructions.

    -   Hitting-touching verbs (highly dynamic, adversative): strong preference for the passive voice, might also prefer se faire passives, contexts that emphasise the subject's affectedness

    -   Communication verbs (favouring the active voice): strongly favour the active construction due to their syntactic disposition and their tendency to use object clauses, may be less compatible with se faire passives, presence of an animate or human subject of the corresponding active construction may further reduce the likelihood of se faire

    -   Neutral verb classes (e.g. psychological verbs, behavioural, movement, transfer): do not show a strong passive preference, distribution on être and se faire passives may depend on additional factors

-   se faire passives might be particularly favoured by highly dynamic hitting-touching verbs such as *frapper*, while verb classes that favour active constructions (e.g. communicative verbs) might be less compatible.

-   (OR merge communication verbs and mental and local states into one category that prefers être over se faire and introduce another category for movement verbs (frequently occuring with se faire according to @raineri_analyse_2010 [90]: "les verbes de mouvement "aller" et "venir", qui mettent en relief la phase initiale du procès (Spang-Hanssen 1967 : 143)136 et les verbes de perception (Novakova 2008 : 9), qui impliquent un procès duratif"); suggested categories:

    -   frapper-type verbs (summarise as: "immediate physical actions" or "applied force" after @bohmann_large-scale_2023 ?)

    -   movement: going, coming, entering, leaving etc.

    -   communication & mental states: emotions etc.

    -   neutral)

+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| Verb cluster                                                      | hypothesised passive preference | Verb examples                                                                                                                                                                                                                       |
+===================================================================+=================================+=====================================================================================================================================================================================================================================+
| giving, depriving (don, privation)                                | neutral                         | aider (*une personne victime, à 57 ans, d'' un accident entraînant l'' apparition d'' un handicap et qui se ferait aider pendant dix, quinze, voire vingt ans par son conjoint sans demander la PCH.*)                              |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| entering, leaving (entrée, sortie)                                | neutral                         | engager (*cela répond à votre attente à l'' égard de fonctionnaires qui quitteraient l'' administration pour se faire engager par ailleurs au moyen de contrats de forme différente.*)                                              |
|                                                                   |                                 |                                                                                                                                                                                                                                     |
|                                                                   |                                 | transplanter (*Afin de protéger nos concitoyens de l'' exposition aux risques sanitaires qu'' ils encourraient en se faisant transplanter à l'' étranger puis en revenant en France sans forcément recourir aux soins nécessaires*) |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| hitting, touching (frapper, toucher)                              | se faire                        | tuer (*ces hommes que l'' on avait tant de difficulté à intégrer furent mobilisés pour aller combattre et se faire tuer au nom de notre pays*)                                                                                      |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| movement in place (mouvement sur place)                           | neutral                         |                                                                                                                                                                                                                                     |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| equipping, disarming (munir, démunir)                             | neutral                         | vacciner (*Il faut une prescription pour se faire vacciner contre la grippe en pharmacie*)                                                                                                                                          |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| transformation, change (transformation, changement)               | neutral                         | remplacer (*afin que le premier puisse se faire remplacer par le second*)                                                                                                                                                           |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| realisation, preparation (réalisation, mise en état)              | neutral                         | photographier (*J '' ajoute la diminution du nombre de points où l'' on peut se faire photographier et déposer son dossier en vue de faire établir ou renouveler sa carte nationale d'' identité ou son passeport*.)                |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| physical states and behaviours (états physiques et comportements) | neutral                         | aider (*afin de permettre aux enfants victimes de haine en ligne de se faire aider même si leurs parents n'' en sont pas informés.*)                                                                                                |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| locative (locatif)                                                | neutral                         | accepter (*Oliver Wendell Holmes écrit ainsi que " la meilleure épreuve de vérité est le pouvoir d'' une pensée de se faire accepter dans la compétition du marché " – competition of the market.*)                                 |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| union, reunion (union, réunion)                                   | neutral                         |                                                                                                                                                                                                                                     |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| psychological verbs (verbes psychologiques)                       | neutral                         | écouter (*Le problème , on l' a dit, est de savoir comment mettre vraiment cela en pratique lorsque l'on constate une difficulté à se faire écouter.*)                                                                              |
|                                                                   |                                 |                                                                                                                                                                                                                                     |
|                                                                   |                                 | suivre (*les patientes seront contraintes de multiplier les déplacements pour se faire suivre.*)                                                                                                                                    |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| seizing, clasping, possessing (saisir, serrer, posséder)          | neutral                         | élire (*avant que M. Hollande ne se fasse élire en affirmant qu'' il ne signerait pas le TSCG en l'' état.*)                                                                                                                        |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| communication                                                     | être                            | accuser (*il est devenu impossible de l'' exprimer dans le débat public sans se faire accuser de haïr les policiers*)                                                                                                               |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+
| auxiliary verbs                                                   | \-                              | \-                                                                                                                                                                                                                                  |
+-------------------------------------------------------------------+---------------------------------+-------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------------+

: Hypothesised passive preference for verbs based on clusters/ classes by @francois_presentation_2007 [4]

We have used the classes *communication* and *hitting-touching* from the lexicon in the annotation, as we expect communication verbs to favour the être passive and hitting-touching verbs to favour se faire passives. We expect all other classes to not have a preference for either passive variant and therefore group them together as neutral. If there was no exact match, we used the closest similar class. In one case (*caillasser*) there was no entry in the lexicon: Here we assigned the category hitting-touching (similar to *attaquer, frapper, bombarder, ...*)

**Aspect and Tense Marking**

The tense and aspect markings French cannot be clearly distinguished from each other because they heavily influence each other and depend on each other.

According to @raineri_analyse_2010 [365], the "type of process denoted by the verb" is a "necessary condition". inchoative/ durative aspect: "get passive and se faire passive focus on onset of the process that affects the subject referent, whereas be and être passive present the process in its entirety, without offering an internal perspective." ?

No direct effect of tense, tense and aspect are not entirely separable categories but are functionally interdependent. Tense can serve as an indicator of aspect in terms of how it sets the temporal focus of a sentence. Constructions in the passé composé tend to focus on the completed event, which could increase the likelihood of the *être* passive over the *se faire* passive. In contrast, the imparfait, which emphasises the ongoing process, might favour the *se faire* passive. In the present tense, we expect both passives to be evenly distributed, as the present tense emphasises neither the end point nor the process. Descriptions of future events tend to emphasise the beginning and ongoing process of an action rather than its end point (@bellec_construction_2014 [210]). We therefore expect that in the future tense, the *se faire* passive is favoured over the *être* passive.

According to @spang-hanssen_quelques_1967 \[141\] and @novakova_construction_2009, periphrases that denote progressivity, such as *être en train de* + INF, *venir de* + INF and *aller* + INF, occur more naturally with the *se faire* passive than with the *être* passive. This is because the *se faire* passive conveys a dynamic, ongoing action (Quelle) compared to the more stative *être* passive. (also: @bellec_construction_2014; @keenan_passive_2007 [341], @raineri_analyse_2010 [91]

We decide against coding tense, aspect (and mood?) as separate factors, but to consider them in the context of a functional category that reflects the difference between process focus (progressive) and result focus (non-progressive). Therefore, we establish a TAM-category with two levels: progressive and non-progressive. The descision to not keep the categories separately is based on findings by @bellec_construction_2014 \[210\] who argues that aspectuality and tense are closely intertwined and together are responsible for whether a sentence is understood to be progressive or non-progressive. This marking reflects the more functional nature of category distinction in French rather than the formal distinction in English for example. And even in English: @bohmann_large-scale_2023 [39]: "Form of auxiliary verb: present, preterite, perfect, infinitive, -ing" OR progressive/non-progressive) -\> only effect found by @bohmann_large-scale_2023 \[49\] that present and preterite disfavour get passive. @hundt_be-_2024 [97] does tag tense and aspect separately but later, in the interpretation, interpret them together in terms of a tam category (result: se faire favours progressive but non-significant effect).

(sentence for conditional marking: Subjonctif and conditional forms excluded because of CQL? Or: We mark them according to the tense they occur with)

**levels**:

-   "progressive": imparfait, futur, periphrase constructions (called "progressive but includes temse)

-   "non-progressive": passé composé, plus-que-parfait, present tense without periphrase construction

    -   \+ imparfait referring to states?

! & "**se faisant remplacer" -\> progressive form**

Progressive category: emphasises ongoing or continuous actions, includes the imparfait, futur (except for futur antérieur - "I will have..."), and periphrastic constructions like *être en train de* + INF, *venir de* + INF or *aller* + INF.

Non-progressive category: completed actions, includes the passé composé, plus-que-parfait, and the present tense without progressive constructions.

Our hypothesis is that progressive marking (whether through tense or periphrasis) favours the *se faire* passive, while non-progressive marking (e.g., passé composé, plus-que-parfait) favours the *être* passive.

### Adversativity/ Non-neutrality

The factor of "Adversativity/ Non-neutrality" is reflected in the semantics of the verb and captures the (negative) consequences of the described action for the subject. It has been found to be relevant to the choice between the be and get passive (@hundt_be-_2024 [89, 96-97]; @bohmann_large-scale_2023). @spang-hanssen_quelques_1967 and @raineri_analyse_2010 draw parallels between the English get passive and the French PSF construction.

@gaatone_desagreable_1983 argues that the se faire passive has a strong "unpleasant sense" ("sens désagreable") because it occurs mainly with verbs denoting unpleasant events [^4]. Although @kupferman_construction_1995 \[77\] rejects the "unpleasant sense" as a characteristic of se faire, he agrees that it is more commonly used in unpleasant contexts ("has tended to specialize in unpleasant roles" (compared to être, which is more neutral)). This is supported by corpus data: @novakova_construction_2009 [110] found a clear preference for PSF to be realised with unpleasant actions, with 30% of verbs in PSF denoting unpleasant events and only 5% denoting pleasant events. Also: @raineri_analyse_2010 [247].

[^4]: According to @gaatone_desagreable_1983, unpleasant events contradict the notion of subject responsibility because subjects would not willingly bring the unpleasant event upon themselves (argument for "primacy of causative reading"). According to him, the reason this is the subject's involvement in se faire constructions: if the subject experiences a pleasant event and this is expressed by se faire -\> mostly the subject has initiated the action -\> causative and not passive reading (language example?)/ in contrast, it is less likely that the subject initiates an unpleasant action.

In their corpus study on the English be - get passive alternation, @hundt_be-_2024 label it as binary on the verb as adversative, non-adversative/neutral. For practicality and clear statistical modelling, We follow them in our annotation: "Adversative" if event negatively affects the subject, "non-adversative" if event is neutral or positive.

<!-- -   \@bohmann_large-scale_2023 [38, 43] Sentiment values of verb participles extracted from SentiWordNet: positivity, negativity, objectivity scores; reversed if sentence is negated -->

As @bohmann_large-scale_2023 notes, there are two hypotheses related to emotions of the verb: The adversativity hypothesis and the non-neutrality hypothesis. Both expect the *get* passive to favour verbs with negative scores; but only the non-neutrality hypothesis predicts get passive to correlate positively with positive emotion scores as well. Since positive emotion verbs have been shown to be quite rare with PSF (@novakova_construction_2009 [110]) we have chosen to not test them/ put them together with "neutral".

**Hypothesis:** *se faire* passive is preferred over the *être* passive for verbs that denote "unpleasant" actions (as reflected by verbs labeled as "adversative").

**Interaction:** Affectedness of the subject may be reflected by an interaction of subject animacy and verb adversativity: Negative consequences of actions, expressed by adversative verbs, affect animate subjects more than inanimate subjects.

## Memory based learning

-   non-parametric: no assumptions about underlying distribution of data
-   no problem with sparse observation of features/ low-frequency events @theijssen2013choosing [243, 244] because it "does not rely on any kind of generative model"
-   only mandatory annotation is class label @theijssen2013choosing [243] (passive choice)
-   "simple-theory neutral features"/ or some linguistic features (animacy, verb, pp-phrase...)
-   compares test instances with labeled training data (stored in the memory)
-   MBL very susceptible to majority classes -\> unbalanced data is a big problem (poor performance for minority classes possible)
-   k-Nearest-Neighbor (kNN):
    -   kNN algorithm is trained on data to predict the most likely passive choice
    -   nearest neighbours out of all instances are determined -\> observation receives class of majority of nearest neighbours
    -   leave-one-out-setting @theijssen2013choosing [244]: train on all but one, test on remaining instance, iterate for each instance
    -   hyperparameters setting (e.g. number of neighbors)? -\> help of wrapper
    -   Tutorial: <https://www.datacamp.com/tutorial/k-nearest-neighbors-knn-classification-with-r-tutorial>

1.  choose the value of K, which is the number of nearest neighbors that will be used to make the prediction.
2.  Calculate the distance between that point and all the points in the training set.
3.  Select the K nearest neighbors based on the distances calculated.
4.  Assign the label of the majority class to the new data point.
5.  Repeat steps 2 to 4 for all the data points in the test set.
6.  Evaluate the accuracy of the algorithm.

-   in R: "class" or "caret" package
-   set seed for reproducibility
-   split set manually or by "caTools" library
-   Preprocessing/ Feature scaling if feautres have different scales, to ensure proportions between features and prevent that features with largest scale dominate distance calculations
-   caret: Model Tuning: find optimal value of "K"
-   caret: train best performing model
-   Model evaluation

# Results

## Data preparation

For the main analysis, the se faire and être data were merged into one "clean" dataset where each column corresponds to one factor. This column includes the annotation of the first annotator (annotator A for half of the observations, annotator B for the other half).

```{r merge-to-main-dataset}
#goal: one dataset with one column for each factor, the column includes the annotation of the "main" annotator
# special case agent pp column: if passive variant is se faire -> take value from column "agent pp_A", if passive variant is être is: same procedure as with the other variables
passive_sample_clean <- passive_sample_combined |> 
  mutate(
    subj_animacy = case_when(annotator == "A" ~ subj_animacy_A,
                             annotator == "B" ~ subj_animacy_B),
    subj_gender = case_when(annotator == "A" ~ subj_gender_A,
                            annotator == "B" ~subj_gender_B),
    agent_pp = case_when(passive_variant == "se faire" ~ agent_pp_A,
                         passive_variant == "être" ~ case_when(
                           annotator == "A" ~ agent_pp_A,
                           annotator == "B" ~ agent_pp_B)),
    modal = case_when(annotator == "A" ~ modal_A,
                      annotator == "B" ~ modal_B),
    tam = case_when(annotator == "A" ~ tam_A, 
                    annotator == "B" ~ tam_B),
    cluster = case_when(annotator == "A" ~ cluster_A, 
                        annotator == "B" ~ cluster_B),
    adversativity = case_when(annotator == "A" ~ adversativity_A, 
                               annotator == "B" ~ adversativity_B)
  )

# remove now-redundant A and B columns, remove ID (irrelevant since it's only there for être)
passive_sample_clean <- passive_sample_clean |> 
  select(-subj_animacy_A, -subj_animacy_B, -subj_gender_A, -subj_gender_B, -agent_pp_A, -agent_pp_B, -modal_A, -modal_B, -tam_A, -tam_B, -cluster_A, -cluster_B, -adversativity_A, -adversativity_B, -ID)
```

## Data exploration

Taking a look at the data:

```{r data-exploration-check, include=FALSE}
str(passive_sample_clean)
head(passive_sample_clean)
```

We only have categorical data; therefore mean and standard deviation are not meaningful. Descriptive statistics for each variable (following Elens textbook Chapter 8 & 10 and @gries_statistics_2021 [441]

-   table to get an overview of distributions of values and the mode (measure of central tendency) -\> sort

-   where appropiate: cross tabulations

-   proportions

-   bar plots, mosaic plots

### Subject animacy

```{r animacy-descriptive-tables}
table(passive_sample_clean$subj_animacy) |> 
  sort(decreasing = TRUE) |> 
  proportions ()

#cross-tabulation for hypothesised interactions
table(passive_sample_clean$subj_animacy, passive_sample_clean$passive_variant)

# not really helpful/ meaningful:

# table(passive_sample_clean$subj_animacy, passive_sample_clean$subj_gender)
# 
# table(passive_sample_clean$subj_animacy, passive_sample_clean$cluster)
# 
# table(passive_sample_clean$subj_animacy, passive_sample_clean$adversativity)
# 
#  ftable(passive_sample_clean$passive_variant, passive_sample_clean$subj_animacy, passive_sample_clean$subj_gender) |> 
#    proportions()
# 
# ftable(passive_sample_clean$passive_variant, passive_sample_clean$subj_animacy, passive_sample_clean$cluster)
# 
# ftable(passive_sample_clean$passive_variant, passive_sample_clean$subj_animacy, passive_sample_clean$adversativity)

#proportions (Gries)
prop.table(table(passive_sample_clean$subj_animacy, passive_sample_clean$passive_variant))
#alternatively: (Elen)
#table(passive_sample_clean$subj_animacy) |> 
#  sort(decreasing = TRUE) |> 
#  proportions()
```

Overall, two-thirds of the constructions have animate subjects (mode: animate; 704 vs. 360 inanimate). However, in *être* passives, inanimate subjects are more frequent than animate ones (332 inanimate, 200 animate), whereas *se faire* passives show a strong preference for animate subjects (504 animate, 28 inanimate), as shown in these plots:

```{r animacy-descriptive-plots}
passive_sample_clean |> 
ggplot(mapping = aes(x = subj_animacy)) +
  geom_bar() +
  facet_wrap(~ passive_variant) +
  #scale_y_continuous(expand = c(0,0)) + 
  labs(x = "Subject Animacy",
       y = "Number of observations") +
  theme_bw()

mosaicplot(passive_variant ~ subj_animacy,
           data = passive_sample_clean)
```

The plot shows clearly that in *se faire* passives, the subject is animate in the great majority of cases. In contrast, for *être*, subject animacy is more evenly distributed, with a slight preference for inanimate subjects.

Animacy x Adversativity:

```{r animacy-adversativity-plots}
passive_sample_clean |> 
ggplot(mapping = aes(x = subj_animacy, fill = adversativity)) +
  geom_bar() +
  facet_wrap(~ passive_variant + adversativity) +
  #scale_y_continuous(expand = c(0,0)) + 
  labs(x = "Subject Animacy",
       y = "Number of observations", 
       fill = "Adversativity" ) +
  theme_bw()
```

For *se faire*, both adversative and non-adversative constructions more frequently occur with animate subjects. In contrast, for *être*, non-adversative constructions are more common with inanimate subjects than with animate ones. Adversative constructions are generally less frequent and fairly evenly distributed.

### Subject gender

```{r gender-descriptive}
table(passive_sample_clean$subj_gender) |> 
  sort(decreasing = TRUE)
#level "mixed" too rare? (11 occurences)
#cross-tabulation
table(passive_sample_clean$subj_gender, passive_sample_clean$passive_variant)
prop.table(table(passive_sample_clean$subj_gender))
```

```{r}

passive_sample_clean_filtered <- passive_sample_clean |> 
  filter(subj_gender != "mixed")

passive_sample_clean_filtered2 <- passive_sample_clean |> 
  filter(subj_gender != "mixed") |>
  filter(subj_gender != "unknown")

```

```{r gender-plot}

ggplot(passive_sample_clean_filtered, 
       aes(x = passive_variant, 
           fill = subj_gender)) +
  geom_bar(position = "dodge") +  # No need for stat="identity"
  scale_y_continuous(expand = c(0,0)) +
  labs(title = "plot",
       x = "passive variant grammatical gender",
       y = "Count") +  # Change y-axis label to "Count" 
  theme_classic()

```

```{r}

ggplot(passive_sample_clean_filtered2, 
       aes(x = subj_animacy, fill = subj_gender)) +
  geom_bar(position = "dodge") +  
  facet_wrap(~ passive_variant) +  
  labs(title = "Gender and Subject Animacy in the Passive Variants",
       x = "Subject Animacy",
       y = "Number of Observations", 
       fill = "Subject Gender") +
  theme_bw()
```

The barplots show that the *être* passive occurs most frequently with inanimate feminine subjects, while the *se faire* passive occurs most frequently with animate masculine subjects. This distribution supports the hypothesis that *se faire* is more frequent with animate male subjects than with animate female subjects.

For both passive variants, animate subjects are predominantly masculine, while inanimate subjects are more frequently feminine. However, the data for *se faire* with inanimate subjects is sparse.

For inanimate subjects, the difference in frequency between masculine and feminine gender is smaller compared to animate subjects, suggesting that the influence of gender on the choice of passive variant could be less strong for inanimate subjects than for animate ones.

### Complément d'agent

### Modal

### TAM

### Cluster

### Adversativity

Some mosaic plots for exploration:

```{r data-exploration-plots}
mosaicplot(passive_variant ~ subj_gender,
           data = passive_sample_clean)

mosaicplot(passive_variant ~ tam,
           data = passive_sample_clean)

mosaicplot(passive_variant ~ adversativity,
           data = passive_sample_clean)

mosaicplot(subj_gender ~ passive_variant,
           data = passive_sample_clean)

```

More in @levshina_how_2015: Chapter 4 and @gries_statistics_2021 Chapter 3/Descriptive statistics & visualisations) and Chapter 6 (Mixed effects regression modeling)

## Binomial logistic regression

```{r packages-regression, include=FALSE}
library(mclm)
library(kableExtra)
library(lmerTest)
library(afex)
library(ggeffects)
library(sjPlot)
#Levshina: Rling not available (Warning in install.packages : package ‘Rling’ is not available for this version of R); therefore no funktion "lrm"
library(rms)
library(visreg)
library(car)
```

@levshina_how_2015 [253] "logit represents the chances of construction A to be chosen in a particular type of context compared with the chances of construction B to be used in the same type of context."

intercept: "value that determines the chances of an outcome when all predictors are equal to zero (for quantitative variables) or the reference value (for categorical variables)".

### Reference levels

ordering levels: by default alphabetical; reference level always the first, i.e. (suggestion in brackets):

-   passive variant: être

-   subj_animacy: animate

-   subj_gender: feminine (better: unknown?)

-   agent_pp: 0

-   modal: no

-   tam: non-progressive

-   cluster: communication (better: neutral?)

-   adversativity: adversative (better: non-adversative?)

idea for reference levels: construct the most "standard" case: être passive, inanimate, unknown, modal: no, non-progressive, cluster: neutral, non-adversative ?

```{r reference-levels}
#exclude "mixed" gender because of low frequency (would likely skew the data)
passive_sample_clean <- passive_sample_clean|>
filter(subj_gender != "mixed") |>
droplevels()
#change reference levels: "animate" to "inanimate", "communication" to "neutral", "adversative" to "non-adversative" (Levshina 2015: 78)
passive_sample_clean$subj_animacy <- relevel(passive_sample_clean$subj_animacy, ref = "inanimate")
passive_sample_clean$subj_gender <- relevel(passive_sample_clean$subj_gender, ref = "unknown")
passive_sample_clean$cluster <- relevel(passive_sample_clean$cluster, ref = "neutral")
passive_sample_clean$adversativity <- relevel(passive_sample_clean$adversativity, ref = "non-adversative")

str(passive_sample_clean)
```

### Model selection: Maximum likelihood

#### glmer

Function following @gries_statistics_2021 [442](&%20Elens%20Scripts)

```{r regression-model, output=FALSE}
passive_glmer <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity) +
               (1 | lemma) + (1 | speaker),
    family = "binomial", 
    data = passive_sample_clean,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(passive_glmer) #AIC: 785.1 
tab_model(passive_glmer) # Marginal R2 / Conditional R2	0.598 / 0.746 // 0.704 / NA
vif(passive_glmer)
```

random effects: variance of the speaker very very small (1.359e-13), does not need to be kept; variance of the lemma is bigger (1.917e) and should be kept.

significant effects:

-   main effects: (subj gender\[unknown\]); cluster\[hitting-touching\]; cluster\[neutral\]

-   interactions: subj animacy \[inanimate\]× subj gender \[unknown\]; **subj animacy \[inanimate\]\
    × adversativity \[non-adversative\]**

removing non-significant interactions (cluster:adversativity and subj_animacy:cluster) and random effect of speaker:

```{r model-1, output=FALSE}
model_1 <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:adversativity) + (1 | lemma),
    family = "binomial", 
    data = passive_sample_clean,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_1) #AIC: 776.8(higher i.e. slightly worse?)
tab_model(model_1) # Marginal R2 / Conditional R2	0.601 / 0.750 (minimally higher, i.e. slightly better?)
```

removing non-significant fixed effects (agent_pp, modal, tam)

```{r model-2}
model_2 <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:adversativity) + (1 | lemma),
    family = "binomial", 
    data = passive_sample_clean,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_2) # AIC: 771.5 without level "mixed")
tab_model(model_2) # Marginal R2 / Conditional R2 0.599 / 0.748 (slightly lower than model_1 but with fewer factors!)
vif(model_2)
```

significant effects in model_2: cluster\[hitting-touching\]; cluster \[neutral\]; subj animacy \[inanimate\] × subj gender \[unknown\]; subj animacy \[inanimate\] × adversativity \[non-adversative\]

**Concordance index C**

```{r concordance-index}
#model_2
Hmisc::somers2(
  predict(model_2), as.numeric(passive_sample_clean$passive_variant)-1
)["C"] 
# C 0.9549516 - pretty good I think?
```

```{r coefficients, include=FALSE}
model_2_coef <- coefficients(summary(model_2))[,"Estimate"]

model_2_coef
#"a negative coefficient expressed as a logit (or log odds) shows that this outcome is less probably than the reference level"

#convert from log odds to odds:
exp(model_2_coef)

#convert odds to probabilities
exp(model_2_coef) / (1 + exp(model_2_coef))
```

Some plots of significant effects:

```{r model-2-plots}
ggpredict(model_2, "cluster") |> 
  plot()

ggpredict(model_2, c("subj_animacy", "subj_gender")) |>
  plot()

ggpredict(model_2, c("subj_animacy", "adversativity")) |> 
  plot()
```

Interpretation of subj_animacy x adversativity: for animate subject se faire is very likely, whether the construction is adversative or not does not make a huge difference. For inanimate subjects however, adversative constructions are realized more often with se faire while non-adversative constructions are less likely with se faire.

**Random intercepts**

```{r}
ranef <- as.data.frame(ranef(model_2))

ranef |> 
  filter(grpvar == "lemma") |>
  filter(condval < -0.8 | condval > 0.8) |> 
  ggplot(aes(x = reorder(grp, condval), y = condval)) +
  geom_point() +
  geom_errorbar(aes(ymin=condval-condsd, ymax=condval+condsd)) +
  labs(x = "Lemma", 
       y = "Variation in intercept estimate (logit)") +
  coord_flip()
```

#### lrm

Function following @levshina_how_2015 [258] (258-276) - no random effects!

```{r lrm-model}
passive_lrm <- lrm(
  passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity),
  data = passive_sample_clean
)

#no random effects
passive_lrm
#intercept: 2.2777
#convert to odds:
exp(2.2777) #9.75422
#meaning: chances of se faire are 9.75422 times greater than être given this context (default reference levels)
```

coefficients: if positive, level boosts the chance of se faire (decreases chance of être)

R-squared not recommended as central measure @levshina_how_2015 [259], Hosmer & Lemeshow (2000: 167)

# Discussion

-   interaction animacy x gender: *personne* (feminine) occurs frequently in the sample

# Conclusion/Outlook

-   Future studies: modality (spoken/written) as a factor (compare modalities), analyses on other modalities/varieties/registers

-   se faire passive more likely in informal settings, être passive more likely in formal settings

-   definiteness & pronominality of the subject

-   it might be interesting to look at the cases we excluded and see how they differ from real passive. (Grauzone: Passiv-Nuance, aber keine echte Alternation; Grade der Passivität untersuchen)

-   Limitation: Lemmas were determined by the lemmas occuring in se faire constructions

-   modal category only comprises modal verbs in our study, other similar constructions denoting subject responsibility could be interesting to examine (e.g. avoir besoin de, arriver à, réussir à, envier de, avoir des difficultés pour, avoir la possibilité de, avoir envie de, ...)

# References
