---
title: "Term paper"
subtitle: ""
author: "Rose Hörsting & Gina Reinhard"
date: "2025-03-15"
engine: knitr
bibliography: references.bib
format: 
  html:
    df-print: default
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
    number-sections: true
    embed-resources: true
link-citations: true
csl: "apa_7th.csl"
number-sections: true
---

```{=html}
<style>
p {
  text-align: justify;
}
</style>
```

```{r}
#| include: false
#install.packages("glossr")

library(glossr)
use_glossr(styling = list(
  source = "b",
  first = "",
  font_family = "Cambria"
))
```

# Introduction

-   rare construction: Only 5% of verbs in the French Treebank (FTB) are in the passive voice ( @da_cunha_lalternance_2020) vs. 9% in English for the same text type (@roland2007frequency).

# The French passive alternation of *être* and *se faire*

## Canonical passive with *être*

### Structure and meaning

-   Function of the passive construction (semantic focus):

-   Diathesis: Concept of the passive (passif long) as an inversion of the order/ syntactic function of the arguments (the subject argument and the object argument) in relation to the active construction. The object of the active "becomes" the subject in the passive. (Per default, in French the subject is coded at position 1 in the sentence)

-   The omission of the agent argument, called the *complément d'agent* (*par* or *de* phrase), is also possible. In this case, the construction is referred to as *passif court*, whereas the *passif long* explicitly includes the complément d'agent:

    *Language examples from our corpus*

    Approx. a quarter of all passive constructions in the French Treebank (FTB) are realised as passif long (@da_cunha_lalternance_2020), while in spoken French, the proportion is about 10% (source Hamma).

-   evtl.: passive vs. *construction moyenne en se* (*La porte a été ouverte* vs. *la porte s'est ouverte*) -\> In the passif court, the agent argument is still implicit (Quellen)

## *se faire* passive

```{=html}
<!-- ### Historical background

@bellec_grammaticalisation_2015 and @bellec_quand_2014 date the se faire + INF passive construction (PSF) to the middle of the 20th century and call it a "modern" construction that has since experienced a rapid increase in popularity. She describes a process of grammaticalisation and reanalysis for the PSF during/through which “se” has lost its reflexive meaning and “faire” has become an auxiliary. In merging, they form a “semi-auxiliary” and obtain a passive interpretation, while the causative meaning is lost (désémantisation). However, @bellec_construction_2014 claims that the causative reading did not disappear but instead coexists with the passive reading (debated, see below).

-   In contrast, @raineri_analyse_2010 [280-305] based on corpus data: sees origin much earlier, exists since origin of Middle French (14-17th century); NOT derived from causative reading, increased productivity from 19th century onward -->
```

### Structure and meaning

The construction consists of the pronoun "se", the conjugated form of the verb "faire" and a verb in the infinitive. According to @labelle_anticausativizing_2013 [235] the construction is "formally causative, but semantically passive". Similar to the être passive, the subject of the passive sentence is the experiencer of the action. As with être, the direct object of the active sentence becomes the subject of the passive sentence. *Language example: active sentence with two (animate) arguments, se faire passive sentence, evtl. etre passive sentence*

Unlike the être passive, however, the diathesis is not restricted to the direct object: The PSF has "greater syntactic flexibility" (@raineri_analyse_2010 [93-95] and can also be formed with ditransitive verbs, where the indirect object becomes the subject: *Language example: active ditransitive sentence, se faire passiv.* Another possibility are possesor constructions where the subject of the passive is only part of the object in the active sentence (Le coiffeur coupe les cheveux de Pierre. -\> Pierre s'est fait couper les cheveux). However, the passive status of these constructions is controversial, as discussed below/ in the next section. <!--Some could also be understood as causatives if the subject both initiates and undergoes the action. (Pierre made the hairdresser cut his hair) (add source)) -->

The peculiarity of the construction has led scholars to wonder: "how can a passive-like construction arise from a combination of a causative and a reflexive marker, where no marking of the passive function is apparent?" (@koenig_semantic_1992 [145]).

@labelle_anticausativizing_2013 [23] suggests that the passive meaning is obtained through a syntactic mechanism that involves binding the empty NP-object position of the embedded verb to the subject through the reflexive particle "se".

<!-- In the se faire passive, the embedded verb (i.e., the infinitive after faire) has an empty NP-object position since only the subject is realised. The reflexive clitic "se" binds the empty object position of the embedded verb. Therefore, "se" connects the (missing) object-NP and the surface subject (coindexation). In contrast to the être passive, the object does not become the new subject but is connected to the subject by "se". The subject is interpreted as affected because it is linked to the missing object of the embedded verb (which would be affected in the active sentence but is absent on the surface in se faire). For this to work, the embedded verb needs an agent as a subject (because otherwise, the missing object would not be affected). -->

Sentences including this construction can have either/both a passive and/or a causative sense:

*Language examples: passive se faire; causative se faire (same construction, different meaning)* @koenig_semantic_1992\[143\]

Crucially, passive se faire and causative se faire are structurally identical, differing only in their semantics. As @labelle_anticausativizing_2013\[236\] explains, the ambiguity is resolved by "the lexical items involved". In the causative reading, "se" is understood as reflexive and "faire" as causative. In contrast, if interpreted as a passive, the meanings of "se" and "faire" are not transparent. Thus, the subject referent does not necessarily cause the event and is not necessarily responsible (@koenig_semantic_1992\[143\]: "any notion of causation is absent"). Sometimes both readings are possible and the ambiguity cannot be resolved (@koenig_semantic_1992\[144\]), for example when cognition and perception verbs are involved (@bellec_construction_2014\[206, 207\]).

<!-- @kupferman_construction_1995 reasons that the PSF is a real, fully adequate passive construction, equivalent to the être passive. Others have different opinions... -->

### Debate of the construction's passive-status (vs. causative)

The above-described stance taken by @bellec_construction_2014, @labelle_anticausativizing_2013, @koenig_semantic_1992 and @kupferman_construction_1995 that the PSF is a full-fledged passive construction is by no means undisputed. The opposing view - that the construction is a type of causative - is also widely represented in the literature. It is most prominently presented by @tasmowski-de_ryck_causatif_1987 and @gaatone_desagreable_1983, and was more recently presented by @sveberg_se_2012. In their view, passive meaning emerges from verb semantics and pragmatics, not from the PSF construction itself. As @raineri_analyse_2010 [89] describes, there is also the possibility of a multitude of readings (@kokutani_sur_2005) and of a continuum between causative and passive (Hobak Haff 2000), later accepted by @raineri_analyse_2010 [154-166].

<!-- As @labelle_anticausativizing_2013 [15] explains: -->

@gaatone_desagreable_1983 as well as @tasmowski-de_ryck_causatif_1987 centre their argumentation that the se faire passive is a weak reading of the causative around the notion that the subject of the se faire construction "is always, even unconsciously, responsible for the occurrence of the embedded event."

On the other hand, @kupferman_construction_1995 sees the causative and passive readings as homonyms with fundamentally different semantics and therefore assigns an inherent passive status to the se faire construction. This agrees with @bellec_construction_2014, according to whom the se faire passive has been grammaticalised into a semi-auxiliary with non-transparent components and there is thus no reason for a causative reading.

@raineri_analyse_2010 \[356-357\] concludes that despite its productivity as a passive construction, the PSF is problematic because it neither fits into two homonyms nor functions as a single weak causative. @bellec_construction_2014[207] defines the following three criteria for a causative reading, which distinguish the causative from the passive interpretation:

-   In the causative, the subject referent is agentive and has the role of both causer and patient. The subject actively initiates what happens to them. In the passive, the subject referent has only the role of the patient and is affected by the event.
-   In the causative reading, the particle “se” has a reflexive meaning whereas in the passive reading, “se” is not reflexive.
-   In the causative reading, “faire” has transparent causative meaning. This causative meaning is lost in the passive reading, where “faire” has no inherent meaning. <!-- add: how will we distinguish them / have we distinguished them in our study -->

This debate has been central to the investigation of the se faire + INF constructions and cannot be left aside. However, even if the status of the PSF has been the subject of disagreement and discussion, the literature clearly shows that there is a passive reading associated with this construction. This is independent of whether the passive sense emerges from within the construction or from the components it occurs with, or whether it has an underlying causative meaning.

By comparing PSF and PE constructions, we may be able to shed light on this debate from a corpuslinguistic perspectivem by examining which factors contribute to the choice between the two passives. <!-- (-> exclude causative readings?). Can our findings contribute to the debate of whether se faire deserves passive-status? If not: no need to spell out the debate in detail. -->

## *être* vs. *se faire*: Competition or complementation? - Defining the variable context

Following @bohmann_large-scale_2023 \[35\] who claim that "only cases where both variants can be used interchangeably should be considered", we aim to exclude constructions that do not allow for a paraphrase with the other. The passive constructions of se faire and être are not always interchangeable: As described above, the canonical passive with être requires a direct object in the active equivalent that becomes the new subject, @koenig_semantic_1992 [144], @washio_when_1993, @labelle_anticausativizing_2013 [16]. Thus, for reflexive and ditransitive verbs, the être passive is not available and does not compete with se faire. These constructions will be excluded and only transitive verbs will be studied.

Since our aim is to investigate the passive alternation of être and se faire, it is essential to ensure that we only compare the être passive with real passive se faire constructions, and not with causative constructions (@bohmann_large-scale_2023 \[35\]: "all extracted forms should be genuine instances of the variable under investigation"). We follow @raineri_analyse_2010 [165] in including only se faire + Inf constructions that have a natural paraphrase with the être passive in order to include only constructions that directly contrast with être.

-   se faire passive has evolved to be used in situations without a causer (passive reading) @bellec_construction_2014
-   figurative, "besoin d’expressivité and extravagance" instead of "used up" canonical passive @bellec_construction_2014 \[206\]; part of lexicalised expressions @novakova_construction_2009 [117]/ idioms @raineri_analyse_2010 [165]
-   while some scholars speak of semantic differences between the canonical être passive and the se faire passive, @kupferman_construction_1995 sees them as semantically synonymous -\> therefore: competition
-   how much is it used today (according to authors/ background literature) and in which settings
    -   se faire less formal, more figurative and emotional, used to build the passive with indirect object as the agent & to express passivity in dynamic/ ongoing processes, more spoken
    -   etre passive more “neutral”, more formal, more written
-   être: exclude adjectives and some participles that do not form the passive (similar to @bohmann_large-scale_2023 \[36\])/ "instances that allow for gradation" @hundt_be-_2024 [92]

### Maybe: Comparison to similar alternations in other languages

-   English @bohmann_large-scale_2023, East Asian languages @washio_when_1993 and @guan_multifactorial_2024 , Swedish @cuypere_corpus-based_2014
-   se faire passive "typologically marked", comparison to other Romance languages @koenig_semantic_1992 \[145\]; similar constructions exist in other Romance languages but passive reading only allowed in French @labelle_anticausativizing_2013 [236]

# Data and Methods

## Data Collection

-   corpus:
-   problem: data are extremely unbalanced (ratio in Parlatalk 1.7 Mio être vs. 3888 se faire (450:1)) -\> making the ratio 50: 50 would distort the reality significantly but if the ratio is kept the model will almost always predict être
-   strategy possibilities:
    -   build a corpus where être is presented more than se faire but not as extreme (approximation, e.g. 34% se faire - 60% etre) ("over/undersampling within limits")
    -   assign more weight to se faire constructions by "weights=" variable of glm function
    -   @bohmann_large-scale_2023 kept the imbalance (ca. 3 Mio be passive vs. 25 000 get passive)
    -   evaluate with precision & recall, not just accuracy (F1-score)

## Logistic Regression

Variables, Hypotheses, Preference constraints

### XXXXXX

Several corpus studies on the active/passive alternation in French have shown that the choice between active and passive constructions is not purely syntactic, but that semantic and pragmatic factors also play a crucial role in determining whether a passive construction is favoured.

-   Subjects in passive sentences are preferably animate, definite, and pronominal. In the study by @da_cunha_lalternance_2020, animate, definite, and pronominal arguments that function as the objects in active sentences were more frequently found in the passive (and vice versa). -\> prominence / informational structure as significant factor. (And: constituent length, shorter constituents are preferred in subject position and favour the passive voice)

-   Animacy has been shown to play an important role in the active/passive alternation in French.

-   Interestingly, gender also seems to play a role in the active/passive alternation. In @da_cunha_lalternance_2020 the passif court model showed a significant effect of grammatical gender: a feminine agent increased the likelihood of the the passive.

-   Surprisingly, number also had an effect: a plural arguments favoured the passive in all models. (+ Relativsätze & Priming als signifikante Faktoren) @da_cunha_lalternance_2020

### Subject responsibility

As indicated above, subject responsibility in the se faire passive is central to the debate about the construction's passive-status (see above; @sveberg_se_2012, @bellec_construction_2014 [212-213]) <!-- whether the passive reading of se faire is a type of the causative reading or its own instance --> <!-- As indicated above, the question whether the referent of the subject in the se faire construction holds any responsibility is central in the debate of the construction's passive-status. However, there is general consensus that
the PSF construction has a preference to mark the subject as an active participant (although according to @raineri_analyse_2010 [98] is is not "universally accepted" as some authors also name counterexamples @kupferman_construction_1995 [59].)
In @raineri_analyse_2010 [95]'s view, the idea that the PSF expresses subject responsibility is the "most common hypothesis" ("l'hypothèse la plus répandue") in the literature. -->

In line with their argument that the passive reading of se faire + INF is a kind of causative, @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987 state that the subject referent of the se faire passive is always a causer and therefore always responsible to some degree. This is contrasted to the être passive's subject referent, which is see as more neutral. @gaatone_desagreable_1983 [164] distinguishes between an "active patient" in the se faire construction and a "passive patient" in the être passive (@raineri_analyse_2010 [96]).

In contrast, @kupferman_construction_1995 [59] gives PSF examples where the subject cannot be interpreted as responsible, one of which is repeated here:

```{r gloss}
#| echo: false
gloss_non_resp <- as_gloss(
  "elle sanglotait parce que son tout jeune bébé venait de se faire enlever les amygdales.",
translation = "She sobbed because her little baby got operated on for tonsilitis.",
source = "Kupferman (1995: 59)"
)
gloss_non_resp
```

According to him, subject responsibility emerges from the context and not from the PSF, which is inherently passive [98, 99]. Further support for this idea comes from @labelle_anticausativizing_2013 and @washio_when_1993. For @martin_construction_2002, subject dynamism, but not subject intentionality, is a necessary condition in the PSF (@raineri_analyse_2010 [90]).

Thus, although the requirement of subject responsibility in the PSF is not "universally accepted" (@raineri_analyse_2010 [98]), the analyses show at least some consensus in that the PSF construction prefers to mark the subject as connected to the event (@raineri_analyse_2010), either through affectedness, responsibility or intentionality. This tendency does not exist for the être passive (source).

-   **Animacy**

As implied in the debate on subject responsibility, subjects of se faire passive constructions are predominantly active participants in the described action. Since inanimate subjects are rarely affected and even less often responsible, PSF is common with animate subjects. While some scholars claim that only animate subjects are possible with the PSF, others provide counterexamples (@labelle_anticausativizing_2013, @kokutani_sur_2005, @novakova_construction_2009 \[4\], @sveberg_se_2012) (@raineri_analyse_2010 [92]). We tag animacy as binary "inanimate"/ "animate" although we recognize that animacy has been prominently considered as a continuum in the literature.

However, these examples appear to be exceptions as there is both corpus (@novakova_construction_2009 [111]) and experimental (@koenig_semantic_1992 \[151, 153\]) evidence that subject animacy and PSF are positively correlated: In @novakova_construction_2009 [111]'s corpus study, less than 3% of the se faire sentences had an inanimate subjects. In a survey by @koenig_semantic_1992, French native speakers rated the acceptability of PSF with animate subjects significantly higher than with inanimate subjects. Inanimate subjecta were only allowed by some speakers when the subject "affects a relevant animate entity" (@koenig_semantic_1992 [151, 153]. The preponderance of animate subjects for the se faire passive seems to be one of the few aspects on which scholars agree (@raineri_analyse_2010 [92]).

<!-- The preference of the se faire passive of animate over inanimate subjects is parallel to the English get passive (@hundt_be-_2024), although @raineri_analyse_2010 [92] finds that the get passive allows more inanimate subjects than the se faire passive. According to @bohmann_large-scale_2023 [43], the get passive has become more likely with inanimate subjects over time which may signify a lessening of subject responsibility. @raineri_analyse_2010 brings up the possibility that a similar process of grammaticalisation may be happening in French. (not sure if relevant)-->

-   **Agent PP** (par-phrase, Complément d'agent)
    -   @hundt_be-_2024, the “by-phrase openly assigns responsibility” to an external agent
    -   @bohmann_large-scale_2023: by-phrase “partially mitigate subject responsibility”, expected to favour be-passive (finding: more in the early data than now)
    -   @kupferman_construction_1995: agentivity requirement for agents denoted by par-phrase; not possible for PSF par-phrase with abstract cause (in contrast to être passive?)

**Hypothesis:**

Both *se faire* and *être* passive favour animate subjects over inanimate subjects but subject animacy is a stronger predictor of the PSF than the être passive.

Further, we hypothesize that the presence of an agent PP ("Complément d'agent") prefers the canonical être passive over the se faire passive (because it assigns responsibility to an external agent that is not the subject which is unusual for PSF, where the subject is usually somewhat agentive.

**Interaction**: @raineri_analyse_2010 [94] suspects that the PSF's animacy (almost-) constraint restricts verb selection and provides an explanation for the construction's incompatibility with certain verbs (i.e. not all verbs can occur with animate subjects).

### Aspectuality

@keenan_passive_2007 [340] cite one of Greenberg's universals: "G-5: If a language has two or more basic passives they are likely to differ semantically with respect to the aspect ranges they cover." Assuming that the PSF is a "basic passive", this suggests that the être and se faire passives will demonstrate different aspect preferences. Aspectuality is reflected in the verb, either in semantics ("Aktionsart") or grammatical marking (Tense, Aspect, Mode) [^1] (source missing!).

[^1]: The overall consensus in the literature is that aspect marks progressivity grammatically and Aktionsart semantically/ lexically. However, this clear cut distinction of Aspect and Aktionart has been doubted, and some theories see Aktionsart und Aspect as two instantiations of Aspectuality (Quelle, not focus).

-   **Verb semantics (Aktionsart), Clusters** (@bohmann_large-scale_2023 [51-52]: 21 levels):
    -   se faire passive works well with dynamic verbs but not with stative verbs @raineri_analyse_2010 \[90\], @kupferman_construction_1995 [67] (but exceptions provided by @novakova_construction_2009); Martin 2002, @raineri_analyse_2010 [100-103]; volitional verbs fit well with PSF because they involve an action that leads to a mental state change while non-volitional verbs do not
    -   Telicity: According to @bellec_construction_2014 \[209\], the être passive is likely to describe a result/ mark an endpoint of an action, while the se faire passive is used for actions in progress and verbs that are in their nature ongoing (e.g. perception verbs, @bellec_construction_2014 [209]). This could be reflected in the telicity of the verb, with telic verbs preferring the être passive and atelic verbs favouring the se faire passive.
-   **TAM (Tense, Aspect, Mode)**
    -   **Tense:** Descriptions of future events are more likely to focus on the start and process of an event/ action than on its endpoint (@bellec_construction_2014 [210]) Therefore, it seems likely that future tense will favour the se faire passive over the être passive. Regarding past tense, the être passive will probably be more popular than se faire for constructions in the passé composé (passé composé also used with se faire if speaker wants to emphasize a process, @bellec_construction_2014). In contrast, the se faire passive may be preferred if the action is described in the imparfait (only a guess).
    -   **Aspect** tagging @hundt_be-_2024 : simple, perfect, progressive; @bohmann_large-scale_2023 [39]: "Form of auxiliary verb: present, preterite, perfect, infinitive, -ing"
        -   @raineri_analyse_2010 [365] sees the "type of process denoted by the verb" as a "necessary condition", inchoative/ durative aspect: "get passive and se faire passive focus on onset of the process that affects the subject referent, whereas be and etre passive present the process in its entirety, without offering an internal perspective."
    -   **Mode?** actuality, possibility, necessity) -
        -   Modal verbs (auxiliary): such as "devoir", "savoir", "pouvoir", "falloir", "vouloir" very frequent with se faire passive \@novakova_construction_2009 \[9\] (in \@raineri_analyse_2010 \[97\], reason: they denote "control" to the subject IF POSSIBLE
-   **(Periphrase constructions**: According to @spang-hanssen_quelques_1967 \[141\] and @novakova_construction_2009, periphrases that denote progressivity such as “être en train de + INF”, “venir de + INF” and “aller + INF” occur more naturally with the se faire passive than with the être passive. This is due to the dynamicity that the se faire passive entails (Quelle) compared to the more stative être passive. (also: @bellec_construction_2014; @keenan_passive_2007 [341], @raineri_analyse_2010 [91])

### Adversativity/ Non-neutrality

The factor of "Adversativity/ Non-neutrality" is reflected in the semantics of the verb and captures the (negative) consequences of the described action for the subject. It has been found to be relevant to the choice between the be and get passive (@hundt_be-_2024 [89, 96-97]; @bohmann_large-scale_2023). @spang-hanssen_quelques_1967 and @raineri_analyse_2010 draw parallels between the English get passive and the French PSF construction. @hundt_be-_2024 label it as binary on the verb as adversative, non-adversative/ neutral.

@gaatone_desagreable_1983 argues that the se faire passive has a strong "unpleasant sense" (sens désagreable") because it occurs mainly with verbs denoting unpleasant events [^2].

[^2]: According to @gaatone_desagreable_1983, unpleasant events contradict notion of subject responsibility because subjects would not willingly bring the unpleasant event onto themselves (argument for "primacy of causative reading") According to him, the reason is subject involvement in se faire constructions: if the subject experiences a pleasant event and this is expressed by se faire -\> mostly the subject has initiated the action -\> causative and not passive reading (language example?)/ in contrast, It is less likely that the subject initiates an unpleasant action.

Although @kupferman_construction_1995 \[77\] rejects the "unpleasant sense" as a characteristic of se faire, he agrees that it is more commonly used in unpleasant contexts ("has tended to specialize in unpleasant roles" (compared to être, which is more neutral). This is supported by corpus data: @novakova_construction_2009 [110] found a clear preference for PSF to be realised with unpleasant actions, with 30% of verbs in PSF denoting unpleasant events and only 5% denoting pleasant events.

-   @bohmann_large-scale_2023 [38, 43] Sentiment values of verb participles extracted from SentiWordNet: positivity, negativity, objectivity scores; reversed if sentence is negated

As @bohmann_large-scale_2023 notes, there are two hypotheses related to emotions of the verb: The adversativity and the non-neutrality hypothesis. Both expect the get passive to favour verbs with negative scores; but only the non-neutrality hypothesis predicts get passive to correlate positively with positive emotion scores as well.

**Hypothesis:** Se faire passive be preferred over the être passive for verbs that denote "unpleasant" actions (as reflected by the sentiment analysis).

**Interaction:** Affecteness of the subject may be reflected by an interaction of subject animacy and verb adversativity: Negative consequences of actions, expressed by adversative verbs, affect animate subjects more than inanimate subjects.

### Mode

spoken vs. written

-   se faire passive favoured in spoken French; être passive favoured in written French @raineri_analyse_2010

### (Maybe: Register)

Text genre, formal vs. informal

-   se faire passive more likely in informal settings, être passive more likely in formal settings

### Morphosyntactic features

Negation, subject person, subject number, subject gender

@bohmann_large-scale_2023 [39]: get passive disfavoured in complex constructions (e.g. negation); parallel in French?

## Memory based learning

-   non-parametric: no assumptions about underlying distribution of data
-   no problem with sparse observation of features/ low-frequency events @theijssen2013choosing [243, 244] because it "does not rely on any kind of generative model"
-   only mandatory annotation is class label @theijssen2013choosing [243] (passive choice)
-   "simple-theory neutral features"/ or some linguistic features (animacy, verb, pp-phrase...)
-   compares test instances with labeled training data (stored in the memory)
-   MBL very susceptible to majority classes -\> unbalanced data is a big problem (poor performance for minority classes possible)
-   k-Nearest-Neighbor (kNN):
    -   kNN algorithm is trained on data to predict the most likely passive choice
    -   nearest neighbours out of all instances are determined -\> observation receives class that majority of nearest neighbours has
    -   leave-one-out-setting @theijssen2013choosing [244]: train on all but one, test on remaining instance, iterate for each instance
    -   hyperparameters setting (e.g. number of neighbors)? -\> help of wrapper
    -   Tutorial: <https://www.datacamp.com/tutorial/k-nearest-neighbors-knn-classification-with-r-tutorial>

1.  choose the value of K, which is the number of nearest neighbors that will be used to make the prediction.
2.  Calculate the distance between that point and all the points in the training set.
3.  Select the K nearest neighbors based on the distances calculated.
4.  Assign the label of the majority class to the new data point.
5.  Repeat steps 2 to 4 for all the data points in the test set.
6.  Evaluate the accuracy of the algorithm.

-   in R: "class" or "caret" package
-   set seed for reproducibility
-   split set manually or by "caTools" library
-   Preprocessing/ Feature scaling if feautres have different scales, to ensure proportions between features and prevent feautre with largest scale to dominate distance calculations
-   caret: Model Tuning: find optimal value of "K"
-   caret: train best performing model
-   Model evaluation

# Results

# Discussion

# Conclusion/Outlook

# References
