---
title: "The French Passive Alternation: A Corpus-Based Study"
author: "Rose Hörsting & Gina Reinhard"
date: "`r format(Sys.Date())`"
engine: knitr
bibliography: references.bib
format: 
  html:
    df-print: default
    toc: true
    toc-depth: 3
    fig-width: 8
    fig-height: 6
    number-sections: true
    embed-resources: true
execute:
  warning: false
link-citations: true
csl: "apa_7th.csl"
number-sections: true
---

```{=html}
<style>

p {
  text-align: justify;
}

</style>
```
```{r load-libraries, include=FALSE}

library(here)
#install.packages("glossr")
library(glossr)
library(kableExtra)
library(tidyverse)
#install.packages("psy")
library(psy)
```

```{r packages-regression, include=FALSE}
library(mclm)
library(kableExtra)
library(lmerTest)
library(afex)
library(ggeffects)
library(sjPlot)
library(rms)
library(visreg)
library(car)
library(arm)
library(DHARMa)
```

```{r styling, include=FALSE}

use_glossr(styling = list(
  source = "b",
  first = "",
  font_family = "Cambria"
))

```

```{r import-data, include=FALSE}

sefaire_sample <- read.csv(file = here("data", "sefaire_sample.csv"))
etre_sample_1 <- read.csv(file = here("data", "etre_sample_1.csv"))
etre_sample_2 <- read.csv(file = here("data", "etre_sample_2.csv"))
etre_sample_3 <- read.csv(file = here("data", "etre_sample_3.csv"))
etre_sample_4 <- read.csv(file = here("data", "etre_sample_4.csv"))
```

```{r etre-sample, include=FALSE}

etre_sample <- rbind(etre_sample_1, etre_sample_2, etre_sample_3, etre_sample_4)

```

```{r randomisation, include=FALSE}

set.seed(2142637384) #lottery numbers too long, deleted last digit

etre_sample_randomised <- etre_sample |>
  slice_sample(n = 700)

write.csv(etre_sample_randomised,"./data/etre_sample_randomised.csv")

```

```{r downsampling, include=FALSE}
etre_sample_clean <- read.csv(file = here("data","etre_sample_clean.csv"))
set.seed(2142637384)

etre_sample_532 <- etre_sample_clean |> 
  slice_sample(n= 532)
```

```{r downsampling-false-positive, include=FALSE}
#one false positive was found after sampling (ID=332) but sampling completely new not possible
set.seed(2142637384)

etre_sample_533 <- etre_sample_clean |> 
  slice_sample(n= 533)

etre_sample_1 <- etre_sample_533 |> 
  filter(!etre_sample_533$ID %in% etre_sample_532$ID)
```

```{r downsampling-correction, include=FALSE}
# ID217 is "next in line" given the seed
# correction: remove 332 and add 217 to sample
etre_sample_532 <- etre_sample_532 |> 
  filter(ID != "332") |> 
  bind_rows(etre_sample_1)

write.csv(etre_sample_532,"./data/etre_sample_532.csv")

```

```{r etre-difference, include=FALSE}
difference_list <- etre_sample_clean |> 
  filter(!etre_sample_clean$ID %in% etre_sample_532$ID)

difference_ID <- difference_list["ID"]

```

```{r cql-lemmas, include=FALSE}
unique_lemma <- unique(sefaire_sample$lemma)
lemma_list <- sprintf('lemma="%s"', unique_lemma)
cat(paste(lemma_list, collapse = " | "))

```

# Introduction {#sec-introduction}

The most common way of expressing passive in French is via [être]{.smallcaps}, as in *Les députés sont élus* ('The members of parliament are elected'), as attested in the ParlaMint corpus [@11356/1431]. However, alternatively, one may use [se faire]{.smallcaps} to express passivity: *Les députés se font élire* ('The members of parliament got elected'). We use [être]{.smallcaps} and [se faire]{.smallcaps} in small capitals to refer specifically to the passive variants, distinguishing these forms from other uses of the verbs *être* and *se faire*. Previous research suggests that the choice between these two variants is not arbitrary: [se faire]{.smallcaps} may be chosen to convey a particular meaning or emphasise something that [être]{.smallcaps} cannot capture [@kupferman_construction_1995: 57-62]. But what exactly determines whether the alternative passive is chosen, or even preferred, over the canonical one?

The [être:se faire]{.smallcaps} alternation, and [se faire]{.smallcaps} in particular, has been widely studied and discussed, for example by @bellec_construction_2014; @gaatone_desagreable_1983; @koenig_semantic_1992; @kupferman_construction_1995; @labelle_anticausativizing_2013; @novakova_construction_2009; @raineri_analyse_2010, and @tasmowski-de_ryck_causatif_1987. The results of these analyses overlap to varying degrees, but are not entirely consistent. This indicates that the alternation is quite nuanced and deserves further investigation.

Quantitative and qualitative analysis agree that [se faire]{.smallcaps} strongly favours animate subjects [@novakova_construction_2009: 111; @koenig_semantic_1992: 151-153]. This, in turn, is often taken to indicate active involvement of the subject, despite the fact that, given the nature of the passive, the referent of the subject undergoes the embedded action. Some go as far as to state that the subject of an [se faire]{.smallcaps} construction is always responsible for the action happening to them [@tasmowski-de_ryck_causatif_1987], while others emphasise subject affectedness, proposing that as a patient, the subject is not necessarily responsible but rather impacted [@kupferman_construction_1995: 59-60; @bellec_construction_2014: 212-213]. This is related to the debate on passivity and causativity: Since the subject of [se faire]{.smallcaps} constructions is often both the initiator and the undergoer, it has been disputed whether the construction should be considered passive at all [@bellec_construction_2014: 204; @raineri_analyse_2010: 88].

Another factor that has been proposed to influence passive choice is adversativity. @gaatone_desagreable_1983 claims that [se faire]{.smallcaps} constructions often carry a negative meaning (compared to [être]{.smallcaps} which is considered semantically more neutral). Speakers may prefer [se faire]{.smallcaps} to [être]{.smallcaps} to express that the subject is negatively affected by the action described.

Thirdly, @raineri_analyse_2010 and @bellec_construction_2014 propose that [se faire]{.smallcaps} and [être]{.smallcaps} behave differently in terms of their aspectual preferences: [se faire]{.smallcaps} is favoured by dynamic verbs denoting ongoing actions with progressive aspect, whereas [être]{.smallcaps} tends to cover stative verbs and completed actions [@raineri_analyse_2010: 90; @bellec_construction_2014: 209]. Research on the English [be:get]{.smallcaps} alternation [@bohmann_large-scale_2023], which has been considered parallel to [être:se faire]{.smallcaps} [@raineri_analyse_2010: 87], allows the hypothesis that verbs describing violent actions may be realised more often with [se faire]{.smallcaps} than with [être]{.smallcaps}, suggesting a potential interaction between aspectuality and adversativity.

These factors and the subtle differences between the passive variants have been studied extensively. However, to our knowledge, no quantitative study has investigated the exact factors that predict [se faire]{.smallcaps} over [être]{.smallcaps} using logistic regression modelling. Our study aims to fill this research gap: By applying a quantitative, multivariate, probabilistic approach to model corpus data, we investigate the factors that influence the choice between [se faire]{.smallcaps} and [être]{.smallcaps} in the passive alternation in spoken French, more precisely in the context of parliamentary debates. Specifically, we aim to understand which factors contribute to speakers’ preference for one construction over the other. Although their generalisability may be limited, parliamentary debates offer a wide linguistic and thematic variety and are a rich source of contemporary spoken French.

According to Gries [-@gries_statistics_2021: 238], regression modelling is suitable for linguistic analyses because it can capture the complex interplay of various factors. Furthermore, binomial logistic regression modelling is particularly well-suited for the study of alternations with two possible outcomes [@gries_statistics_2021: 238-239; @levshina_how_2015: 253-254]. It has been successfully employed for investigation of passive variations in other languages, such as English [@bohmann_large-scale_2023; @hundt_be-_2024], Swedish [@cuypere_corpus-based_2014], and Chinese [@guan_multifactorial_2024]. Therefore, such a model promises insightful results that may contribute to the discussion of the passive alternation and potentially provide new insights into the use of both [se faire]{.smallcaps} and [être]{.smallcaps}.

We begin by describing the [être: se faire]{.smallcaps} alternation in detail in @sec-alternation. In the following subsections (@sec-etre, @sec-se-faire), we provide detailed information on the two participants of the alternation in order to establish the background and motivation for this study. @sec-factors concludes the theoretical background by introducing the variable context and outlining the three main factors that we hypothesise, based on the literature, to fundamentally determine the choice between [se faire]{.smallcaps} and [être]{.smallcaps}. @sec-data-and-methods details the methodology of this study: we explain the process of data collection (@sec-data-collection, including manual filtering, annotation and preprocessing steps) and the process of modelling binomial logistic regression (@sec-log-regression, including model selection and model diagnostics). We then report the results of our study in @sec-results, interpreting them within the key concepts previously introduced: subject responsibility (@sec-results-subj-resp), adversativity (@sec-results-adversativity), and aspectuality (@sec-results-aspectuality). This is followed by the discussion (@sec-discussion), in which our findings are embedded with the previous literature and our results are evaluated in the light of the larger research picture. In the same vein, the limitations of our study are noted and discussed. The paper concludes with @sec-conclusion, which concisely summarises our main findings and how they relate to passive alternation research, and provides directions for further corpus-based investigation.

<!-- ------------------------------------------------------------------------ -->

<!-- The French construction *se faire* followed by an infinitive has evolved through analogy with the passive voice. Originally, it had a causative meaning, where the subject played an active role in the action. However, in some contexts, it can also be interpreted as passive, alternating with the *être* passive construction. -->

<!-- This study investigates the factors that influence the choice between *se faire* and *être* in the passive alternation in spoken French, with a particular focus on parliamentary debates. Specifically, we aim to understand what factors contribute to speakers’ preference for one construction over the other. -->

<!-- To address this question, we use a quantitative, multivariate, probabilistic approach to model corpus data. Our results show a main effect of semantic verb clusters, with communication verbs favouring *se faire* over *être* in passive constructions. -->

<!-- and the logistic regression model revealed a significant interaction between subject animacy and adversativity -->

# The Passive Alternation of [être]{.smallcaps} and [se faire]{.smallcaps} {#sec-alternation}

This section introduces the French passive variants *être* and *se faire*, hereinafter [être]{.smallcaps} and [se faire]{.smallcaps}, focusing on their structure and meaning in the context of passive alternation. While [être]{.smallcaps} is the unmarked passive construction in French, [se faire]{.smallcaps} can also function as a passive variant in certain contexts. We outline the factors used as predictors in our mixed-effects logistic regression model of the [être:se faire]{.smallcaps} alternation.

Unless otherwise specified, all examples we provide in this study are from our sample from the French ParlaMint 2.1 corpus [@11356/1431].

## Canonical Passive: [être]{.smallcaps} {#sec-etre}

[être]{.smallcaps} is the unmarked passive variant in French. It involves an alternation of verb valency that reverses the syntactic roles of the arguments within a clause. In the active construction, the first argument (the actor) is realised as the subject and the second argument (the undergoer) as the direct object. In the passive, this order is reversed: the second argument is promoted to subject position, while the first argument either becomes a prepositional phrase or is omitted. This active/passive alternation, traditionally referred to as voice or diathesis, shifts the focus of the discourse. While the verb retains its core meaning, the attention is shifted away from the actor, emphasising the action itself or its result rather than the referent performing it [@abeille2021grande: 209-210].

[être]{.smallcaps} consists of a form of the attributive verb *être* 'to be' and the *participe passif* (passive participle), which is identical in form to the *participe passé* (past participle), e.g. *élus* 'elected'. Although other attributive verbs, such as *rester* 'to stay' and *sembler* 'to seem', can also be used in passive constructions, they are less common. This term paper will focus specifically on the passive formed with *être*. In the [être]{.smallcaps} construction, the subject of the active sentence becomes an adjunct phrase, which is called *complément d'agent*, introduced by the preposition *par* or *de*. When the *complément d'agent* is present, the construction is referred to as *passif long*; when the agent is omitted, it is called *passif court* [@abeille2021grande: 210-211].

For example, in the following sentences, in 1.a the agent *nos concitoyens* appears in the *complément d'agent* introduced by *par*, while in 1.b the agent is omitted:

> (1) 
>
>     (a) *\[...\] ils sont élus par nos concitoyens au suffrage universel.*\
>         '\[...\] they are elected by our fellow citizens through universal suffrage.'
>     (b) *Les députés sont élus pour contrôler votre action.*\
>         'The members of parliament are elected to monitor your actions.'

The *passif long* can favour the interpretation of a habitual process, as in *La bière est servie par le patron* 'The beer is served by the owner'. In contrast, the *passif court* focuses on the action or its result, without specifying who performed it [@abeille2021grande: 219].

An impersonal passive (*passif impersonnel*) is also possible in French. In impersonal passives, the subject position is occupied by the expletive *il* 'it'. In this construction, the direct object can remain in the complement position [@abeille2021grande: 210-211]. An example of an impersonal passive from our sample is:

> 2.  *Il est prévu un suivi des progrès accomplis d'ici à fin juillet* *2018 \[...\]*\
>     'A follow-up of the progress made is planned by the end of July 2018 \[...\]'

Some grammars distinguish between a *passif verbal* (verbal passive), with an event interpretation, and *passif adjectival* (adjectival passive), with a resultative interpretation. According to *La grande grammaire du français* [@abeille2021grande], this distinction reflects semantic differences concerning the aspect of the event. With stative verbs, active and passive sentences are generally semantically equivalent, as seen in the examples *Tout le monde aime Marie* 'Everyone loves Marie' vs. *Marie est aimée de tout le monde* 'Marie is loved by everyone'. However, with event verbs, which involve a change in state or location, the passive form often emphasises the result or outcome of the change. For example, in the *passé composé* (compound past tense), *On a servi la bière* 'They served the beer' describes the event, while *La bière a été servie* 'The beer has been served' emphasises the result of the action [@abeille2021grande: 219].

Finally, distinguishing between passive constructions and adjectives in French is not always straightforward. One difference between them is that the *participe passif* can form impersonal constructions with a nominal complement, as seen in example 2. Additionally, adjectives often have antonyms with a negative prefix, such as *connu* 'known' and *inconnu* "unknown". In some cases, the adjective may have a meaning or construction that differs significantly from the passive participle, as in the contrast between *Ce restaurant est très couru* 'This restaurant is very popular' and *On court beaucoup ce restaurant* [@abeille2021grande: 212]

To summarise, [être]{.smallcaps} is the canonical form of passivisation in French, providing a means for speakers to shift the focus away from the actor and onto the action and its result. The next section examines the passive variant [se faire]{.smallcaps}, which, while sharing some functional similarities with [être]{.smallcaps}, also presents distinct structural and semantic differences.

## Alternative Passive: [se faire]{.smallcaps} {#sec-se-faire}

### Structure and Meaning {#sec-sefaire-structure-meaning}

When constructing a passive sentence, French speakers may choose to use [se faire]{.smallcaps} instead of the canonical [être]{.smallcaps}, as [se faire]{.smallcaps} provides an alternative way to express passive meaning in French. As example 3 shows, [se faire]{.smallcaps} consists of the pronoun *se* (which shows agreement with the subject), a form of the verb *faire* ('to do'), and a verb in the infinitive.

> 3.  *Des policiers se font agresser.*\
>     'Police officers are attacked.'

If used with an intransitive verb, as in example 4, [se faire]{.smallcaps} does not have a passive sense. Its semantics are then composed of the literal meaning of its components.

> 4.  *Elle se faire rire.* [@kupferman_construction_1995: 61]\
>     'She made herself laugh.'

[se faire]{.smallcaps} resembles [être]{.smallcaps} in that the direct object of the active sentence becomes the subject of the passive sentence. This subject is the experiencer of the action described. Different than for [être]{.smallcaps}, however, the diathesis is not restricted to the direct object: [se faire]{.smallcaps} can also occur with ditransitive verbs, where the indirect object becomes the subject [@raineri_analyse_2010: 93-95, @koenig_semantic_1992: 144], as demonstrated by example 5.

> 5.  *Les intéressés peuvent se faire expliquer tout ce qui touche aux retraites.*\
>     'Interested parties can have everything related to pensions explained to them.'

[se faire]{.smallcaps} can also be used to express active involvement of the subject while still conveying a sense of subject undergoing.

> 6.  *Oh... tu t'es fait couper les cheveux pour l' été!* [@raineri_analyse_2010: 156]\
>     'Oh.. you've had your hair cut for summer!'

However, constructions like example 6 are controversial: Sveberg [-@sveberg_se_2012: 164-165] argues that such instances of [se faire]{.smallcaps} are fundamentally causative rather than passive, since the subject both initiates and undergoes the action (as in: 'You've made the hairdresser cut your hair').

<!-- Labelle [-@labelle_anticausativizing_2013: 235] classifies [se faire]{.smallcaps} as “formally causative, but semantically passive”. -->

The semantic tension between the fundamentally causative nature of the verb *faire* and the passive reading has led scholars to question the nature of [se faire]{.smallcaps}, as Koenig and Pederson [-@koenig_semantic_1992: 145] put it: "how can a passive-like construction arise from the combination of a causative and a reflexive marker, where no marking of the passive function is apparent?". Various answers to this question have been proposed: @kupferman_construction_1995 sees the causative and passive readings as homonyms with fundamentally different semantics. According to him, [se faire]{.smallcaps} is always causative when used with an intransitive verb (ex. 5.), but always passive when used with "non-agentive" predicates [@kupferman_construction_1995: 61]. This is in line with Bellec [-@bellec_construction_2014; -@bellec_quand_2014], who takes a diachronic approach: she describes a process of grammaticalisation and reanalysis for [se faire]{.smallcaps} through which *se* has lost its reflexive meaning and *faire* has become an auxiliary. According to her, the causative meaning has not disappeared but now coexists with the passive reading. Notably, Raineri [-@raineri_analyse_2010: 280-305] suggests, based on her corpus study, that [se faire]{.smallcaps} did not derive from the causative reading but evolved independently.

<!-- Labelle [-@labelle_anticausativizing_2013: 23] adopts a syntactic approach, proposing that the passive meaning is derived through a mechanism that links the empty NP-object position of the embedded verb to the subject via *se*. This binding causes the subject to be interpreted as affected, as it is coindexed with the missing object. This object would be affected in the active sentence, but is absent on the surface in [se faire]{.smallcaps}. This interpretation works only when the embedded verb has an agentive subject, as the missing object must be affected. The subject therefore remains the agent of the action, but is at the same time affected by the action of the embedded verb, which creates the passive interpretation. -->

These accounts offer only a brief glimpse into the complex analysis of [se faire]{.smallcaps}.The questions how the passive reading emerges, and how it differs from the syntactically identical causative reading remain controversial, and will be explored in the following @sec-debate.

### Passive or Causative: the Debate {#sec-debate}

The above-described stance taken by @bellec_construction_2014, @labelle_anticausativizing_2013, @koenig_semantic_1992 and @kupferman_construction_1995 that [se faire]{.smallcaps} is a full-fledged passive construction and clearly separate from the causative is by no means undisputed. The opposing view - that the construction is a type of causative - is also widely represented in the literature. It is most prominently presented by @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987 and has more recently been reworked by @sveberg_se_2012. In their view, the passive sense emerges only from the context, rather than from the construction itself. They centre their argumentation around the notion that the subject of *se faire* is inherently responsible for the described event [@sveberg_se_2012: 160]. <!-- "always, even unconsciously, responsible for the occurrence of the embedded event." (Labelle) -->

However, Koenig [-@koenig_semantic_1992: 143] points out that sentences including *se faire* can have both a passive and a causative sense. Example 7 illustrates this twofold interpretation:

> 7.  *il y a \[…\] des parlementaires qui se sont fait élire sur la promesse de ne pas supprimer 120 000 postes.*\
>     'there are \[…\] Members of Parliament who were elected on the promise not to cut 120,000 jobs.'\
>     'there are \[…\] Members of Parliament who made themselves be elected on the promise not to cut 120,000 jobs.'

As Washio [-@washio_when_1993] describes, the linguistic proximity of causativity and passivity is not exclusive for French but quite typical across the world's languages, as "passive and causative senses are shown to share a basic conceptual structure" [@washio_when_1993: 45]. Crucially, passive and causative *se faire* are structurally identical, differing only in their semantics. According to Labelle [-@labelle_anticausativizing_2013: 236] the ambiguity is resolved by the semantic nature of the three main components: the subject referent, *se*, and *faire*. Referring to these three components, Bellec [-@bellec_construction_2014: 207] defines three criteria to distinguish the causative from the passive interpretation: She claims that in the causative, the subject referent is agentive and has the role of both causer and patient, whereas in the passive, the referent is affected by the event and therefore only has the role of the patient. Secondly, according to her, the particle *se* has a reflexive meaning in the causative but is non-reflexive in the passive. Similarly, *faire* is said to have a transparent causative meaning only in the causative and not in the passive [@bellec_construction_2014: 207; see also @koenig_semantic_1992: 143].

Sometimes both readings are possible and the ambiguity cannot be resolved [@koenig_semantic_1992: 144], especially with cognition and perception verbs [@bellec_construction_2014: 206-207]. Therefore, despite its productivity as a passive construction, Raineri [-@raineri_analyse_2010: 356-357] concludes that *se faire* is problematic because it does not fit into two distinct categories nor does it function as a single weak causative. She points to the possibility of multiple readings [@kokutani_sur_2005] or of a continuum between causative and passive [-@raineri_analyse_2010: 89, 154-166]. The idea of the continuum implies that *se faire* constructions shift between passive and causative interpretations depending on the surrounding context. Their inherent flexibility suggests that a strict binary classification is inadequate, reinforcing the idea of a passive–causative continuum. This continuum can be understood in terms of the degree of agentivity attributed to the subject referent. When its role is reduced to mere initiation of the event, the construction can be fully paraphrased with [être]{.smallcaps} without altering the overall meaning of the sentence. For example, *Il se fait remarquer par une patrouille de police* 'He is noticed by a police patrol' and *Il est remarqué par une patrouille de police* 'He is noticed by a police patrol' illustrate a clear equivalence between [se faire]{.smallcaps} and [être]{.smallcaps}. Moreover, other contextual elements can weaken the construction's causativity and strengthen the passive interpretation, reflecting varying agentivity depending on the specific discourse context [@raineri_analyse_2010: 154-160].

This debate has been central to the investigation of *se faire* and cannot be left aside. However, even though its passive-status has been the subject of disagreement and discussion, the literature clearly shows that there is a passive reading associated with this construction. This is independent of whether the passive sense emerges from within the construction or from the components it occurs with, or whether it has an underlying causative meaning. By comparing [se faire]{.smallcaps} and [être]{.smallcaps} constructions, and examining the factors contributing to the choice of one or the other, we aim to shed light on this debate from a corpuslinguistics perspective.

## [être]{.smallcaps} and [se faire]{.smallcaps}: Competition or Complementation? {#sec-factors}

### Defining the Variable Context {#sec-sefaire-context}

Both [se faire]{.smallcaps} and [être]{.smallcaps} passives are available for the active diathesis [@novakova_construction_2009: 108] and assign the role of the undergoer to the subject referent (see @sec-etre, @sec-se-faire). Therefore, the two variants form an alternation and are expected to compete with each other in contexts where passive is expressed. We find that both variants may be used in the same context, as example 8 illustrates:

> 8.  *Je remarque d' ailleurs que ce sont souvent ceux qui filment qui **se font éborgner** ou **sont blessés** par des tirs de LBD*.\
>     'Incidentally, I've noticed that it's often those filming who **get their eyes poked out** or **are injured** by LBD fire.'

However, some scholars speak of semantic differences between canonical [être]{.smallcaps} and [se faire]{.smallcaps}: Novakova [-@novakova_construction_2009: 108] notes that the constructions differ in semantic nuances, as they have *un sens proche mais pas identique* ('a similar but not identical sense'). These differences may be more or less pronounced depending on the context: according to Raineri [-@raineri_analyse_2010: 87], [se faire]{.smallcaps} is marked and predominantly used in informal contexts. Additionally, [se faire]{.smallcaps} has a strongly figurative sense that [être]{.smallcaps} is missing: it is often used as part of idiomatic or lexicalised expressions [@raineri_analyse_2010: 165; @novakova_construction_2009: 117], and is associated with a *besoin d’expressivité and extravagance* ('a need of expressivity and extravagance') [@bellec_construction_2014: 206]. As discussed in @sec-debate, previous research also suggests that [se faire]{.smallcaps} is more likely to have a causative undertone while [être]{.smallcaps} is more neutral, with varying degrees of passivity for [se faire]{.smallcaps} [@raineri_analyse_2010: 154-166].

These subtle differences in meaning pose certain challenges for the contrastive analysis of the alternation. Bohmann et al. [-@bohmann_large-scale_2023: 35-36] face similar obstacles when defining the variable context of the English passive alternation of [be]{.smallcaps} and [get]{.smallcaps}. Since our aim is to investigate constructions that participate in the passive alternation of [être]{.smallcaps} and [se faire]{.smallcaps}, we follow Raineri [-@raineri_analyse_2010: 165] in including only [se faire]{.smallcaps} constructions that have a natural paraphrase with [être]{.smallcaps}. In line with Bohmann et al. [-@bohmann_large-scale_2023: 35], who argue that "all extracted forms should be genuine instances of the variable under investigation" and that "only cases where both variants can be used interchangeably should be considered", we intend to exclude non-contrastive constructions, such as idioms and ditransitive verbs. Our criteria for excluding constructions are explained in detail in @sec-filtering-criteria.

We recognize this "fuzzy" [@bohmann_large-scale_2023: 50] variable context as both a challenge and an opportunity. It is precisely these subtle differences that make the analysis interesting and our research valuable. Consequently, our study seeks to identify the exact factors influencing the choice between the two variants by statistically modeling real language data. Our central research question is: what factors determine the [se faire: être]{.smallcaps} passive alternation in spoken French, specifically in the context of parliamentary debates? In other words: what factors influence the speaker's choice between these two constructions? In the following sections, we outline the factors that we expect to shape this choice and present our hypotheses on passive preferences.

```{=html}
<!-- Several corpus studies on voice choice in French have shown that the choice between active and passive constructions (*être* passive) is not purely syntactic, but that semantic and pragmatic factors also play a crucial role in determining whether a passive construction is favoured.

-   Subjects in passive sentences are preferably animate, definite, and pronominal. In the study by @da_cunha_lalternance_2020, animate, definite, and pronominal arguments that function as the objects in active sentences were more frequently found in the passive (and vice versa). -\> prominence / informational structure as significant factor. (And: constituent length, shorter constituents are preferred in subject position and favour the passive voice) -->
```
### Subject Responsibility {#sec-subj-resp}

As indicated in @sec-debate, the question whether the referent of the subject of [se faire]{.smallcaps} holds any responsibility is central to the debate of the construction's passive-status [@sveberg_se_2012; @bellec_construction_2014: 212-213]. In line with their argument that [se faire]{.smallcaps} is a kind of causative, @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987 state that the subject referent of [se faire]{.smallcaps} is always a causer and therefore always responsible to some degree. They contrast it to [être]{.smallcaps}'s subject referent, which is seen as more neutral [@gaatone_desagreable_1983: 164-166].

In contrast, Kupferman [-@kupferman_construction_1995: 59] gives [se faire]{.smallcaps} examples where the subject cannot possibly be interpreted as responsible, one of which is repeated in here:

> 9.  *Elle sanglotait parce que son tout jeune bébé venait de se faire enlever les amygdales.* [@kupferman_construction_1995: 59]\
>     'She sobbed because her little baby got operated on for tonsilitis.'

In example 9, *son tout jeune bébé* 'her little baby' does not have any responsibility for the event happening to them. The sentence can only be interpreted as passive. According to Kupferman [@kupferman_construction_1995: 98-99], subject responsibility emerges from the context and not from the inherently passive [se faire]{.smallcaps}. Further support for this idea comes from @labelle_anticausativizing_2013 and @washio_when_1993. For @martin_construction_2002, subject dynamism, but not subject intentionality, is a necessary condition [@raineri_analyse_2010: 90][^1].

[^1]: Unfortunately, we could not get access to the original paper of @martin_construction_2002 and therefore need to rely on @raineri_analyse_2010's report of @martin_construction_2002's analysis.

Thus, although the requirement of subject responsibility in [se faire]{.smallcaps} is not universally accepted [@raineri_analyse_2010: 98], the analyses show some consensus in that [se faire]{.smallcaps} prefers to mark the subject as connected to the event, either through affectedness, responsibility or intentionality. We expect that this tendency is less pronounced for [être]{.smallcaps}. In our study, we model subject responsibility by the factors subject animacy, subject gender, *complément d'agent*, and modal verbs. Our overarching hypothesis is that [se faire]{.smallcaps} will be favoured over [être]{.smallcaps} in constructions with responsible subjects, indicated by animate, masculine subjects, the absence of *complément d'agents* and the presence of modal verbs.

**Subject Animacy**

Corpus studies have shown that animacy significantly influences voice choice in French. A human undergoer (i.e. the direct object in the active voice and the subject in the passive) increases the likelihood of using [être]{.smallcaps} over the active voice in written as well as in spoken French [@da_cunha_lalternance_2020; @da2022alternance]. According to @da_cunha_lalternance_2020 and @da2022alternance, this is due to animate referents being more salient than inanimate ones, which makes them more likely to be promoted to subject position. In parallel, we also expect the animacy factor to play an important role in the distribution of [se faire]{.smallcaps} in relation to subject responsibility.

<!-- As discussed, subjects of [se faire]{.smallcaps} are typically active participants in the described action. Since inanimate subjects are rarely affected, and even less often responsible, [se faire]{.smallcaps} is predominantly used with animate subjects.  -->

While Gaatone [-@gaatone_desagreable_1983: 166-167] argues that [se faire]{.smallcaps} requires animate subjects, Novakova [-@novakova_construction_2009: 111] provides counterexamples. However, these examples appear to be exceptions: both corpus data [@novakova_construction_2009: 111] and experimental evidence [@koenig_semantic_1992: 151-153] point to a strong correlation between subject animacy and the use of [se faire]{.smallcaps}. In Novakova [-@novakova_construction_2009: 111]'s corpus study, fewer than 3% of [se faire]{.smallcaps} constructions have inanimate subjects. This is also supported by a survey by @koenig_semantic_1992, where French native speakers rated [se faire]{.smallcaps} with animate subjects as significantly more acceptable than with inanimate subjects. As Raineri [-@raineri_analyse_2010: 92] puts it, the preference for animate subjects with [se faire]{.smallcaps} is one of the few aspects on which scholars agree.

The preference for animate subjects with [se faire]{.smallcaps} finds a parallel in the English [get]{.smallcaps} which also tends to occur with animate subjects [@hundt_be-_2024: 102]. There is evidence that [get]{.smallcaps} has become more common with inanimate subjects over time [@bohmann_large-scale_2023: 43], which may indicate a shift away from subject responsibility. Raineri [-@raineri_analyse_2010: 92] suggests that a similar change may be occurring in French, with [se faire]{.smallcaps} potentially becoming less constrained in terms of subject animacy. However, Raineri [-@raineri_analyse_2010: 92] notes that [get]{.smallcaps} allows for inanimate subjects more frequently than [se faire]{.smallcaps}.

Animacy is considered a hierarchical scale in the literature, most prominently by @zaenen_animacy_2004. However, for the scope of our study, we define animacy as a binary factor (inanimate/animate). Following @zaenen_animacy_2004, we classify metonymic references to organisations, understood as "collectivities of humans", as animate. In line with the literature, we hypothesise that animate subjects favour [se faire]{.smallcaps} over [être]{.smallcaps}. However, we expect that both [se faire]{.smallcaps} and [être]{.smallcaps} are more frequent with animate subjects than with inanimate subjects. Additionally, we propose that passive choice may not be determined by animacy alone, but rather by its interaction with the semantic verb cluster. We describe this interaction hypothesis in the discussion of verb clusters in @sec-aspectuality.

**Gender**

Since previous research suggests that grammatical gender plays a role in voice choice, it is worth investigating whether this effect extends to the alternation between [être]{.smallcaps} and [se faire]{.smallcaps}.

@da_cunha_lalternance_2020 found that in written French, a feminine agent argument increased the likelihood of [être]{.smallcaps} over the active voice in their *passif court* model - i.e. where the agent is omitted - with no interaction with animacy (human/non-human) or definiteness. This suggests that grammatical gender itself may directly influence voice choice. However, given the study’s small sample size, it is also possible that an interaction effect exists but was not statistically significant. @da2022objectifying observed a similar syntactic bias in both French and English, where female referents were less likely to appear as subjects than male referents. In French, they found an interaction between the grammatical gender and animacy, as the bias was only evident for human referents.

Since we do not infer the actor from context in this study, our analysis focuses solely on the gender of the subject in the passive construction. If [se faire]{.smallcaps} favours subjects with high salience, responsibility, or affectedness, the grammatical gender of the subject may influence the alternation between [se faire]{.smallcaps} and [être]{.smallcaps}. Specifically, if feminine referents are perceived as less active or agentive in certain contexts, a feminine subject may favour [être]{.smallcaps} over [se faire]{.smallcaps}.

To test this, we categorise gender into four levels: unknown, mixed, masculine, and feminine. Our hypothesis is that a feminine subject will favour [être]{.smallcaps} and disfavour [se faire]{.smallcaps}. We will test the potential interaction between gender and animacy, as the effect of grammatical gender may depend on whether the subject is animate or inanimate, potentially linking to the social gender of animate subject referents. Additionally, we will examine how unknown or mixed-gender subjects behave in terms of this alternation.

**Complément d'agent**

The *complément d'agent* introduces an external argument into the passive construction, typically referring to the causer or instigator of the event. It is derived through diathesis from the subject of the corresponding active sentence (@sec-etre).

The *complément d'agent* is optional, usually formed with *par*, sometimes with *de*, and very rarely with *dans*. We do not consider the special case with *dans* in our analysis. In most cases, *de* alternates with *par*, but the two are not always semantically interchangeable. In contrast to *par*, *de* is generally associated with non-dynamic, stative, or non-agentive interpretations. For instance, the verb *accompagner* 'to accompany' implies a less active role of the argument with *de* than with *par*. *de* is frequent with stative verbs across various semantic classes and some verbs, such as *constituer* 'to constitute', naturally prefer *de* over *par*. In addition, when the subject of the active construction is an infinitive or a subordinate clause, *de* is used in the passive voice. With a nominal subject, however, the passive with *de* is a remnant of Classical French and is therefore very rare, accounting for only 2% of cases [@abeille2021grande: 216-218].

We acknowledge that introducing the *complément d'agent* as a predictor in our analysis adds an additional source of variance. Nevertheless, we have chosen to include it, as we believe it plays a crucial role in subject responsibility. Corpus studies on English passive constructions by @hundt_be-_2024 and @bohmann_large-scale_2023 suggest that the agentive prepositional phrase (*by*-phrase) reduces subject responsibility by introducing another entity that carries the agentive role. We believe the same may hold for French, where the *complément d'agent* could similarly influence the distribution of passive constructions.

We hypothesise that the *complément d'agent* influences the choice between [être]{.smallcaps} and [se faire]{.smallcaps}. The presence of a *complément d'agent* should favour [être]{.smallcaps}, as it introduces a responsible external argument and shifts responsibility away from the subject. Conversely, the absence of a *complément d'agent* is expected to favour [se faire]{.smallcaps}, as [se faire]{.smallcaps} implies some degree of subject responsibility. We expect phrases with *par* to have a stronger effect than phrases with *de*, as they typically refer to more agentive referents.

**Modal Verb**

Novakova [-@novakova_construction_2009: 116-117] notes that modal verbs frequently occur with [se faire]{.smallcaps}. According to Raineri [-@raineri_analyse_2010: 97, 162] one possible reason is that modals contribute to subject involvement by expressing agentivity. Raineri [-@raineri_analyse_2010: 97] presents @martin_construction_2002's observation that volitional verbs fit well with [se faire]{.smallcaps} because they involve an intentional action that results in a change of mental state. Since modals usually convey some kind of subject intention or involvement, we expect [se faire]{.smallcaps} to appear more often with modal verbs than [être]{.smallcaps}.

We assign the label modal if a form of one of the modals *devoir*, *vouloir* or *pouvoir* has scope on [se faire]{.smallcaps}. We limit the modal category to these three modal verbs to ensure consitent annotation; verbs that could have a modal sense in some contexts (e.g. *savoir*) and paraphrasal constructions indicating subject responsibility (e.g. *avoir besoin de, arriver à, réussir à, envier de, avoir des difficultés pour, avoir la possibilité de, avoir envie de*) are not assigned modal status.

### Adversativity {#sec-adversativity}

The factor of adversativity is reflected in the semantics of the verb and captures the negative consequences of the described action for the subject. This factor has been shown to influence the choice between the English [be]{.smallcaps} and [get]{.smallcaps} passives [@chappell1980get; @hundt_be-_2024: 89, 96-97; @bohmann_large-scale_2023], with [get]{.smallcaps} being realised more often with adversative verbs than [be]{.smallcaps}. Since Raineri [-@raineri_analyse_2010: 87] draws parallels between the [get]{.smallcaps} and [se faire]{.smallcaps} passive, we see an adversativity effect for [se faire]{.smallcaps} as likely.

@gaatone_desagreable_1983 argues that [se faire]{.smallcaps} has a strong 'unpleasant' sense (*désagreable*, @gaatone_desagreable_1983: 161), as it occurs mainly with verbs that have a negative connotation. Although Kupferman [-@kupferman_construction_1995: 77] rejects this unpleasantness as a characteristic of [se faire]{.smallcaps}, he agrees that it is more commonly used in negative contexts than [être]{.smallcaps}. This is supported by corpus data: Novakova [-@novakova_construction_2009: 110] found a clear preference for [se faire]{.smallcaps} to be realised with negative actions, with 30% of verbs in [se faire]{.smallcaps} denoting unpleasant events, and only 5% denoting pleasant events. <!-- Also: @raineri_analyse_2010 [247].-->

As @bohmann_large-scale_2023 note in their investigation of the [be:get]{.smallcaps} alternation, two hypotheses relate to verb sentiment: the adversativity hypothesis and the non-neutrality hypothesis. Both suggest that [get]{.smallcaps} favours verbs with negative connotations, but only the non-neutrality hypothesis also predicts [get]{.smallcaps} to correlate positively with positive sentiment. Since verbs with positive sentiment have been shown to appear rarely with [se faire]{.smallcaps} [@novakova_construction_2009: 110], we have chosen not to test them. In their corpus study on the [be:get]{.smallcaps} alternation, @hundt_be-_2024 categorise this factor as binary - 'adversative' versus 'non-adversative/neutral'. For practicality and clear statistical modelling, we adopt this approach, tagging verbs as adversative if the event negatively affects the subject, and non-adversative if the event is neutral or positive. In line with @bohmann_large-scale_2023 and @hundt_be-_2024, we hypothesise that [se faire]{.smallcaps} will be preferred over [être]{.smallcaps} for verbs that denote unpleasant actions (i.e. those labeled as adversative).

According to Gaatone [-@gaatone_desagreable_1983: 170], the concept of adversativity conflicts with subject responsibility in [se faire]{.smallcaps} because subjects would not willingly bring negative events upon themselves. He argues that unpleasant actions expressed with [se faire]{.smallcaps} typically have a purely passive interpretation, whereas pleasant actions tend to be initiated by the subject, resulting in a causative reading. This contrast suggests an interaction between adversativity and subject animacy that reflects the subject's affectedness. We hypothesise that adversative events are perceived differently depending on the animacy of the subject, with animate subjects being more likely to be perceived as negatively affected than inanimate subjects. Consequently, adversative verbs could favour [se faire]{.smallcaps} more strongly with animate subjects than with inanimate ones.

### Aspectuality {#sec-aspectuality}

Regarding the factor of aspectuality, Keenan and Dryer [-@keenan_passive_2007: 340] propose that "If a language has two or more basic passives they are likely to differ semantically with respect to the aspect ranges they cover." Assuming that [se faire]{.smallcaps} is a "basic passive", this suggests that [être]{.smallcaps} and [se faire]{.smallcaps} have different aspectual preferences. Aspectuality is expressed through the verb, either via inherent lexical semantics (Aktionsart) or grammatical marking (TAM: tense, aspect, mood)[^2] [@kortmann_triad_1991: 13]. Overall, we predict a preponderance of [se faire]{.smallcaps} for hitting-touching verbs and progressive aspect. <!-- We do not examine mood in this study.-->

[^2]: The general consensus in the literature is that aspect marks progressivity grammatically and Aktionsart marks it semantically. However, this clear distinction between aspect and Aktionart has been doubted, and some theories see Aktionsart und Aspect as two instantiations of Aspectuality [@Dik+1997+217+246: 221].

**Semantic Verb Clusters**

[se faire]{.smallcaps} tends to occur with dynamic processes, particularly in contexts where the source role is agentive and the target role is affected, which is the typical case of semantic transitivity [@raineri_analyse_2010: 90, 238]. Several hypotheses in the literature attempt to explain the compatibility of certain verbs with [se faire]{.smallcaps}. For example, Kupferman [-@kupferman_construction_1995: 67] and Raineri [-@raineri_analyse_2010: 90] argue that [se faire]{.smallcaps} is compatible with dynamic verbs but not with state verbs or verbs of state change with an external "cause" argument. *(However, exceptions to this pattern are noted by Novakova [-@novakova_construction_2009: 110], @martin_construction_2002 pages? and Raineri [-@raineri_analyse_2010: 100-103].)*

As reported by Raineri [-@raineri_analyse_2010: 90], Martin [-@martin_construction_2002: 14] challenges the notion that an external source always implies agency. She argues that a source can induce a change of state without intentionality and should therefore be classified as an "effector" rather than an agent. To support this claim, she contrasts two examples:

> 10. 
>
>     a.  *Je me suis fait étrangler par mon écharpe.*\
>         'I was strangled by my scarf.'
>
>     b.  *Isadora Duncan s'est fait étrangler par son écharpe qui s'est prise dans l'essieu de sa voiture.*\
>         'Isadora Duncan was strangled by her scarf, which got caught in the axle of her car.'

While the first sentence is ungrammatical, the second is more acceptable, suggesting that the role of the source does not necessarily imply agency. @martin_construction_2002 also argues that [se faire]{.smallcaps} with psychological verbs, such as in *Elle s'est fait séduire par Pierre* 'She was seduced by Pierre', is not always causative but can have a passive reading. Novakova [-@novakova_construction_2009: 110] also notes that, while rare, [se faire]{.smallcaps} can sometimes be interpreted passively with state verbs, as in *Il peine encore à se faire aimer par tous les siens* 'He still struggles to be loved by all his own'.

The semantic class of the verb has been shown to play a crucial role in the use of the passive voice in spoken French [@da2022alternance] and in the alternation between [be]{.smallcaps} and [get]{.smallcaps} passives in English [@bohmann_large-scale_2023]. In the corpus study by @da2022alternance, verbs in the *frapper* class ("hitting-touching" verbs describing violent actions) occurred predominantely in the passive voice. This aligns with findings from journalistic English, where verbs describing violent actions are frequently used in the passive voice [@henley1995syntax]. According to @da2022alternance, this preference can be explained by two factors: (1) the topical character of the undergoer affected by a hitting-touching action, which makes it more likely to be promoted to subject position, and (2) the ability to omit an unknown or less salient agent, which relates with the concept of subject responsibility. We expect this effect to hold for [se faire]{.smallcaps} as well. Even though both [se faire]{.smallcaps} and [être]{.smallcaps} topicalise the undergoer, [se faire]{.smallcaps} typically implies some degree of subject involvement. Consequently, highly dynamic verbs such as *frapper* might increase the likelihood of [se faire]{.smallcaps} over [être]{.smallcaps}, where the subject is typically a more passive patient.

As shown in @fig-clusters, some verb clusters in the study by @da2022alternance did not show a strong preference for passive constructions, with active and passive forms appearing in approximately equal proportions (e.g. verbs of perception, cognition, psychological states, behaviour, movement and transfer). However, locative and communication verbs were more frequently used in the active voice. While this preference was not statistically significant for locative verbs, communication verbs showed a significantly higher tendency towards active constructions.

![Proportions of Active and Passive Constructions by Semantic Verb Cluster [@da2022alternance: 72]](images/clusters_dacunha-abeillé.png){#fig-clusters fig-align="center" width="470"}

@da2022alternance suggest that this pattern can be explained by the verbal disposition hypothesis by @stallings1998phrasal, according to which verbs show a preference for specific syntactic constructions that speakers store in their mental lexicon. Communication verbs frequently occur with object complements, which rarely function as subjects. Since the active voice allows these verbs to maintain the object complement, they naturally favour active constructions over passive ones. @thuilier2012lemme makes a similar argument regarding communication verbs and postverbal complements. Additionally, communication verbs often feature an animate or human actor, which is more likely to be encoded as the subject in active constructions [@da2022alternance].

Hitting-touching verbs are dynamic and adversative, making them more likely to appear in passive constructions. Since these verbs emphasise the subject’s affectedness, they may be particularly compatible with [se faire,]{.smallcaps} where the subject typically is an affected undergoer. In contrast, with communication verbs, the subject is not necessarily affected in the same way. While the subject is involved in the communication event, it does not imply a significant change or affectivity, possibly making [se faire]{.smallcaps} less compatible with these verbs. Communication verbs tend to favour the active voice due to their syntactic properties, particularly their frequent use of object complements. Additionally, as they typically feature a salient animate or human subject in the corresponding active construction, they may be less likely to occur with [se faire]{.smallcaps}. We therefore hypothesise that hitting-touching verbs favour [se faire,]{.smallcaps} while communication verbs favour [être]{.smallcaps}. We expect all other classes to not have a preference for either passive variant and group them together as 'neutral'. These include psychological, behavispoken, movement, and transfer verbs. Their distribution between [être]{.smallcaps} and [se faire]{.smallcaps} may depend on additional contextual factors.

To test this hypothesis, we categorised the verbs according to their lexical class. Following @da2022alternance, we used the verb cluster categories from the online version of the lexicon "Les Verbes Français" [@lvf], based on the work by @dubois1997verbes. A table presenting the French verb clusters defined by @francois_presentation_2007, on which this dictionary is based, along with examples from our dataset, is included in the appendix of this paper. If there was no exact match, we used the closest similar class. In one case, *caillasser*, there was no entry in the lexicon: here we classified it as a hitting-touching verb, given its similarity to verbs like *attaquer*, *frapper*, or *bombarder*.

However, the effect of the verb cluster may interact with other factors, particularly animacy and adversativity. We also test the interactions with these factors. We expect the animacy of the subject to have a stronger effect if the verb is from the "hitting-touching" or the "neutral" category. [se faire]{.smallcaps} could be strongly favoured with animate subjects compared to inanimate subjects, while this effect could be weaker with communication verbs, which could favour [être]{.smallcaps} regardless of animacy. The adversativity of the verb could favour [se faire]{.smallcaps} more in communication and neutral verbs than in hitting-touching verbs. While communication verbs may favour [être]{.smallcaps} and neutral verbs may not have a strong preference, adversative constructions may shift both toward [se faire]{.smallcaps}. Since hitting-touching verbs are inherently highly adversative, this factor is unlikely to strongly affect their passive preference.

**Tense-Aspect-Mood (TAM)**

Previous literature has established that grammatical aspect plays a central role in the [être:se faire]{.smallcaps} alternation. According to Raineri [-@raineri_analyse_2010: 356ff.], among the various factors commonly associated with [se faire]{.smallcaps}, such as subject animacy, subject responsibility, and adversativity, the only necessary condition for its occurence is the type of process referred to by the verb. In her corpus, nearly all [se faire]{.smallcaps} constructions involve dynamic processes. She argues that [se faire]{.smallcaps} conveys an inchoative or durative perspective, emphasising the onset or unfolding of the event and the moment the subject begins to be affected. In contrast, she describes [être]{.smallcaps} as more typically presenting the event in its entirety. Raineri [-@raineri_analyse_2010: 356ff.] also observes a comparable distinction for the English [be]{.smallcaps}:[get]{.smallcaps} passive alternation. Bellec [-@bellec_construction_2014: 209], similarly, notes that [se faire]{.smallcaps} is often used with verbs referring to progressive or ongoing actions, whereas [être]{.smallcaps} is more likely to describe a result or endpoint.

Because tense and aspect are closely intertwined in French, we treat them as a single combined factor in our analysis. For example, constructions in the *passé composé*, which focus on completed events, may be more likely to occur with [être]{.smallcaps} than with [se faire]{.smallcaps}. In contrast, the *imparfait* emphasises the ongoing process and may therefore favour [se faire]{.smallcaps} over [être]{.smallcaps}. In the present tense, which does not focus on either endpoint or process, we expect a more even distribution of both passive variants. Descriptions of future events tend to emphasise the beginning and ongoing process of an action, rather than its endpoint, which should similarly favour [se faire]{.smallcaps} over [être]{.smallcaps} [@bellec_construction_2014: 210]. Although tense, aspect, and mood are often summarised under the term TAM, in our analysis we focus exclusively on tense and aspect, as we assume that mood does not play a crucial role in the [être:se faire]{.smallcaps} alternation. Additionally, we decided to exclude imperative constructions from the dataset (see @sec-filtering-criteria).

Consequently, rather than coding tense, aspect, and mood as separate factors, we establish a combined TAM category with two levels: progressive and non-progressive. This decision follows Bellec [-@bellec_construction_2014: 210], who argues that in French, tense and aspect are interrelated and jointly influence whether a clause is interpreted as progressive or not. Even in English, where these categories are more formally distinct, recent studies have treated them as a combined category. For instance, Bohmann et al. [-@bohmann_large-scale_2023: 39, 49] grouped tense marking on auxiliary verb forms (present, preterite, perfect, infinitive) with the progressive suffix *-ing* into a broader category and found that present and preterite forms disfavoured [get]{.smallcaps} over [be]{.smallcaps}. Similarly, Hundt et al. [-@hundt_be-_2024: 97] initially coded tense and aspect separately but ultimately interpret their effects together in terms of a TAM category.

Another indication of the association of [se faire]{.smallcaps} with progressive aspect is its frequent co-occurrence with periphrastic constructions that mark ongoing or recently initiated actions, such as *être en train de*, *venir de*, and *aller* + infinitive. These periphrastic constructions have been shown to co-occur more naturally with [se faire]{.smallcaps} than [être]{.smallcaps}, as [se faire]{.smallcaps} conveys a dynamic, ongoing action compared to the more stative [être]{.smallcaps} [@spang-hanssen_quelques_1967: 141; @novakova_construction_2009: 115-117; @raineri_analyse_2010: 90-91].

Based on these considerations, we hypothesise that progressive aspect, whether encoded through tense (*imparfait*, *futur*), periphrastic constructions (*être en train de*, *venir de*, *aller* + INF), or gerundive constructions (*se faisant*) favours [se faire]{.smallcaps}. In contrast, non-progressive aspect (*passé composé*, *plus-que-parfait*, *futur antérieur*, simple present without periphrastic construction) is expected to favour [être]{.smallcaps}.

# Data and Methods {#sec-data-and-methods}

Prior to conducting our study, we preregistered its protocol on the platform [AsPredicted](https://aspredicted.org) [@AsPredictedWharton] to enhance transparency and accountability [@huensch2024preregistration]. The preregistered study can be accessed [here](https://aspredicted.org/gp6y-dd8p.pdf).

To maintain anonymity while ensuring transparency and consistency, we refer to ourselves as Annotator A and Annotator B when describing the manual steps.

## Data Collection {#sec-data-collection}

### Corpus {#sec-corpus}

In our investigation of spoken French language, we aimed to select a corpus that authentically reflects natural language use. After careful consideration of available resources, we chose the French ParlaMint corpus version 2.1 [@11356/1431]. Although our preregistration document refers to version 2.0, this was a typographical error. Version 2.1 was the intended and only version available on [SketchEngine](http://www.sketchengine.eu) [@kilgarriff2014sketch] at the time of data collection. As such, this does not represent a deviation from our planned methodology.

ParlaMint comprises French parliamentary debates from 2017 to 2020 and contains 38,702,395 tokens (32,214,147 words). We selected this corpus not only for its size and accessibility but also because it is well-documented, POS-tagged, and contains detailed metadata. We specifically chose a spoken-language corpus because [se faire]{.smallcaps} is known to occur more frequently in speech than in writing [@raineri_analyse_2010: 105].

The genre of parliamentary debates is somewhat specialised and our results may not fully generalise to all French spoken language. However, it still offers a wide topical range and a high level of linguistic diversity. It has also been the subject of previous linguistic research [e.g. @skubic2022networks]. As such, the ParlaMint corpus provides a valuable and representative cross-section of contemporary spoken French.

### Semi-automated Approach {#sec-semi-automatic-approach}

To extract our sample, we used a semi-automatic approach combining CQL expression-based extraction and manual filtering. Our aim was to obtain an equal number of occurrences of [être]{.smallcaps} and [se faire]{.smallcaps} constructions.

First, we used a CQL expression to automatically extract a sample of 1,254 [se faire]{.smallcaps} constructions from the French ParlaMint corpus [@11356/1431]. After this, Annotator A manually filtered the data to exclude false-positives, i.e. constructions that did not meet our predefined criteria for [se faire]{.smallcaps} passives (see @sec-filtering-criteria). During the filtering process, Annotator A also annotated the lemmas of the constructions and compiled a list. After filtering, the final sample consisted of 532 valid [se faire]{.smallcaps} constructions.

```{r lemma-list, include=FALSE}

lemmas <- sefaire_sample |> 
  count(lemma)

```

Next, we extracted the [être]{.smallcaps} constructions using a CQL expression in SketchEngine. To filter the concordances, we used the list of lemmas derived from the [se faire]{.smallcaps} sample, ensuring comparability between the two construction types by controlling for verb-specific differences that might influence the results. Due to SketchEngine’s download limit, we retrieved the [être]{.smallcaps} constructions in four subsamples, each containing between 9,295 and 9,794 concordances. After downloading, all filtered instances were stored in a .csv file and merged into a single dataframe in `R`. To ensure reproducibility, we set a random seed using the function `set.seed()`, with the seed value derived from the lottery numbers on Wednesday, 12th March 2025, as listed on the official [Lotto website](https://www.lotto.de/lotto-6aus49/lottozahlen). The last digit had to be removed to match the required length. We then randomly sampled a subset of concordance lines that was equal in size to the manually filtered [se faire]{.smallcaps} sample (i.e. 532 instances). After sampling, one false positive was detected in the [être]{.smallcaps} data and was subsequently removed. At this stage, complete resampling was not feasible. However, since setting a fixed seed ensures that the order of concordance retrieval remains constant, we retrieved the 533rd concordance using the same seed and added it to the sample, thereby matching the number of 532 data points across both variants.

### Manual Filtering Exclusion Criteria and Borderline Cases {#sec-filtering-criteria}

For our study of [être]{.smallcaps} and [se faire]{.smallcaps} as a passive alternation, our main criterion was to include only those [se faire]{.smallcaps} constructions that have a passive (and not an exclusively causative) reading, i.e. constructions that could be paraphrased with [être]{.smallcaps} (see @sec-sefaire-context). Consequently, we excluded expressions that could not be rephrased with [être]{.smallcaps}, either because the [être]{.smallcaps} construction was ungrammatical or the meaning was fundamentally altered.

We defined the following additional criteria, which follow from our main criterion:

-   **Transitivity of the verb**: the verb is a transitive verb (the direct object of the active sentence is the subject of the passive sentence). As described in @sec-etre, [être]{.smallcaps} requires a direct object in the active equivalent which becomes the subject in the passive voice. For reflexive and ditransitive verbs, [être]{.smallcaps} is not available and does not compete with [se faire]{.smallcaps}, which is why we excluded such constructions from our analysis. Additionally, we excluded impersonal passive constructions (*passif impersonnel*) due to the lack of a subject referent.

-   **Exclusion of fixed idiomatic expressions**: the [se faire]{.smallcaps} construction is not a fixed idiomatic expression with a figurative meaning. We specifically excluded constructions classified as *idiomes lexicalisés* ('lexicalized idioms') [@raineri_analyse_2010: 163]. These constructions neither carry a true causative nor passive meaning[^3] and cannot occur with a *complément d'agent*: *se faire entendre*, *se faire sentir*, *se faire attendre*, and *se faire désirer* [@raineri_analyse_2010: 163]. Examples of additional fixed idiomatic expressions that we filtered out include *se faire passer pour* 'to impersonate sb.', *se faire avoir* 'to be fooled', *se faire mousser* 'to boast/brag', and *se faire taper sur les doigts* 'to get a slap on the fingers, to be scolded'.

-   **Exclusion of adjectival constructions**: we excluded [être]{.smallcaps} constructions that are essentially adjectival constructions, such as *elle est connue* 'she is known'. Accordingly, [se faire]{.smallcaps} constructions that would typically be rephrased with an adjectival [être]{.smallcaps} construction were also excluded.

[^3]: These verbs mainly occur with inanimate subjects (effects, consequences, discontent). Raineri [-@raineri_analyse_2010: 163-164] proposes that the causative meaning is not available for these verbs since the subject is not causing the event. There is also no passive meaning since they cannot cannot occur with a *complément d’agent* and do not have a paraphrase with être.

To ensure consistency and practicality, we decided to include or exclude verbs based on their lemma, rather than on individual occurrences. After manually evaluating all occurrences, the following lemmas were excluded from our sample: *entendre*, *sentir*, *ressentir*, *attendre*, *désirer*, *voir*, *connaître*, *passer (pour)*, *mousser*, *avoir*, *rouler (se faire rouler dans la farine)*, *mener (se faire mener par le bout du nez)*.

We also excluded imperative constructions because of their rarity in the passive and because of the semantic contradiction. While the passive focuses on the undergoer of an action, the imperative demands the performance of an action. We anticipated that imperatives would be unevenly distributed, potentially leading to statistical distortions.

Regarding tense, we chose to include all tenses in our analysis, including the present tense, based on previous literature suggesting that the present tense favours [se faire]{.smallcaps} over [être]{.smallcaps} (see @sec-aspectuality). As part of the TAM factor, tense is an important feature in our study. This approach contrasts with Da Cunha and Abeillé [@da_cunha_lalternance_2020; @da2022alternance], who focused exclusively on passive constructions in the compound past tense (i.e. *NP a été VP*) to prioritise verbal passives over adjectival passives, such as *La porte est ouverte* 'The door is open'. By including all tenses, we ensured to include as many [se faire]{.smallcaps} constructions as possible and obtain a sufficient number of occurrences for the analysis. This decision was therefore motivated both by the aim to capture the overall syntactic range of the passive alternation and by the importance of tense within the TAM factor. While we recognise that this approach increases complexity, we believe that it allows for a more comprehensive analysis of the [être:se faire]{.smallcaps} alternation.

A typical example of a [se faire]{.smallcaps} construcion that met our criteria is:

> *Les élus se font très souvent insulter \[...\]\
> *'The elected officials are often insulted \[...\]'

The corresponding [être]{.smallcaps} paraphrase would be:

> *Les élus sont très souvent insultés.\
> *'The elected officials are often insulted.'

In contrast, expressions such as *se faire connaître* were excluded. For example:

> *Cette baisse est ensuite une conséquence de la crise économique de 2018 dont les effets se font connaître aujourd'hui.\
> *'This decline is then a consequence of the 2018 economic crisis, the effects of which are becoming known today.'

When paraphrased with [être]{.smallcaps}, *connus* takes on an adjectival function:

> *Cette baisse est ensuite une conséquence de la crise économique de 2018 dont les effets sont connus aujourd'hui.*\
> 'This decline is then a consequence of the 2018 economic crisis, the effects of which are known today.'

While we excluded purely causative constructions, the distinction between causative and passive constructions was not always clear. The question arose: when is a construction semantically close enough to a passive alternation, and when does its meaning lean too much toward a causative interpretation? As expected, given the ongoing debate about the passive status of [se faire]{.smallcaps} (see @sec-debate) and the difficulties noted by Bohmann et al. [-@bohmann_large-scale_2023] in their study of the English [be]{.smallcaps}:[get]{.smallcaps} passive alternation, we encountered a number of unclear cases during the manual filtering of constructions. We acknowledge that, similar to their study, there is no straightforward distinction between causative and passive readings, and some degree of subjectivity in categorization is unavoidable.

This approach is also supported by linguistic analyses that suggest a continuum between causative and passive meanings in constructions with *se faire* (see @sec-debate). We identified numerous cases where a clear binary distinction between passive and causative was not possible, for example when the subject is both the undergoer and the causer of the action. In general, we opted for a broad approach, including as many constructions as possible. This means that even when the causative reading appeared more natural (because the subject initiates the event), we included the construction as long as a passive understanding was also plausible (since the subject also undergoes the action). Representative examples of such constructions in our sample include *se faire vacciner* 'to be/get vaccinated', *se faire élire* 'to be/get elected' or 'to stand for election', *se faire remarquer* 'to be/get noticed' or 'to draw attention to oneself'.

### Inter-Annotator Agreement {#sec-iaa}

In preparation for statistical modelling, we annotated the filtered sample for the features outlined in @sec-factors. A blind annotation process was used, where Annotator A and Annotator B were unaware of each other’s annotations. As [se faire]{.smallcaps} and [être]{.smallcaps} data were collected separately, some steps were necessary before conducting analyses on the data. First, the data was cleaned by aligning the speaker and reference columns. Since annotator A annotated all *compléments d' agents* of [se faire]{.smallcaps} during initial filtering, the column `agent_pp` was renamed to `agent_pp_A`. The datasets were then combined, variables were converted from characters to factors and empty values were replaced with `NA` to ensure proper handling of missing values in statistical analysis.

```{r read-annotated-data, include = FALSE}
sefaire_sample_annotated <- read.csv(file = here("data", "sefaire_sample_annotated.csv"))
etre_sample_annotated <- read.csv(file = here("data", "etre_sample_annotated.csv"))
```

```{r data-cleaning-colnames, include = FALSE}
#| code-fold: true
# data preprocessing
colnames(etre_sample_annotated)
colnames(sefaire_sample_annotated)

# Differences between the two datasets mainly regard columns speaker and reference: se faire sample has a "speaker" column which includes the speaker id and the name and a "reference" column that includes the speaker ID and the date; être sample has only a "reference" column that includes speaker ID, date, name
```

```{r data-cleaning-colnames-2, echo = FALSE}
#| code-fold: true
# Goal: two columns consistent across both datasets, one "speaker" (name), one "reference" (ID), drop date
# 1. se faire
# separate "speaker" column into reference and speaker
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  mutate(reference = sapply(str_split(speaker, ",", n = 2), `[`, 1),
         speaker = sapply(str_split(speaker, ",", n = 2), `[`, 2))

# remove now redundant "Reference", remove "Kommentar" column, put new "reference" column in first position
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  dplyr::select(-Reference, -Kommentar) |> 
  relocate(reference, speaker)

# 2. être
# separate "Reference" column into reference and speaker
etre_sample_annotated <- etre_sample_annotated |> 
  mutate(reference = sapply(str_split(Reference, ",", n = 3), `[`, 1),
         speaker = sapply(str_split(Reference, ",", n = 3), `[`, 3))

# remove now redundant "Reference" column, remove "Kommentar" column, put new "reference" und "speaker" column in first position
etre_sample_annotated <- etre_sample_annotated |> 
  dplyr::select(-Reference, -Kommentar) |> 
  relocate(reference, speaker)

#still some differences left: être has an "ID" column which se faire does not; être has two agent_pp columns because it was annotated by A and B while se faire only has one, since agent_pp of se faire constructions was annotated there one person (annotator A)

#rename agent_pp for clarity
sefaire_sample_annotated <- sefaire_sample_annotated |> 
  rename(agent_pp_A = agent_pp)
```

```{r combine-data, include = FALSE}
#| code-fold: true
#combine the data
passive_sample_combined <- bind_rows(etre_sample_annotated, sefaire_sample_annotated) 
```

```{r data-cleaning-2, include = FALSE}
#| code-fold: true
# check for typos/ tagging inconsistencies in combined data; convert to factor; add "NA" as value for all cases that are only annotated by one person
#check data type
class(passive_sample_combined$subj_animacy_A)
#"character", convert to factor and fill "empty" values with NA
passive_sample_combined <- passive_sample_combined |> 
  mutate(across(where(is.character), ~ na_if(.x, ""))) |> 
  mutate(across(where(is.character), factor)) |> 
  mutate(across(where(is.factor), droplevels)) # to not include NA as a level
# summary
passive_sample_combined |> 
  dplyr::select(subj_animacy_A, subj_animacy_B, subj_gender_A, subj_gender_B, agent_pp_A, agent_pp_B, modal_A, modal_B, tam_A, tam_B, cluster_A, cluster_B, adversativity_A, adversativity_B) |> 
  summary()
```

Since the analysis required detailed (and time-consuming) manual annotation, we decided to split the annotation work between us. However, to ensure consistent annotation, a representative sample of 320 observations (`r round(320/1064*100, 2)` of all instances) were annotated by both of us. For these instances, we calculated Cohen's ϰ (Kappa) as a measure of inter-annotator agreement, following Levshina [-@levshina_how_2015: 201]. The values and their interpretation are displayed in @tbl-kappa-values.

```{r cohens-kappa, include = FALSE}
kappa_animacy <- ckappa(passive_sample_combined[c("subj_animacy_A", "subj_animacy_B")]) 
kappa_animacy #0.9096219

kappa_gender <- ckappa(passive_sample_combined[c("subj_gender_A", "subj_gender_B")])
kappa_gender #0.8746109

kappa_agent_pp <- ckappa(passive_sample_combined[c("agent_pp_A", "agent_pp_B")])
kappa_agent_pp #0.9118457

kappa_modal <- ckappa(passive_sample_combined[c("modal_A", "modal_B")])
kappa_modal #0.9469328

kappa_adversativity <- ckappa(passive_sample_combined[c("adversativity_A", "adversativity_B")])
kappa_adversativity #0.6700877

kappa_cluster <- ckappa(passive_sample_combined[c("cluster_A", "cluster_B")])
kappa_cluster #0.8326557

kappa_tam <- ckappa(passive_sample_combined[c("tam_A", "tam_B")])
kappa_tam  #0.7702391

kappa_values <- c(kappa_animacy$kappa, kappa_gender$kappa, kappa_agent_pp$kappa, kappa_modal$kappa, kappa_adversativity$kappa, kappa_cluster$kappa, kappa_tam$kappa)

mean_kappa <- mean(kappa_values)
mean_kappa #0.845142, very high agreement :)
```

```{r cohens-kappa-table, echo = FALSE}
kappa_values <- c(kappa_values, mean_kappa)

factors_passive <- c("Subject Animacy", "Subject Gender", "Complément d'Agent", "Modal", "Adversativity", "Cluster", "TAM","Mean")

kappa_strength <- c("Almost Perfect", "Almost Perfect", "Almost Perfect", "Almost Perfect", "Substantial", "Almost Perfect", "Substantial", "Almost Perfect")

kappa_table <- tibble(
  Factor = factors_passive,
  Kappa = kappa_values,
  Interpretation = kappa_strength
)

```

```{r tbl-kappa-values, echo = FALSE}
#| label: tbl-kappa-values
#| tbl-cap: Cohen's ϰ Values and their Interpretation, Following Landis and Koch (1977)

kappa_table |> 
  kbl(digits = 2, col.names = c("Factor", "Cohen's ϰ", "Strength of Agreement" )) |>
  row_spec(8, bold = T)

```

We classified the ϰ-values using the influential benchmark categorisation of Landis and Koch [-@landis1977measurement: 165], who define an agreement of 0.81 - 1.00 as "Almost Perfect". The mean inter-annotator agreement is high at `r  round(mean_kappa, 2)`. Agreement is lowest for the cateogories adversativity (`r round(kappa_adversativity$kappa, 2)`) and TAM (`r round(kappa_tam$kappa, 2)`).

### Data Preprocessing {#sec-data-preprocessing}

For the main analysis, the combined data of [se faire]{.smallcaps} and [être]{.smallcaps} data were merged into one dataset where each column corresponds to one factor. The columns indicating factors include the annotations of the first annotator (i.e. annotator A for half of the observations, annotator B for the other half). The sample comprises 1064 observations, 532 for [se faire]{.smallcaps} and [être]{.smallcaps} each. We looked at the data distribution among the categories and levels to identify scarce combinations and patterns that could be problematic for analysis.

```{r merge-to-main-dataset}
#| code-fold: true
#goal: one dataset with one column for each factor, the column includes the annotation of the "main" annotator
# special case agent pp column: if passive variant is se faire -> take value from column "agent pp_A", if passive variant is être is: same procedure as with the other variables
passive_sample_clean <- passive_sample_combined |> 
  mutate(
    subj_animacy = case_when(annotator == "A" ~ subj_animacy_A,
                             annotator == "B" ~ subj_animacy_B),
    subj_gender = case_when(annotator == "A" ~ subj_gender_A,
                            annotator == "B" ~subj_gender_B),
    agent_pp = case_when(passive_variant == "se faire" ~ agent_pp_A,
                         passive_variant == "être" ~ case_when(
                           annotator == "A" ~ agent_pp_A,
                           annotator == "B" ~ agent_pp_B)),
    modal = case_when(annotator == "A" ~ modal_A,
                      annotator == "B" ~ modal_B),
    tam = case_when(annotator == "A" ~ tam_A, 
                    annotator == "B" ~ tam_B),
    cluster = case_when(annotator == "A" ~ cluster_A, 
                        annotator == "B" ~ cluster_B),
    adversativity = case_when(annotator == "A" ~ adversativity_A, 
                               annotator == "B" ~ adversativity_B)
  )

# remove now-redundant A and B columns, remove ID (irrelevant since it's only there for être)
passive_sample_clean <- passive_sample_clean |> 
  dplyr::select(-subj_animacy_A, -subj_animacy_B, -subj_gender_A, -subj_gender_B, -agent_pp_A, -agent_pp_B, -modal_A, -modal_B, -tam_A, -tam_B, -cluster_A, -cluster_B, -adversativity_A, -adversativity_B, -ID)
```

```{r first-look-main-data, include = FALSE}
#| code-fold: true
table(passive_sample_clean$passive_variant, passive_sample_clean$subj_animacy)
table(passive_sample_clean$passive_variant, passive_sample_clean$subj_gender)
table(passive_sample_clean$passive_variant, passive_sample_clean$agent_pp)
table(passive_sample_clean$passive_variant, passive_sample_clean$modal)
table(passive_sample_clean$passive_variant, passive_sample_clean$tam)
table(passive_sample_clean$passive_variant, passive_sample_clean$cluster)
table(passive_sample_clean$passive_variant, passive_sample_clean$adversativity)
```

We observed certain patterns in the distribution of passive constructions across the levels of each factor: the subject animacy level "inanimate" has comparatively few occurrences with [se faire]{.smallcaps}. The subject gender level "mixed" appears infrequently overall and was excluded to avoid distorting the analysis and to ensure interpretability.

The subject gender level "unknown" appears much more often for [se faire]{.smallcaps} than for [être]{.smallcaps}. We suspected that this may be an annotation-related effect: when annotating gender, we took grammatical gender marking into account. Since [être]{.smallcaps} is realised with a participle that always agrees in gender with the subject, gender was always identifiable, and only impersonal constructions were annotated as "unknown". In contrast, [se faire]{.smallcaps} is very common in infinitive constructions, which do not indicate gender. Therefore, when gender was not inferable from the context, it was labelled "unknown". We suspect that this created an artifact, as "unknown" is strongly correlated with infinitive constructions, and thus with [se faire]{.smallcaps}. To test this, we created two separate datasets, one with "unknown" and one without, which will be compared in the analysis. The dataset without "unknown" contained 503 [être]{.smallcaps} and 367 [se faire]{.smallcaps} constructions.

The *complément d'agent* level "de" is also relatively rare with [se faire,]{.smallcaps} but we decided to keep it for interpretative reasons. However, we flagged it as a potential problem and kept the option open to exclude it later, if it proved to have a negative impact on the analysis.

```{r cleaning-main-data, include=FALSE}
# exclude "mixed" gender because of low frequency (would likely skew the data)
#sample with "unknown" kept
passive_sample_unknown <- passive_sample_clean|>
  filter(subj_gender != "mixed") |>
  droplevels()

# exclude "unknown" gender (needs to be reasoned and explained!)
# (merge "de" and "par" into a "present category")
passive_sample_clean <- passive_sample_clean|> 
  filter(subj_gender != "mixed") |> 
  filter(subj_gender != "unknown") |>
  droplevels()

# passive_sample_clean <- passive_sample_clean |> 
#   mutate(agent_pp = case_when( #to merge par and de
#     agent_pp %in% c("par", "de") ~ "present",
#     TRUE ~ agent_pp
#   )) |> 
#   mutate(agent_pp = factor(agent_pp))

#check levels
levels(passive_sample_clean$subj_gender)
levels(passive_sample_unknown$subj_gender)

#check passive distribution without unknown
table(passive_sample_clean$passive_variant)

```

## Binomial Logistic Regression {#sec-log-regression}

In addition to conducting a descriptive analysis of the data, we employed a binomial logistic regression model to test our hypotheses. This method provides a logit value for each variable that "represents the chances of construction A to be chosen in a particular type of context compared with the chances of construction B to be used in the same type of context" [@levshina_how_2015: 253]. The model allowed us to exactly track the influence of factors and their interactions on the chances of [se faire]{.smallcaps} being used over [être]{.smallcaps}.

### Levels {#sec-levels}

The levels of each factors are ordered alphabetically by default. We rearranged the reference levels to be most useful for both modelling and interpreting: where possible, we chose the most unmarked case as the reference level (passive variant: être; *complément d' agent*: 0; modal: no; TAM: non-progressive, cluster: neutral). If this would negatively affect the stability of the model, we instead selected the most frequent level (subject animacy: animate, subject gender: masculine). In the case of adversativity, the most balanced level (adversative) was chosen to facilitate interpretation and maintain consistency with the other factors.

```{r reference-levels, include = FALSE}
#change reference levels: "animate" to "inanimate", "communication" to "neutral", "adversative" to "non-adversative"
#passive_sample_clean$subj_animacy <- relevel(passive_sample_clean$subj_animacy, ref = "inanimate")
passive_sample_clean$subj_gender <- relevel(passive_sample_clean$subj_gender, ref = "masculine")
passive_sample_clean$cluster <- relevel(passive_sample_clean$cluster, ref = "neutral")
#passive_sample_clean$adversativity <- relevel(passive_sample_clean$adversativity, ref = "non-adversative")

ftable(passive_sample_clean$passive_variant, passive_sample_clean$subj_animacy, passive_sample_clean$adversativity)

passive_sample_unknown$subj_gender <- relevel(passive_sample_unknown$subj_gender, ref = "masculine")
#passive_sample_unknown$subj_animacy <- relevel(passive_sample_unknown$subj_animacy, ref = "inanimate")
passive_sample_unknown$cluster <- relevel(passive_sample_unknown$cluster, ref = "neutral")
#passive_sample_unknown$adversativity <- relevel(passive_sample_unknown$adversativity, ref = "non-adversative")

levels(passive_sample_clean$subj_gender)
levels(passive_sample_clean$agent_pp)
levels(passive_sample_unknown$subj_gender)
```

```{r levels-table, echo = FALSE}
factors <- c("Passive Variant", "Subject Animacy", "Subject Gender", "Complément d'Agent", "Modal", "Adversativity", "Cluster", "TAM")

labels <- c("passive_variant", "subj_animacy", "subj_gender", "agent_pp", "modal", "adversativity", "cluster", "tam")

levels <- c("être, se faire", "animate, inanimate", "masculine, feminine", "0, de, par", "no, yes", "adversative, non-adversative", "neutral, communication, hitting-touching", "non-progressive, progressive")

levels_table <- tibble(
  Factor = factors,
  Label = labels,
  Levels = levels
)
```

```{r tbl-levels-kable, echo = FALSE}
#| label: tbl-levels
#| tbl-cap: Overview of Variables and their Reference Levels

kbl(levels_table)

```

### Model Selection {#sec-model-selection}

We fit a mixed-effects binomial logistic regression model in `R` following Levshina [-@levshina_how_2015: 253-276] and Gries [-@gries_statistics_2021: 393-452]. The model estimates the log of the odds that a given construction is realized with [se faire]{.smallcaps}. To account for individual variation, we included the verb lemma and the speaker (as indicated in the meta data of ParlaMint 2.1) as random effects.

We set a significance threshold $p$-value of 0.05 and applied backward stepwise selection [@gries_statistics_2021: 366]. Model quality was monitored and compared using the Aikaike Information Criterion (AIC), with preference given to the model with the lower AIC value [@gries_statistics_2021: 366]. Following Gries [-@gries_statistics_2021: 366], non-significant factors were removed one-by-one, starting with the one with the highest non-significant $p$-value. Additionally, we evaluated model fit using R^2^, aiming to find the model with the highest possible R^2^, that therefore explains most of the variance.

::: panel-tabset
#### Maximal Model

First, we constructed a model that includes all the factors described in @sec-levels, the interactions we hypothesised to be significant, and the random effects of speaker and lemma. While the model does not include all theoretically possible interactions, we considered it maximal in the sense that it encompasses all factors and interactions identified in the literature review as potentially meaningful and contributing to passive choice.

`passive_variant ~ subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity + (1|lemma) + 1|speaker)`

```{r max-model-unknown, output = FALSE}
#| code-fold: true
model_max_unknown <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity) + (1|speaker) + (1 | lemma),
    family = "binomial", 
    data = passive_sample_unknown,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_max_unknown) #AIC: 771.3 
```

@tbl-model-max-unknown shows the odds ratios, confidence intervals and $p$-values of the maximal model, along with the random effects, R^2^ and AIC. Although the main effect of `subj_gender` could not be interpreted due to its participation in an interaction, it is striking that the odds ratios of `subj_gender[unknown]` are very high with a huge confidence interval of 10.49 – 124.99. The AIC is `r round(AIC(model_max_unknown), 2)` with an R^2^ of 0.599 when excluding random effects and 0.748 when accounting for them. The vif-analysis, as displayed in @tbl-max-model-collinearity, revealed no high collinearity values for `subj_gender` or the interaction `subj_animacy:subj_gender`. However, as explained in @sec-data-preprocessing, we suspected an annotated-related artifact related to the level `unknown`. For that reason, we decided to compare the model to one fitted on the dataset without `unknown` constructions in the next step.

```{r max-model-unknown-tab, echo = FALSE}
#| label: tbl-model-max-unknown
#| tbl-cap: Results of the Maximal Logistic Regression Model Fitted on the Full Dataset

tab_model(model_max_unknown, show.aic = TRUE) # Marginal R2 / Conditional R2: 0.599 / 0.748
```

**Collinearity**

```{r collinearity-model-unknown, echo = FALSE}
#| label: tbl-max-model-collinearity
#| tbl-cap: Collinearity Values of the Maximal Logistic Regression Model

kbl(vif(model_max_unknown))
```

#### Model without `unknown`

```{r full-passive-glmer, output = FALSE}
#| code-fold: true
model_max <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity) +
               (1 | lemma) + (1 | speaker),
    family = "binomial",
    data = passive_sample_clean,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_max) #AIC: 732.5
vif(model_max)
```

After excluding `unknown`, the model's AIC improved substantially to `r round(AIC(model_max), 2)` (@tbl-model-max). The model has an R^2^ of 0.46 / 0.67. Although this is lower than for the model with `unknown`, and therefore explains less variance, overall, the model seems to make better (i.e. more stable and more precise) predictions. Since the variance explained by the random effects is larger than zero, the random effects had to be kept. Due to its better performance, the data without `unknown` were used in future models. However, the model with `unknown` was also optimised using backward stepwise selection and is reported in the appendix.

```{r full-passive-glmer-tab, echo = FALSE}
#| label: tbl-model-max
#| tbl-cap: Results of the Maximal Logistic Regression Model without `unknown`
tab_model(model_max, show.aic = TRUE) #Marginal R2 / Conditional R2:	0.460 / 0.670
```

#### Model without Non-significant Interactions

```{r model-back-1, output = FALSE}
#| code-fold: true
model_back_1 <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:adversativity) + (1 | lemma) + (1 | speaker),
    family = "binomial",
    data = passive_sample_clean,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_back_1) #AIC: 725.4
vif(model_back_1)
```

All non-significant interactions (`cluster:adversativity`, `subj_animacy:cluster`, `subj_animacy:subj_gender`) were removed, improving the AIC to `r round(AIC(model_back_1), 2)` with an R^2^ of 0.450 / 0.664 (@tbl-model-back-1). Thus, without the interactions, model accuracy increased while the explained variance decreased slightly. With $p$-values \> 0.05, the categories `subj_gender`, `agent_pp`, `modal` and `tam` were still found to be non-significant.

```{r model-back-1-tab, echo = FALSE}
#| label: tbl-model-back-1
#| tbl-cap: Results of the Logistic Regression Model without Non-significant Interactions
tab_model(model_back_1, show.aic = TRUE) #	0.450 / 0.664 
```

#### Final Model

```{r final-model, output = FALSE}
#| code-fold: true
final_model <- glmer(
  passive_variant ~ (subj_animacy + subj_gender + cluster + adversativity + subj_animacy:adversativity) + 
                    (1 | lemma) + (1 | speaker),
  family = "binomial",
  data = passive_sample_clean,
  control = glmerControl(optimizer = "bobyqa")
)
final_model_summary <- summary(final_model) # AIC: 719.1
vif(final_model)
```

Non-significant fixed effects were removed one-by-one, starting with `tam`, which has the highest $p$-value. At each step, the main effect with the highest $p$-value was eliminated. The results of the final model are reported in @tbl-final-model. Although `subj_gender` remained non-significant ($p$ \> 0.05), we retained it for theoretical and interpretative reasons. The AIC of `r round(AIC(final_model), 2)` indicates that removing `agent_pp`, `modal` and `tam` improved the model quality while R^2^ stayed constant at 0.451 / 0.662.

```{r final-model-tab, echo = FALSE}
#| label: tbl-final-model
#| tbl-cap: Results of the Logistic Regression Model Chosen as Final
tab_model(final_model, show.aic = TRUE) #Marginal R2 / Conditional R2	0.451 / 0.662
```
:::

### Model Diagnostics {#sec-model-diagnostics}

**Residuals**

To check goodness of fit and usability of the final model, we examined the residuals visually. Gelman and Hill [-@Gelman_Hill_2006: 97] state that due to to the discreteness of data in logistic regression, "plots of raw residuals from logistic regression are generally not useful". Instead, they recommend plotting binned residuals. Following this suggestion, we checked the residuals of the model using the `binnedplot()` function from the `arm` package [@R-arm], which sorts residuals into categories, so called "bins". By default, the number of bins equals the rounded squareroot of the total amount of data points (for our model: `r floor(sqrt(848))`).

```{r residuals-diagnostics-binned, echo = FALSE}
#| label: fig-binned-residuals
#| fig-cap: Plot of Binned Residuals

predictions <- predict(final_model, type = "response", na.action = na.omit)
binnedplot(predictions, residuals(final_model, type = "response"))
```

At first glance, the binned residuals visualised in @fig-binned-residuals appear to be quite random with no obvious patterns and with most residuals close to 0. The grey lines mark the bounds of the standard error; ideally, these bounds should emcompass 95% of the binned residuals [@Gelman_Hill_2006: 97]. Although the SE bounds essentially capture the shape of the residuals and the majority of the residuals fall within them, a considerable proportion also fall outside the desired area: 11 out of 29 (`r paste0(round(11/29, 2) * 100, "%")`), which is well above the desired 5%.

In this context, a slight pattern emerges: negative deviations are more common for low to medium expected values (i.e. low predicted probability of [se faire]{.smallcaps}), while positive deviations occur more often for high expected values (i.e. high predicted probability of [se faire]{.smallcaps}). In the mid-range (0.3-0.7), where [être]{.smallcaps} and [se faire]{.smallcaps} are predicted to be about equally likely, the binned residuals are exclusively negative (`y < 0`). Thus, when the correct level was [être]{.smallcaps}, the model often predicted a higher probability of [se faire]{.smallcaps}. This pattern is reversed for predicted probabilities of `x > 0.7`: all deviations are positive (`y > 0`), indicating that when [se faire]{.smallcaps} was the correct variant, the model slightly underestimated the probabilities of [se faire]{.smallcaps} occurring and could have been even more confident. In addition, there is one extremely negative bin between `x = 0.4 - 0.5`, indicating that the model drastically overestimated [se faire]{.smallcaps} in this range.

To investigate these potentially problematic characteristics further and determine whether they indicate systematic prediction errors of the model, we conducted additional residuals tests. Following the recommendations of Gries [-@gries_statistics_2021: 445], we used functions from the `DHARMa` package [@R-DHARMa] to gain more detailed insight into the residuals structure.

```{r residuals-diagnostics-DHARMa, include = FALSE}
sim_res <- simulateResiduals(final_model)
testDispersion(sim_res)
testUniformity(sim_res)
testOutliers(sim_res)
testQuantiles(sim_res)
```

```{r residuals-diagnostics-DHARMa-dispersion, echo = FALSE}
#| label: fig-dharma-dispersion
#| fig-cap: DHARMa Dispersion Plot of Residuals
testDispersion(sim_res)
```

@fig-dharma-dispersion illustrates that the residuals are slightly positively-skewed but approximately normally distributed. The results of the dispersion test indicate no evidence of over- or underdispersion (dispersion = `r round(testDispersion(sim_res)$statistic[[1]], 3)`, $p$ = `r round(testDispersion(sim_res)$p.value, 3)`). Since $p$ is greater than 0.05, the "null hypothesis of normality cannot be rejected" [@levshina_how_2015: 162], suggesting that the deviation is non-significant.

```{r residuals-diagnostics-DHARMa-qq, echo = FALSE}
#| label: fig-dharma-qq
#| fig-cap: DHARMa Q-Q Plot Testing the Deviation of Residuals
plotQQunif(sim_res)
```

The Q-Q Plot (@fig-dharma-qq) reveals a significant deviation ($p$ = `r round(testUniformity(sim_res)$p.value, 3)`) as tested by the Kolmogorov-Smirnov test, which may point to unreliable inferential measures. However, as Levshina [-@levshina_how_2015: 56] notes, deviation tests depend on sample size: in large samples, test algorithms are highly sensitive to small deviations, many of which naturally occur in large datasets such as ours. Levshina [-@levshina_how_2015: 56] therefore recommends prioritising visual inspection over test results. Based on the overall distribution of the residuals, we assessed the magnitude of the deviation to be minor. The outlier test shows no significant deviations from the expected distribution.

Overall, we did not interpret the `DHARMa` plots as evidence of poor model performance. Among the residual test results, the most concerning was the pattern of non-optimal binned residuals. (@fig-binned-residuals). They revealed the model's overall overoptimism toward [se faire]{.smallcaps}, indicating that it slightly overpredicts [se faire]{.smallcaps} when [être]{.smallcaps} is correct. On the other hand, for very high probabilities of [se faire,]{.smallcaps} the model could be more confident in its predictions of [se faire]{.smallcaps}. Considering all residual diagnostics, we understand them to indicate a stable model structure, but we acknowledge that the model is by no means perfect and could be further improved.

**Concordance Index**

```{r concordance-index, echo = FALSE}
# outcome <- as.numeric((passive_sample_clean)$passive_variant) - 1

outcome <- as.numeric(na.omit(passive_sample_clean)$passive_variant) - 1

concordance_final_model <- Hmisc::somers2(predictions, outcome)["C"]
```

As Levshina [-@levshina_how_2015: 256] describes, the concordance index measures the ability of the model to discriminate correctly by providing the proportion of times when the model predicts the use of the variant correctly. Our model has a high concordance index of `r   round(concordance_final_model, 2)`. According to the scale reported in Levshina [-@levshina_how_2015: 259], this indicates "outstanding discrimination".

**Significance**

Additonally, we applied an overall significance test by comparing our final model to a model without predictors using an ANOVA [@gries_statistics_2021: 445-446].

```{r model2-significance, include = FALSE}
null_model <- glmer(passive_variant ~ 1 + (1 | lemma) + (1 | speaker), data = passive_sample_clean, family = binomial)

# null_model <- glmer(passive_variant ~ 1 + (1 | lemma), data = passive_sample_clean, family = binomial)

anova_final_model <- anova(final_model, null_model, test="Chisq")
```

```{r anova-tab, include = FALSE}
kbl(anova_final_model)
 #relative likelihood
(exp((AIC(null_model)-AIC(final_model))/2))

rel_likelihood <- exp((AIC(null_model) - AIC(final_model)) / 2)

# ANOVA to data frame
anova_df <- as.data.frame(anova_final_model)
anova_final_model

# p-value
anova_df$`Pr(>Chisq)` <- format(anova_df$`Pr(>Chisq)`, scientific = TRUE, digits = 3)

# add relative likelihood
anova_df$`Relative Likelihood` <- c(NA, rel_likelihood)
```

```{r anova-kable-tab, echo = FALSE}
#| label: tbl-anova
#| tbl-cap: ANOVA of the Final Model
kbl(anova_df, digits = 3) |> 
  kable_styling(bootstrap_options = c("hover","responsive")) |> 
  row_spec(2, bold = TRUE)
```

The ANOVA, as captured by @tbl-anova, shows that our model is highly significant (LR-test=181.02, $p$ \<0.0001), indicating a strong improvement over the null model. The model has a vastly higher relative likelihood compared to the null model.

**Bootstrap Validation**

We used bootstrapping to validate the stability of our model's coefficient estimates. This method involves random resampling with replacement from the original dataset, analysing each resample, and comparing the resulting estimates to the original. It allows for internal model validation without the need for new external test data [@levshina_how_2015: 166].

For this, the `bootMer()` function from the `lme4` package [@R-lme4] was applied to our mixed-effects model, resampling 1000 times with the same random seed that we used during data sampling for reproducibility. When running `bootMer()` to validate the stability of the fixed effects in our model, we encountered singularity warnings in most of the bootstrap iterations. Convergence during bootstrapping was achieved only when the `bobyqa` optimiser was used. Although `isSingular()` returned `FALSE` for the final model, the instability across resamples indicates limited reliability. Removing the `speaker` effect resolved the singularity issues. This suggests that there may be too little variance between speakers to reliably estimate this random effect. However, removing `speaker` also increases the model’s AIC from 719.2 to 732.0, indicating a poorer overall fit. Additionally, excluding the random effect would change the model’s structure and require re-specification and re-evaluation, which exceeds the scope of this term paper. For interpretation, we therefore rely on our original model, while acknowledging its limitations in terms of robustness. We consider the simplified model without `speaker` (see @tbl-model-without-speaker) a potentially more stable alternative.

```{r bootstrap-validation, include = FALSE}

# bootstrap_final_model <- bootMer(
# final_model,
# FUN = fixef,
# nsim = 1000,
# seed = 2142637384,
# type = "parametric",
# verbose = TRUE
# )

```

```{r test-singularity, include = FALSE}

isSingular(final_model) #returns FALSE

```

<div>

```{r final-model-no-speaker, include = TRUE}
#| code-fold: true

final_model_no_speaker <- glmer(
  passive_variant ~ (subj_animacy + subj_gender + cluster + adversativity + subj_animacy:adversativity) + 
                    (1 | lemma),
  family = "binomial",
  data = passive_sample_clean,
  control = glmerControl(optimizer = "bobyqa")
)

```

</div>

```{r tab-model-no-speaker, echo = FALSE}
#| label: tbl-model-without-speaker
#| tbl-cap: Results of the Final Model without the `speaker` Random Effect

tab_model(final_model_no_speaker, show.aic = TRUE) 

```

As noted by Gries [-@gries_statistics_2021: 385], models often perform better on the data on which they were trained, but may not generalise to other samples if they adapt too closely to the structure and random variation of the training sample. This is called overfitting. According to Levshina [-@levshina_how_2015: 167], large differences between training and test performance (referred to as optimism) can indicate overfitting. In the summary of our bootstrap statistics for our final model in @tbl-bootstrap-summary, the differences between our original model and the means of the bootstrapped resamples can be seen in the bootstrap bias (`bootBias`). Note, however, that bootstrap validation does not provide information about how the model will perform on unseen data.

::: {#tbl-bootstrap-summary}
```{r bootstrap-summary, echo = FALSE}

# bootstrap_summary <- summary(bootstrap_final_model)
# 
# kbl(bootstrap_summary)
```

Bootstrap Statistics for the Final Model
:::

For most predictors, the bootstrap bias is very small (at or below 0.05), indicating that they are stable. The bias for `subj_animacyinanimate`, however, stands out slightly at -0.051, which may indicate some instability across bootstrap samples. This could also affect the interaction `subj_animacyinanimate:adversativitynon-adversative`, which is based on the same predictor. As can be seen in @fig-density-animacy, the distribution for `subj_animacyinanimate` deviates somewhat from a bell shape, centering around the bootstrap median -1.46 (`bootMed`), which supports the interpretation of reduced robustness.

```{r bootstrap-df, include = FALSE}

# bootstrap_df <- as.data.frame(bootstrap_final_model$t)
# 
# colnames(bootstrap_df) <- names(fixef(final_model))

```

::: {#fig-density-animacy}
```{r density-animacy, echo = FALSE}

# ggplot(bootstrap_df, aes(x = subj_animacyinanimate)) +
#  geom_density(fill = "lightblue") +
#  labs(x = "Coefficient Estimate", y = "Density") +
#  theme_minimal()

```

Bootstrapped Distribution of the `subj_animacyinanimate` Variable
:::

In contrast, the distribution for the significant main effect `clustercommunication` in @fig-density-cluster is approximately bell-shaped and centers closely around the original value of 1.80, which confirms its stability and positive effect. However, both verb cluster predictors, `clustercommunication` and `clusterhitting-touching`, have higher standard errors in the bootstrap statistics compared to the other fixed effects, which indicates greater uncertainty.

::: {#fig-density-cluster}
```{r density-cluster, echo = FALSE}

# ggplot(bootstrap_df, aes(x = clustercommunication)) +
# geom_density(fill = "lightblue") +
# labs(x = "Coefficient Estimate", y = "Density") +
# theme_minimal()

```

Bootstrapped Distribution of the `clustercommunication` Variable
:::

# Results {#sec-results}

```{r data-exploration-check, include=FALSE}

str(passive_sample_clean)

head(passive_sample_clean)

table(passive_sample_clean$passive_variant)

```

```{r freq-functions, include = FALSE}

# define a function to summarise the absolute and relative frequencies for each factor
summarise_factor_freq <- function(data, factor) {
  data |>
    count(passive_variant, {{factor}}) |>
    group_by(passive_variant) |>
    mutate(relative = round(proportions(n), 2)) |>
    pivot_wider(
      names_from = passive_variant,
      values_from = c(n, relative),
      values_fill = list(n = 0, relative = 0)
    ) |>
    ungroup() |> 
    dplyr::select({{factor}}, n_être, relative_être, `n_se faire`, `relative_se faire`)
}

# define a function to create separate dataframes for the individual factor frequency tables in preparation for row-binding
prep_row_bind <- function(data, factor) {
  data_bind <- data
  # rename first column to "level"
  colnames(data_bind)[1] <- "level"
  # add a "factor" column
  data_bind$factor <- factor
  return(data_bind)
}

# define a function to pull the relative frequencies of individual levels (for inline code)
count_level_relative <- function(data, factor_column, factor_level, relative_column) {
  data |>
    filter({{factor_column}} == factor_level) |>
    pull({{relative_column}})
}

```

```{r freq-animacy, include=FALSE}

freq_animacy <- summarise_factor_freq(passive_sample_clean, subj_animacy)

animate_etre_rel <- count_level_relative(freq_animacy, subj_animacy, "animate", relative_être)
animate_sefaire_rel <- count_level_relative(freq_animacy, subj_animacy, "animate", `relative_se faire`)
inaninmate_etre_rel <- count_level_relative(freq_animacy, subj_animacy, "inanimate", relative_être)
inanimate_sefaire_rel <- count_level_relative(freq_animacy, subj_animacy, "inanimate", `relative_se faire`)

```

```{r freq-gender, include = FALSE}

freq_gender <- summarise_factor_freq(passive_sample_clean, subj_gender)

table(passive_sample_clean$subj_gender) |> 
  sort(decreasing = TRUE)

gender_table <- table(passive_sample_clean$subj_gender, passive_sample_clean$passive_variant)

gender_prop <- prop.table(table(passive_sample_clean$subj_gender))
  
```

```{r freq-adversativity, include = FALSE}

freq_adversativity <- summarise_factor_freq(passive_sample_clean, adversativity)

table(passive_sample_clean$adversativity) |> 
  sort(decreasing = TRUE)

adversativity_table <- table(passive_sample_clean$adversativity, passive_sample_clean$passive_variant)

adversativity_prop <- prop.table(table(passive_sample_clean$adversativity))

adversative_etre_rel <- count_level_relative(freq_adversativity, adversativity, "adversative", relative_être)
adversative_sefaire_rel <- count_level_relative(freq_adversativity, adversativity, "adversative", `relative_se faire`)
nonadversative_etre_rel <- count_level_relative(freq_adversativity, adversativity, "non-adversative", relative_être)
nonadversative_sefaire_rel <- count_level_relative(freq_adversativity, adversativity, "non-adversative", `relative_se faire`)

```

```{r freq-cluster, include=FALSE}

freq_cluster <- summarise_factor_freq(passive_sample_clean, cluster)

neutral_etre_rel <- count_level_relative(freq_cluster, cluster, "neutral", relative_être)
neutral_sefaire_rel <- count_level_relative(freq_cluster, cluster, "neutral", `relative_se faire`)
communication_etre_rel <- count_level_relative(freq_cluster, cluster, "communication", relative_être)
communication_sefaire_rel <- count_level_relative(freq_cluster, cluster, "communication", `relative_se faire`)
hittingtouching_etre_rel <- count_level_relative(freq_cluster, cluster, "hitting-touching", relative_être)
hittingtouching_sefaire_rel <- count_level_relative(freq_cluster, cluster, "hitting-touching", `relative_se faire`)

```

```{r freq-other-factors, include = FALSE}

freq_complement <- summarise_factor_freq(passive_sample_clean, agent_pp)
freq_modal <- summarise_factor_freq(passive_sample_clean, modal)
freq_tam <- summarise_factor_freq(passive_sample_clean, tam)

```

```{r bind-freq, include = FALSE}

freq_animacy_bind <- prep_row_bind(freq_animacy, "Animacy")
freq_gender_bind <- prep_row_bind(freq_gender, "Gender")
freq_complement_bind <- prep_row_bind(freq_complement, "Complément d'agent")
freq_modal_bind <- prep_row_bind(freq_modal, "Modal Verb")
freq_adversativity_bind <- prep_row_bind(freq_adversativity, "Adversativity")
freq_cluster_bind <- prep_row_bind(freq_cluster, "Verb Cluster")
freq_tam_bind <- prep_row_bind(freq_tam, "TAM")

freq_table <- bind_rows(freq_animacy_bind, freq_gender_bind, freq_complement_bind, freq_modal_bind, freq_adversativity_bind, freq_cluster_bind, freq_tam_bind) |> 
  dplyr::select(factor, everything())

```

```{r coefficients-log-odds, include = FALSE}
final_model_coef <- coefficients(summary(final_model))[,"Estimate"]
final_model_coef_table <- tibble(
  Factor = names(final_model_coef),
  Estimate = round(unname(final_model_coef), 3)
)

```

```{r coefficients-log-odds-tab, include = FALSE}
#| label: tbl-log-odds
kbl(final_model_coef_table, caption = "Final model log odds")

```

```{r coefficients-odds, include=FALSE}
#convert from log odds to odds:
final_model_odds_ratios <- exp(final_model_coef)
#convert odds to probabilities
#exp(final_model_coef) / (1 + exp(final_model_coef))
```

This section is divided into three parts, corresponding to the overarching concepts: subject responsibility, adversativity, and aspectuality. Contrary to our original hypotheses, the only significant main effect was found for the semantic verb cluster, with communicative verbs favouring [se faire]{.smallcaps} over [être]{.smallcaps}. In addition, the model shows a significant interaction between subject animacy and adversativity. @tbl-frequencies summarises the absolute and relative frequencies for each factor level, with the relative frequencies calculated on the basis of the total number of occurrences of [être]{.smallcaps} or [se faire]{.smallcaps}, respectively.

Logistic regression produces log odds, which were automatically converted to odds ratios via exponentiation during model selection using the `tab()` function. Since odds ratios are more intuitive to interpret than log odds, all effect estimates discussed below are presented as odds ratios. As @tbl-final-model indicates, the intercept of the final model is `r round(final_model_odds_ratios[1], 2)`. This means that in the context of the reference levels (the subject is animate and masculine, the cluster of the verb is neutral, and the construction has an adversative sense), the chances of [se faire]{.smallcaps} being used are `r round(final_model_odds_ratios[1], 2)` times greater than that of [être]{.smallcaps}.

::: {#tbl-frequencies}
```{r tbl-freq, echo = FALSE}

kbl(freq_table, col.names = c("Factor", "Level", "Absolute", "Relative", "Absolute", "Relative")) |>  
  add_header_above(c(" " = 1, " " = 1, "être" = 2, "se faire" = 2)) |>  
  kable_styling(bootstrap_options = c("hover", "responsive", fixed_thead = TRUE)) |> 
  scroll_box(height = "400px")

```

Absolute and Relative Frequencies of the Factors Levels for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

## Subject Responsibility {#sec-results-subj-resp}

While several predictors were tested in relation to subject responsibility, only subject animacy showed a significant effect, though this is influenced by its interaction with adversativity. No significant effects were found for the *complément d’agent*, or modal verbs factors, and they are therefore not discussed further. While subject gender does not reach statistical significance, its distribution may suggest a potential tendency and is included here for exploratory purposes.

### Animacy {#sec-results-animacy}

As outlined in @sec-subj-resp, subject animacy is linked to subject responsibility or affectedness. @tbl-animacy illustrates the absolute and relative frequencies of the subject animacy levels for [être]{.smallcaps} and [se faire]{.smallcaps} in our data. With [être]{.smallcaps}, inanimate subjects occur more frequently than animate ones (`r   paste0(animate_etre_rel * 100, "%")` animate, `r   paste0(inaninmate_etre_rel * 100, "%")` inanimate), whereas [se faire]{.smallcaps} shows a strong preference for animate subjects (`r   paste0(animate_sefaire_rel * 100, "%")` animate, `r   paste0(inanimate_sefaire_rel * 100, "%")` inanimate). This distribution aligns with our expectations for [se faire]{.smallcaps}, but contradicts our hypothesis for [être]{.smallcaps}, which we also expected to occur more frequently with animate than with inanimate subjects. The findings for [se faire]{.smallcaps} replicate previous research [e.g. @novakova_construction_2009] and support the consensus that [se faire]{.smallcaps} strongly prefers animate subjects (@sec-subj-resp). If subject animacy is taken as an indicator of subject affectedness, the distribution suggests that subjects of [se faire]{.smallcaps} are likely being more affected by the event than those of [être.]{.smallcaps}

::: {#tbl-animacy}
```{r tbl-animacy, echo = FALSE}

kbl(freq_animacy, col.names = c("Animacy", "Absolute", "Relative", "Absolute", "Relative")) |>  
  add_header_above(c(" " = 1, "être" = 2, "se faire" = 2))

```

Absolute and Relative Frequencies of the Subject Animacy Levels for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

The main effect of subject animacy in the logistic regression model cannot be interpreted in isolation as it is distorted by the significant interaction with adversativity. The interaction of these two factors and their predictions for the choice between [se faire]{.smallcaps} and [être]{.smallcaps} are described in @sec-results-adversativity, while their implications for subject responsibility are further explored in @sec-discussion.

### Gender {#sec-results-gender}

In total, our sample comprises `r   sum(gender_table["masculine", ])` masculine subjects (`r   paste0(round(gender_prop["masculine"], 2) * 100, "%")`) and `r   sum(gender_table["feminine", ])` feminine subjects (`r   paste0(round(gender_prop["feminine"], 2) * 100, "%")`) across all [être]{.smallcaps} and [se faire]{.smallcaps} constructions. This distribution reflects an overall lower representation of feminine subjects compared to masculine subjects, particularly in [se faire]{.smallcaps}, as shown in @tbl-gender, in line with our hypothesis in @sec-subj-resp. While these frequencies suggest a bias towards masculine subjects, gender did not have a statistically significant effect in the logistic regression model and the frequencies should therefore be interpreted with caution.

::: {#tbl-gender}
```{r tbl-gender, echo=FALSE}

kbl(freq_gender, col.names = c("Gender", "Absolute", "Relative", "Absolute", "Relative")) |>
  add_header_above(c(" " = 1, "être" = 2, "se faire" = 2))

```

Absolute and Relative Frequencies of the Subject Gender Levels for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

The interaction between gender and animacy was also not significant. However, as can be seen in the barplots in @fig-gender-animacy, for both passive variants animate subjects are predominantly masculine, while inanimate subjects are more frequently feminine. [être]{.smallcaps} occurs most frequently with inanimate feminine subjects, while [se faire]{.smallcaps} occurs most frequently with animate masculine subjects. This distribution supports the hypothesis that [se faire]{.smallcaps} is more frequent with animate male subjects than with animate female subjects. For inanimate subjects, the difference in frequency between masculine and feminine gender is smaller compared to animate subjects, suggesting that the influence of gender on the choice of passive variant could be less strong for inanimate subjects than for animate ones. Given the sparse data for [se faire]{.smallcaps} with inanimate subjects and the lack of statistical significance, these patterns should be interpreted with caution.

::: {#fig-gender-animacy}
```{r gender-animacy-plot, echo=FALSE}

ggplot(passive_sample_clean, 
       aes(x = subj_animacy, fill = subj_gender)) +
  geom_bar(position = "dodge") +  
  facet_wrap(~ passive_variant) +  
  labs(x = "Subject Animacy",
       y = "Number of Observations", 
       fill = "Subject Gender") +
  scale_fill_manual(values = c("#8DA0CB", "#FC8D62")) +
  theme_bw()

```

Frequencies of [être]{.smallcaps} and [se faire]{.smallcaps} by Subject Gender and Subject Animacy
:::

## Adversativity {#sec-results-adversativity}

Overall, non-adversative constructions occured more frequently than adversative constructions in our data, with a ratio of `r   paste0(round(adversativity_prop["non-adversative"], 2) * 100, "%")` to `r   paste0(round(adversativity_prop["adversative"], 2) * 100, "%")`. However, adversative constructions were more common with [se faire]{.smallcaps} than with [être]{.smallcaps}: while only about one fifth (`r   paste0(adversative_etre_rel * 100, "%")`) of [être]{.smallcaps} were adversative, this was the case for two fifths (`r   paste0(adversative_sefaire_rel * 100, "%")`) of [se faire]{.smallcaps}. The mosaic plot in @fig-adversativity illustrates the proportional distribution of [être]{.smallcaps} and [se faire]{.smallcaps} across the adversativity levels.

::: {#fig-adversativity}
```{r adversativity-plot, echo = FALSE}

mosaicplot(passive_variant ~ adversativity,
           data = passive_sample_clean,
           main = "",
           xlab = "Passive Variant",
           ylab = "Adversativity",
           col = c("lightblue", "orange"))

```

Proportional Distribution of [être]{.smallcaps} and [se faire]{.smallcaps} across Adversativity Levels
:::

```{r interaction-values, include = FALSE}

coefficients <- final_model_summary$coefficients  

interaction_odds <- coefficients["subj_animacyinanimate:adversativitynon-adversative", 1]
interaction_odds_ratio <- exp(interaction_odds)

sderror_interaction <- coefficients["subj_animacyinanimate:adversativitynon-adversative", 2]

interaction_ci_min_log <- interaction_odds - 1.96 * sderror_interaction
interaction_ci_max_log <- interaction_odds + 1.96 * sderror_interaction
interaction_ci_min <- exp(interaction_ci_min_log)
interaction_ci_max <- exp(interaction_ci_max_log)

interaction_pvalue <- coefficients["subj_animacyinanimate:adversativitynon-adversative", 4]

```

The logistic regression model revealed a highly significant interaction between subject animacy and adversativity with an odds ratio of `r   round(interaction_odds_ratio, 2)` and a confidence interval of `r   paste0(round(interaction_ci_min, 2), "-", round(interaction_ci_max, 2))` ($p$-value = `r   format(round(interaction_pvalue, 5), scientific = FALSE)`, rounded to \<0.001 in the model output). This means that [se faire]{.smallcaps} is `r   round(interaction_odds_ratio, 2)` times as likely as [être]{.smallcaps} when the subject is inanimate and the context is non-adversative, so in these contexts, [se faire]{.smallcaps} is significantly less likely than [être]{.smallcaps}. Conversely, since the categories are binary, this also implies that [se faire]{.smallcaps} is much more likely in adversative contexts and with animate subjects than [être]{.smallcaps}.

```{r interaction-pred, include = FALSE}

interaction_prediction <- ggpredict(final_model, terms = c("subj_animacy", "adversativity"))

inanimate_adversative_prediction <- interaction_prediction |> 
  filter(x == "inanimate" & group == "adversative") |> 
  pull(`predicted`)

```

As @fig-animacy-adversativity-log shows, animate subjects strongly favour [se faire]{.smallcaps} over [être]{.smallcaps} regardless of adversativity, with the probability increasing only slightly in adversative contexts. With inanimate subjects, however, [se faire]{.smallcaps} is much more likely in adversative contexts (`r   paste0(round(inanimate_adversative_prediction * 100, 2), "%")`) than in non-adversative contexts, where it is almost never used. However, even in adversative contexts, [être]{.smallcaps} is predicted to be more common for inanimate subjects. For inanimate subjects, then, adversativity seems to be an important condition for [se faire]{.smallcaps} to be used at all.

::: {#fig-animacy-adversativity-log}
```{r interaction-model-plot, echo=FALSE}

ggpredict(final_model, c("subj_animacy", "adversativity")) |>  
  plot() +
  labs(
    title = "",
    x = "Subject Animacy",
    y = "Predicted Probability",
    color = "Adversativity"
  ) +
  scale_color_manual(values = c("lightblue", "orange")) +
  theme_bw()

```

Predicted Probabilities of [se faire]{.smallcaps} by Adversativity and Subject Animacy
:::

## Aspectuality: Verb Cluster {#sec-results-aspectuality}

As can be seen from the distribution of [être]{.smallcaps} and [se faire]{.smallcaps} across different verb clusters in @tbl-cluster, neutral verbs are the most frequent category in both passive variants, making up `r   paste0(neutral_etre_rel * 100, "%")` of all [être]{.smallcaps} constructions and `r   paste0(neutral_sefaire_rel * 100, "%")` of all [se faire]{.smallcaps} constructions. [se faire]{.smallcaps} is relatively more frequent with communication verbs (`r   paste0(communication_sefaire_rel * 100, "%")`) and hitting-touching verbs (`r   paste0(hittingtouching_sefaire_rel * 100, "%")`) compared to [être]{.smallcaps} (`r   paste0(communication_etre_rel * 100, "%")` and `r   paste0(hittingtouching_etre_rel * 100, "%")`).

::: {#tbl-cluster}
```{r tbl-cluster, echo=FALSE}

freq_cluster |> 
  kbl(format = "html", digits = 2, col.names = c("Verb Cluster", "Absolute", "Relative", "Absolute", "Relative")) |>  
  add_header_above(c(" " = 1, "être" = 2, "se faire" = 2))

```

Frequencies of [être]{.smallcaps} and [se faire]{.smallcaps} Across the Verb Clusters
:::

```{r communication-values, echo = FALSE}

communication_odds <- coefficients["clustercommunication", 1]
communication_odds_ratio <- exp(communication_odds)

sderror_communication <- coefficients["clustercommunication", 2]

ci_min_log <- communication_odds - 1.96 * sderror_communication
ci_max_log <- communication_odds + 1.96 * sderror_communication
ci_min <- exp(ci_min_log)
ci_max <- exp(ci_max_log)

communication_pvalue <- coefficients["clustercommunication", 4]

```

Surprisingly, and contrary to our initial hypothesis, the only significant main effect in our model is the semantic verb cluster, with communication verbs favouring [se faire]{.smallcaps} over [être]{.smallcaps}. We had predicted that hitting-touching verbs would favour [se faire]{.smallcaps}, given their dynamic and adversative nature, while communication verbs would be more compatible with [être]{.smallcaps}, as they typically feature an animate subject and frequently occur with object complements. However, in our model, the odds of [se faire]{.smallcaps} occurring with communication verbs are `r   round(communication_odds_ratio, 2)` times higher than with verbs in the "neutral" category (CI: `r   paste0(round(ci_min, 2), "-", round(ci_max, 2))`, $p$ = `r   round(communication_pvalue, 3)`). In contrast, hitting-touching verbs, which we expected to strongly favour [se faire]{.smallcaps}, did not show a significant effect.

```{r cluster-predict, echo = FALSE}

cluster_predicted <- ggpredict(final_model, "cluster")

neutral_predicted <- cluster_predicted |>
  filter(x == "neutral") |> 
  pull(predicted)

communication_predicted <- cluster_predicted |>
  filter(x == "communication") |> 
  pull(predicted)

hittingtouching_predicted <- cluster_predicted |>
  filter(x == "hitting-touching") |> 
  pull(predicted)

```

@fig-cluster-model-plot shows the predicted probabilities of [se faire]{.smallcaps} for each verb cluster, along with the confidence intervals. Note that these predictions are based on the context of animate subjects and adversative verbs, which are the reference levels in our model. For this reason, [se faire]{.smallcaps} is favoured over [être]{.smallcaps} for all three verb cluster categories. Communication verbs have the highest predicted probability of [se faire]{.smallcaps}, (`r   paste0(round(communication_predicted * 100, 2), "%")`). This supports our surprising finding that [se faire]{.smallcaps} is preferred for communication verbs, contrary to our hypothesis. Neutral verbs show a lower probability (`r   paste0(round(neutral_predicted * 100, 2), "%")`) suggesting a weaker preference for [se faire]{.smallcaps} over [être]{.smallcaps}. Hitting-touching verbs have a probability of (`r   paste0(round(hittingtouching_predicted * 100, 2), "%")`), but their confidence interval overlaps with that of neutral verbs, indicating no significant difference between these two categories. The wide confidence intervals, particularly for hitting-touching verbs, suggest a variability or uncertainty in the estimates, which aligns with the non-significant effect found for this category in our statistical model.

::: {#fig-cluster-model-plot}
```{r cluster-model-plot, echo = FALSE}

ggpredict(final_model, "cluster") |> 
  plot() +  labs(
    title = "",
    x = "Verb Cluster",
    y = "Predicted Probability"
  ) +
  theme_bw()
```

Predicted Probabilities of [se faire]{.smallcaps} Across the Verb Clusters
:::

Examining the most frequent communication verb lemmas in our data in @tbl-cluster-lemmas, we find that *insulter* 'to insult' (10 occurrences), *engueuler* 'to yell at' (7 occurrences), and *représenter* 'to represent' (5 occurrences) appear most often with [se faire]{.smallcaps}. Notably, *insulter* and *engueuler* are dynamic and adversative verbs, which may behave similarly to hitting-touching verbs, and typically occur with an animate undergoer - possibly explaining their preference for [se faire]{.smallcaps}. Conversely, the most frequent lemmas with [être]{.smallcaps} are *appeler* 'to call', *inscrire* 'to register/write', *rappeler* 'to remind', and also *représenter* 'to represent' (each with 4 occurences), which are annotated as non-adversative in our dataset, except for one occurence.

```{r cluster-lemmas, include=FALSE}

freq_communication <- passive_sample_clean |> 
  filter(cluster == "communication") |> 
  count(passive_variant, lemma) |> 
  group_by(passive_variant) |> 
  mutate(relative = round(proportions(n), 2)) |>
  pivot_wider(names_from = passive_variant, values_from = c(n, relative), values_fill = list(n = 0, Relative = 0)) |> 
  ungroup() |> 
    dplyr::select(lemma, n_être, relative_être, `n_se faire`, `relative_se faire`) |> 
  mutate(total = n_être + `n_se faire`) |> 
  arrange(desc(total))

```

::: {#tbl-cluster-lemmas}
```{r tbl-cluster-lemmas, echo=FALSE}

freq_communication |> 
  kbl(format = "html", digits = 2, col.names = c("Lemma", "Absolute", "Relative", "Absolute", "Relative", "Total")) |>  
  add_header_above(c(" " = 1, "être" = 2, "se faire" = 2, "")) |>  
  kable_styling(bootstrap_options = c("hover", "responsive", fixed_thead = TRUE)) |> 
  scroll_box(height = "400px")

```

Frequencies of Lemmas in the Communication Verb Cluster for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

Given this pattern, in line with our initial hypothesis, one possible explanation is that adversativity influences the choice of [se faire]{.smallcaps} within the communication verb cluster. However, there was no significant interaction between the verb cluster and the adversativity factor. As shown by the proportions of verb clusters across adversativity levels in @fig-interaction-adv-cluster, [se faire]{.smallcaps} is used more frequently with adversative communication verbs than with non-adversative communication verbs, while the opposite pattern can be observed for [être]{.smallcaps}. Note, however, that this trend is not strong enough to be statistically significant in the logistic regression.

::: {#fig-interaction-adv-cluster}
```{r interaction-adv-cluster, echo=FALSE}

ggplot(passive_sample_clean, 
       aes(x = adversativity, fill = cluster)) +
  geom_bar(position = "fill") +  
  facet_wrap(~ passive_variant) +  
  labs(x = "Adversativity",
       y = "Proportion", 
       fill = "Verb Cluster") +
  scale_fill_manual(values = c("lightblue", "#FC8D62", "#8DA0CB")) +
  theme_bw()

```

Proportions of the Verb Clusters by Adversativity for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

Similarly to the adversativity of the verb, the animacy of the subject could play a role in the distribution of [se faire]{.smallcaps} and [être]{.smallcaps} with communication verbs like *insulter* and *engueuler*, which typically occur with an animate undergoer. Again, a trend can be observed in the frequencies, but there is no significant interaction between the verb cluster and the animacy factors. @fig-interaction-anim-cluster shows the proportions of [être]{.smallcaps} and [se faire]{.smallcaps} across verb clusters and subject animacy. [être]{.smallcaps} is more frequent with communication verbs when the subject is inanimate, whereas [se faire]{.smallcaps} is more common with communication verbs when the subject is animate. As with the frequencies discussed in @sec-results-gender, these plots are included solely for exploratory interpretation, as neither the interaction between verb cluster and adversativity nor the interaction with animacy reached statistical significance.

::: {#fig-interaction-anim-cluster}
```{r interaction-anim-cluster, echo=FALSE}

ggplot(passive_sample_clean, 
       aes(x = subj_animacy, fill = cluster)) +
  geom_bar(position = "fill") +  
  facet_wrap(~ passive_variant) +  
  labs(x = "Subject Animacy",
       y = "Proportion", 
       fill = "Verb Cluster") +
  scale_fill_manual(values = c("lightblue", "#FC8D62", "#8DA0CB")) +
  theme_bw()

```

Proportions of the Verb Clusters by Subject Animacy for [être]{.smallcaps} and [se faire]{.smallcaps}
:::

# Discussion {#sec-discussion}

As the multifaceted and partly conflicting literature suggests, the passive voice is a complex and nuanced phenomenon. This entails some challenges for the systematic study of the French [être:se faire]{.smallcaps} alternation. Firstly, the variable context is necessarily broad due to [se faire]{.smallcaps}'s similarity to the structurally identical causative construction. As has been shown, many borderline cases can be interpreted as either causative or passive, and sometimes even both, depending on the specific context. Similarly, [être]{.smallcaps} constructions range between the stative passive and the progressive passive, with the former often being difficult to distinguish from adjectival constructions.

For our analysis, we had to make decisions regarding the inclusion and exclusion criteria of borderline cases. However, a different set of choices could have been equally justified and might have yielded different results. It could be valuable to examine the cases we excluded from our analysis to understand how they differ from genuine passive constructions. This would provide insight into the gray area where passive nuances are present, but no true alternation occurs, allowing for a closer examination of varying degrees of passivity.

With regard to our hypotheses, although few significant effects were found, the significant interaction of subject animacy and adversativity clearly shows that [se faire]{.smallcaps} predominantly appears with animate subjects and in adversative contexts. [se faire]{.smallcaps} is very likely to occur with animate subjects in both adversative and non-adversative contexts, while inanimate subjects clearly favour adversative contexts. This is in line with the vast previous evidence on [se faire]{.smallcaps}'s preference for animate subjects [@koenig_semantic_1992; @novakova_construction_2009: 111-112; @spang-hanssen_quelques_1967; @tasmowski-de_ryck_causatif_1987; @raineri_analyse_2010: 92]. However, our analysis challenges the claim by @tasmowski-de_ryck_causatif_1987 and @spang-hanssen_quelques_1967 that [se faire]{.smallcaps} strictly requires an animate subject: our data provide evidence that, while [se faire]{.smallcaps} has a high chance to occur with animate subjects, inanimate subjects can occur with [se faire]{.smallcaps} (`r   paste0(inanimate_sefaire_rel * 100, "%")`), particularly in adversative contexts.

Importantly, even a strong correlation of animate subjects, adversative verbs and [se faire]{.smallcaps} does not necessarily imply that the subject is responsible for the event and brings negative consequences upon themselves [@gaatone_desagreable_1983: 170; @kupferman_construction_1995: 59-60]. Rather, the increased likelihood of [se faire]{.smallcaps} with inanimate subjects in adversative contexts suggests that the construction emphasises subject affectedness more than [être]{.smallcaps}. Note, however, that inter-annotator agreement was lowest for adversativity, indicating some inaccuracy regarding this factor, although it was still considered "substantial". Overall, we consider these findings to support the view that [se faire]{.smallcaps} is preferred over canonical [être]{.smallcaps} when emphasising subject affectedness.

Although grammatical gender, like animacy, was tested as an aspect of subject responsibility, it did not show a statistically significant effect in our logistic regression model, nor an interaction with animacy. For this reason, we do not draw any conclusions regarding gender, but consider the analysis of its absolute frequencies to be exploratory and tentative: in both passive variants, animate subjects are predominantly masculine in our sample, while inanimate subjects are more frequently feminine. [se faire]{.smallcaps}, which is often associated with higher subject responsibility or affectedness, occurs most frequently with animate masculine subjects, whereas [être]{.smallcaps} is more common with inanimate feminine subjects. This is consistent with our hypothesis that [se faire]{.smallcaps} is more typical for male subjects, suggesting a bias towards masculine subjects, with masculine subjects dominating in contexts of higher agency. In contrast, feminine gender appears more frequently with inanimate referents, possibly reflecting stereotypical associations with passivity or reduced agency. However, for inanimate subjects the difference in frequency between male and female gender is smaller, which might indicate that gender plays a lesser role for inanimate subjects.

In the data annotation, we did not distinguish between human and non-human referents. However, in previous work on French by @da2022objectifying, where female referents were shown to appear less likely as subjects than male referents, this syntactic gender bias was significant only when the referents were human. It is therefore possible that our broader animate/inanimate categorisation contributed to the lack of a significant gender effect. In addition, the number of [se faire]{.smallcaps} constructions with inanimate subjects was scarce in our dataset, which could also have influenced the significance of the effect. It is also worth noting that particular subject NPs, such as *les personnes* 'the people/individuals', which is grammatically feminine but semantically gender-neutral and animate, are frequent in the data and may have influenced the interaction between gender and animacy. Such cases are worthy of further investigation in future work.

The other factors used to operationalise subject responsibility - the presence of a *complément d'agent* and the use of a modal verb - also yielded no significant effects. This could further support the interpretation that [se faire]{.smallcaps} is more strongly related to subject affectedness than subject responsibility. It is also worth mentioning that our annotation of modality was limited to modal verbs. Other constructions expressing subject involvement or volition, such as *avoir besoin de*, *arriver à*, *réussir à*, *envier de*, *avoir des difficultés pour*, *avoir la possibilité de*, and *avoir envie de*, were not considered. However, we realised during the annotation process that these expressions are remarkably frequent in the context of [se faire]{.smallcaps}. Therefore, they may be worth considering, to further investigate the relationship between subject responsibility and the use of [se faire]{.smallcaps}.

The findings regarding aspectuality in our logistic regression model are unexpected and contradict our initial hypothesis. Our model shows a significant effect for the semantic verb cluster, with communication verbs significantly favouring [se faire]{.smallcaps}, which contrasts with our expectation that [être]{.smallcaps} would be more likely in these contexts. We could not explain this result, as no significant interactions were found. The relationship between verb cluster and adversativity, as well as with subject animacy, may be more complex than initially anticipated, or our dataset may not have been optimally suited to detect a significant interaction effect.

Seeking an explanation within our theoretical framework, we examined the most frequent communication verb lemmas occurring with [se faire]{.smallcaps}. The most frequent lemmas (*insulter* and *engueuler*) carry a dynamic and adversative nuance and typically take an animate undergoer. These verbs may therefore behave similarly to hitting-touching verbs, emphasising subject affectedness. This could also be supported by the finding that [se faire]{.smallcaps} is more frequent with hitting-touching verbs than [être]{.smallcaps}, but this effect is also not statistically significant. Alternatively, other factors not considered in our analysis may account for these findings. Such factors could be register and medium: [se faire]{.smallcaps} is predominantly used in informal contexts and spoken language [@raineri_analyse_2010: 87, 105], whereas [être]{.smallcaps} is more frequently used in written than in spoken French [@da2022alternance: 76-77]. Verbs like *insulter* and *engueuler* could also be more typical of colloquial speech.

An alternative explanation for this result is the inherent lexical properties of verbs. In our model, the random intercept for `lemma` (τ₀₀ = 1.94) indicates a substantial contribution to the variance, suggesting that certain verbs are more likely to occur with [se faire]{.smallcaps} due to their lexical properties. This emphasises the important role of lexical effects for the likelihood of [se faire]{.smallcaps} and aligns with Bohmann et al.[-@bohmann_large-scale_2023: 48]'s findings that English [get]{.smallcaps} is largely lexically determined. We believe that the application of memory-based learning methods, such as k-nearest neighbour classifiers, could provide more insights into the lexical conditioning of [se faire]{.smallcaps}, as they capture lexical properties without requiring explicit rule-based abstractions [@theijssen2013choosing: 243-244].

In terms of aspectuality and TAM, our model did not reveal any significant effects. This contradicts previous observations, such as those by Bellec [-@bellec_construction_2014: 209], who notes that [être]{.smallcaps} is more likely to describe an outcome or endpoint of an event, while [se faire]{.smallcaps} is more likely to occur with verbs referring to progressive or ongoing actions. Based on this, we had hypothesised that progressive aspect would favour [se faire]{.smallcaps}. In our analysis, TAM was broadly grouped into two categories: "progressive" and "non-progressive". We chose this approach to facilitate our annotation and ensure inter-annotator agreement, especially since neither annotator was a native French speaker. However, this annotation may not have been fine-grained enough to capture subtle aspectual nuances. This may partly account for the lack of a significant TAM effect in our results.

<!-- Another aspectual factor not included in our annotation, but worth considering as a factor in future research, is telicity. According to Bellec [-@bellec_construction_2014: 209], telic verbs, which mark the endpoint of an action, appear more typically with [être]{.smallcaps}, while atelic verbs, which are are inherently ongoing (e.g. perception verbs), are more likely with [se faire]{.smallcaps} [@bellec_construction_2014: 209]. This factor could also overlap with the verb cluster factor. -->

One general limitation of our dataset is that it does not fully represent actual language use, where [être]{.smallcaps} is the unmarked default passive and far more frequent than [se faire]{.smallcaps}, which is a marked and rare variant. In the ParlaMint 2.1 corpus, [être]{.smallcaps} occurs 324,836 times, compared to just 1,254 instances of [se faire]{.smallcaps}. To prevent the model from being skewed, we created a balanced dataset based on the lemmas occurring with [se faire]{.smallcaps}. While this approach allowed us to focus on the factors influencing the passive alternation, it introduced an artificial ratio. It may also have biased the lexical distribution towards verbs that are particularly compatible with [se faire]{.smallcaps}. In addition, we excluded all occurrences with "unknown" gender to avoid potential bias and collinearity in the model. This removed 165 [se faire]{.smallcaps} and only 29 [être]{.smallcaps} constructions, thereby leading to some imbalance in the final dataset. This potentially limits the generalisability of our results.

Despite these limitations, the model diagnostics indicate a good overall fit. Although certain aspects, such as suboptimal fits for certain data, suggest potential areas for refinement, the visual inspection indicated a stable model that makes reliable predictions. The bootstrap validation supports our model's stability, but indicates that some results should be interpreted with caution. We retained the original model for interpretation, acknowledging the trade-off between stability and goodness of fit. In this context, we identified the simplified model (@tbl-model-without-speaker), which omits the speaker effect as a potential robust alternative. Most fixed effects show minimal bias, although the bias for subject animacy (animate) suggests that its effect may be slightly less reliable than the other estimates. The verb cluster predictors have higher standard errors, which indicates some uncertainty. Overall, the estimates appear meaningful, but further work would be required to refine the model. While bootstrap validation provides insight into model stability, the application of additional validation techniques would help identifying overfitting, thereby ensuring that the model performs well on unseen data. As a possible approach we suggest cross-validation, where the model is trained on several separate subsets of data and tested on the remaining data.

<!-- Dataset/sample: problem: data are extremely unbalanced (ratio in ParlaMint être vs. se faire: ) -\\\> making the ratio 50: 50 would distort the reality significantly but if the ratio is kept the model will almost always predict êtrestrategy possibilities: - build a corpus where être is presented more than se faire but not as extreme (approximation, e.g. 40% se faire - 60% etre) ("over/undersampling within limits") - assign more weight to se faire constructions by "weights=" variable of glm function - \@bohmann_large-scale_2023 kept the imbalance (ca. 3 Mio be passive vs. 25 000 get passive) - evaluate with precision & recall, not just accuracy (F1-score) -->

<!-- As some linguists (e.g., Martin, Novakova / ... as above?) argue, *se faire* + infinitive can sometimes have a 'passive effect' even without a direct 1:1 alternation with *être*. -->

<!-- -   implications for causativity debate? -->

# Conclusion {#sec-conclusion}

<!-- restate the problem, find novel way to circle back to the problem -->

When constructing a passive sentence, French native speakers usually choose intuitively between the two variants [être]{.smallcaps} and [se faire]{.smallcaps}. However, this choice is not random, but influenced by many linguistic factors. Based on previous research, we narrowed down these factors and constructed a mixed-effects binomial logistic regression model that predicts the occurrence of either [être]{.smallcaps} or [se faire]{.smallcaps}. In order to provide a theoretically stable basis for the model, we clustered the factors under the overarching concepts of subject responsibility, adversativity, and aspectuality. We hypothesised that [se faire]{.smallcaps} would be chosen more often with animate, male subjects, in adversative, inchoative contexts, and in combination with verbs from the semantic cluster of "hitting-touching". Since [se faire]{.smallcaps} is attested to occur more frequently in spoken than in written French, we chose to investigate its use and alternation in the context of parliamentary debates. Throughout this process, we aimed to make our study reproducible and transparent by preregistering it and documenting every step accurately.

During the in-depth investigation of the alternation, the complexity of the construction and the difficulty of our chosen task became apparent. We were faced with challenges of annotation, analysis, and interpretation, which we navigated collaboratively to the best of our ability and knowledge. Although the complexity of language necessarily exceeds what can be measured by the included factors and their interplay, our model produces noteworthy results.

<!-- summary -->

<!-- although the French [être:se faire]{.smallcaps} alternation presents considerable analytical challenges due to its complexity,-->

To summarise, the interaction between subject animacy and adversativity suggests that [se faire]{.smallcaps} emphasises subject affectedness more than [être]{.smallcaps}. The results also show that [se faire]{.smallcaps} is not used exclusively with animate subjects, challenging previous research. While [être]{.smallcaps} remains clearly preferred for inanimate subjects, adversative verbs increase the likelihood of inanimate subjects being used with [se faire]{.smallcaps}. Furthermore, a main effect of the semantic verb cluster on this alternation was found, with communication verbs surprisingly favouring [se faire]{.smallcaps}. Notably, the exact magnitude of these findings may be specific to the specialised genre of parliamentary debates. Nevertheless, since the effects of animacy and adversativity were so strong, we expect them to be applicable to a variety of spoken French. The random effect of the lemma shows the important role of lexical effects in this alternation. However, the limitations of our dataset, annotation scheme, and model design require further refinement, particularly with regard to animacy, gender, aspect, and lexical variation. Methodologically, our model proves to be robust, but additional validation techniques may improve the assessment of its generalisability.

<!-- discuss implications, key takeways; call for action for future research -->

<!-- Finish your paper on a forward-looking note by suggesting how you or other researchers might build on this topic in the future and address any limitations of the current paper. -->

Our analysis reveals interesting results that could pave the way for further investigation. For instance, it might be illuminating to model the lemma as a fixed effect rather than a random effect for better control and interpretability. In addition, a model trained on comparable written and spoken data would allow the inclusion of mode as a factor, thereby assessing the exact distribution of both passive variants and tracking how much more common [se faire]{.smallcaps} really is in spoken language. An approach outside of regression modelling may also be valuable: we suggest memory-based learning as a method that makes predictions of class membership (i.e. passive choice) based on basic features.

Overall, the corpus-based approach allowed us to meticulously model real-world spoken French and make substantiated inferences based on extensive statistical evidence. Therefore, we believe that this study offers a basis for future refined analyses of passive alternations in French and other languages. Our findings provide insight into the complex [être:se faire]{.smallcaps} alternation, and highlight the value of combining corpus-based methods with a theoretical framework when modelling linguistic alternations.

# References

::: {#refs}
:::

## Packages used

```{r package-versions, echo=FALSE}
sessionInfo()
```

## Package references

```{r generateBibliography, results="asis", echo=FALSE}

packages.bib <- sapply(1:length(loadedNamespaces()), function(i) toBibtex(citation(loadedNamespaces()[i])))

knitr::write_bib(c(.packages(), "knitr"), "packages.bib")

require("knitcitations")
cleanbib()
options("citation_format" = "pandoc")
read.bibtex(file = "packages.bib")
```

# Declaration of Collaboration {#sec-declaration-collab}

We, Gina Reinhard and Rose Hörsting, wrote this term paper in close collaboration. All decisions were made by mutual agreement. We both contributed equally to the planning, design, and analysis of this study. The following authors' statement is based on the [CRediT author statement](https://www.elsevier.com/researcher/author/policies-and-guidelines/credit-author-statement) and provides a detailed description of who was responsible for which parts and to what extent. The categories suggested by CRediT are adapted for our purposes and supplemented by references to the corresponding sections of this paper. Notably, many sections were true team efforts and cannot be clearly attributed to one person alone. In such cases, the contributions are listed for both authors.

**Gina Reinhard**:

-   Conceptualisation

-   Methodology

    -   Data Collection/ Curation: CQL for se faire, manual filtering of 1,254 se faire constructions, manual annotation of 648 passive constructions

    -   Reference levels

-   Validation

    -   Model Diagnostics: Bootstrap Validation

-   Formal analysis / Investigation

    -   Results: Overall frequencies, Subject Responsibility (Subject Animacy and Subject Gender), Adversativity, Aspectuality

-   Writing-Original Draft (including Code)

    -   @sec-etre, contributions to @sec-debate, @sec-subj-resp (Subject Animacy first and last paragraph, Gender, Complément d'agent, contributions to Modal Verb), contributions to @sec-adversativity, @sec-aspectuality (including verb cluster table in appendix)

    -   @sec-corpus, @sec-semi-automatic-approach, @sec-filtering-criteria, @sec-log-regression (Bootstrap Validation)

    -   @sec-results

    -   @sec-discussion

    -   contributions to @sec-conclusion

-   Writing - Review & Editing: throughout the whole paper

-   Visualisation

    -   @tbl-model-without-speaker, @tbl-bootstrap-summary, @tbl-frequencies, @tbl-animacy, @tbl-gender, @tbl-cluster, @tbl-cluster-lemmas

    -   (@fig-clusters), @fig-density-animacy, @fig-density-cluster, @fig-gender-animacy, contributions to @fig-adversativity, @fig-cluster-model-plot, @fig-interaction-anim-cluster, @fig-interaction-adv-cluster

**Rose Hörsting**:

-   Conceptualisation

-   Methodology

    -   Data Collection/ Curation: CQL for être, manual annotation of 648 passive constructions

    -   Reference Levels, Model Selection

-   Validation

    -   Model Diagnostics: Residuals, Concordance Index, Significance

-   Formal analysis, Investigation

    -   Results: Log odds & Intercept, Subject Animacy, Adversativity

-   Writing - Original Draft (including Code)

    -   @sec-introduction

    -   @sec-se-faire, @sec-factors: @sec-sefaire-context, @sec-subj-resp (Intro Subject Responsibility, Subject Animacy, Modal Verb), @sec-adversativity, contributions to Tense-Aspect-Mood (TAM) in @sec-aspectuality

    -   @sec-corpus, @sec-semi-automatic-approach, contributions to @sec-filtering-criteria, @sec-iaa, @sec-data-preprocessing, @sec-log-regression (except bootstrap validation)

    -   @sec-results-animacy, @sec-results-adversativity

    -   contributions to @sec-discussion (first three paragraphs & second to last paragraph)

    -   @sec-conclusion

-   Writing - Review & Editing: throughout the whole paper

-   Visualisation

    -   @tbl-kappa-values, @tbl-levels, @tbl-model-max-unknown, @tbl-max-model-collinearity, @tbl-model-max, @tbl-model-back-1, @tbl-final-model, @tbl-anova, contributions to @tbl-frequencies

    -   @fig-binned-residuals, @fig-dharma-dispersion, @fig-dharma-qq, @fig-adversativity, @fig-animacy-adversativity-log

For a full overview of the exact contributions of each author, please consult our Git repository: <https://github.com/ginareinhard/sefairePassiveAlternation>

# Appendix {.appendix}

## Verb Cluster Categorisation with Example Verbs {.appendix}

<div>

| Verb Clusters [@francois_presentation_2007, 4]                    | Grouped Verb Clusters (Our Study) | Example Verbs                                                                                                                                                                                                 |
|--------------------|--------------------|--------------------------------|
| giving, depriving (don, privation)                                | neutral                           | aider (*\[...\] une personne victime, à 57 ans, d'un accident entraînant l'apparition d'un handicap et qui se ferait aider pendant dix, quinze, voire vingt ans par son conjoint sans demander la PCH.*)      |
| entering, leaving (entrée, sortie)                                | neutral                           | engager (*\[...\] cela répond à votre attente à l'égard de fonctionnaires qui quitteraient l'administration pour se faire engager par ailleurs au moyen de contrats de forme différente.*)                    |
| hitting, touching (frapper, toucher)                              | hitting-touching                  | tuer (*\[...\] ces hommes que l'on avait tant de difficulté à intégrer furent mobilisés pour aller combattre et se faire tuer au nom de notre pays \[...\]*)                                                  |
| movement in place (mouvement sur place)                           | neutral                           | \-                                                                                                                                                                                                            |
| equipping, disarming (munir, démunir)                             | neutral                           | vacciner (*Il faut une prescription pour se faire vacciner contre la grippe en pharmacie \[...\]*)                                                                                                            |
| transformation, change (transformation, changement)               | neutral                           | remplacer (*\[...\] afin que le premier puisse se faire remplacer par le second \[...\]*)                                                                                                                     |
| realisation, preparation (réalisation, mise en état)              | neutral                           | photographier (*J'ajoute la diminution du nombre de points où l'on peut se faire photographier et déposer son dossier en vue de faire établir ou renouveler sa carte nationale d'identité ou son passeport*.) |
| physical states and behaviours (états physiques et comportements) | neutral                           | aider (*\[...\] afin de permettre aux enfants victimes de haine en ligne de se faire aider même si leurs parents n'en sont pas informés.*)                                                                    |
| locative (locatif)                                                | neutral                           | domicilier (*Beaucoup trichent en se faisant domicilier chez un parent \[...\]*)                                                                                                                              |
| union, reunion (union, réunion)                                   | neutral                           | \-                                                                                                                                                                                                            |
| psychological verbs (verbes psychologiques)                       | neutral                           | écouter (*Le problème, on l'a dit, est de savoir comment mettre vraiment cela en pratique lorsque l'on constate une difficulté à se faire écouter.*)                                                          |
| seizing, clasping, possessing (saisir, serrer, posséder)          | neutral                           | élire (*\[...\] avant que M. Hollande ne se fasse élire en affirmant qu'il ne signerait pas le TSCG en l'état.*)                                                                                              |
| communication                                                     | communication                     | accuser (*\[...\] il est devenu impossible de l'exprimer dans le débat public sans se faire accuser de haïr les policiers \[...\]*)                                                                           |
| auxiliary verbs                                                   | \-                                | \-                                                                                                                                                                                                            |

Verb Cluster Categorisation with Example Verbs Based on the Classes by @francois_presentation_2007 [4]

</div>

## Alternative models {.appendix}

```{r model-with-unknown, include = FALSE}
passive_glmer_unknown <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:cluster + subj_animacy:adversativity + cluster:adversativity) +
               (1 | lemma) + (1 | speaker),
    family = "binomial", 
    data = passive_sample_unknown,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(passive_glmer_unknown) #AIC: 731.2
tab_model(passive_glmer_unknown) # Marginal R2 / Conditional R2	0.462 / 0.668 // 0.704 / NA
vif(passive_glmer_unknown)
```

```{r, include = FALSE}
model_1_unknown <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + agent_pp + modal + tam + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:adversativity) + (1 | lemma),
    family = "binomial",
    data = passive_sample_unknown,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(model_1_unknown) # AIC: 738
tab_model(model_1_unknown) #Marginal R2 / Conditional R2	0.453 / 0.659
```

```{r final-model-unknown, include = FALSE}
final_model_unknown <- glmer(
    passive_variant ~ (subj_animacy + subj_gender + cluster + adversativity + subj_animacy:subj_gender + subj_animacy:adversativity) + (1|speaker) + (1 | lemma),
    family = "binomial", 
    data = passive_sample_unknown,
    control = glmerControl(optimizer = "bobyqa")
  )

summary(final_model_unknown) # AIC: 759.3
vif(final_model_unknown)
```

```{r final-model-unknown-tab, echo = FALSE}
#| tbl-cap: Optimised model with level "unknown"
tab_model(final_model_unknown)
```

```{r residuals-model-unknown, include = FALSE}
plot(residuals(final_model_unknown) ~ predict(final_model_unknown))

sim_res <- simulateResiduals(final_model_unknown)
binnedplot(predict(final_model_unknown), residuals(final_model_unknown))
plot(sim_res)
testDispersion(sim_res)
```
