---
title: "Term paper"
subtitle: ""
author: "Rose Hörsting & Gina Reinhard"
date: "2025-03-15"
engine: knitr
bibliography: references.bib
format: 
  html:
    df-print: default
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 8
    number-sections: true
    embed-resources: true
link-citations: true
csl: "apa_7th.csl"
number-sections: true
---

```{=html}
<style>
p {
  text-align: justify;
}
</style>
```

```{r}
#| include: false
#install.packages("glossr")

library(glossr)
use_glossr(styling = list(
  source = "b",
  first = "",
  font_family = "Cambria"
))
```

# Introduction

-   rare construction: Only 5% of verbs in the French Treebank (FTB) are in the passive voice ( @da_cunha_lalternance_2020) vs. 9% in English for the same text type (@roland2007frequency).

# The French passive alternation of *être* and *se faire*

## Canonical passive with *être*

### Structure and meaning

-   Function of the passive construction (semantic focus):

-   Diathesis: Concept of the passive (passif long) as an inversion of the order/ syntactic function of the arguments (the subject argument and the object argument) in relation to the active construction. The object of the active "becomes" the subject in the passive. (Per default, in French the subject is coded at position 1 in the sentence)

-   The omission of the agent argument, called the *complément d'agent* (*par* or *de* phrase), is also possible. In this case, the construction is referred to as *passif court*, whereas the *passif long* explicitly includes the complément d'agent:

    *Language examples from our corpus*

    Approx. a quarter of all passive constructions in the French Treebank (FTB) are realised as passif long (@da_cunha_lalternance_2020), while in spoken French, the proportion is about 10% (source Hamma).

-   evtl.: passive vs. *construction moyenne en se* (*La porte a été ouverte* vs. *la porte s'est ouverte*) -\> In the passif court, the agent argument is still implicit (Quellen)

## *se faire* passive

```{=html}
<!-- ### Historical background

@bellec_grammaticalisation_2015 and @bellec_quand_2014 date the se faire + INF passive construction (PSF) to the middle of the 20th century and call it a "modern" construction that has since experienced a rapid increase in popularity. She describes a process of grammaticalisation and reanalysis for the PSF during/through which “se” has lost its reflexive meaning and “faire” has become an auxiliary. In merging, they form a “semi-auxiliary” and obtain a passive interpretation, while the causative meaning is lost (désémantisation). However, @bellec_construction_2014 claims that the causative reading did not disappear but instead coexists with the passive reading (debated, see below).

-   In contrast, @raineri_analyse_2010 [280-305] based on corpus data: sees origin much earlier, exists since origin of Middle French (14-17th century); NOT derived from causative reading, increased productivity from 19th century onward -->
```

### Structure and meaning

The construction consists of the pronoun "se", the conjugated form of the verb "faire" and a verb in the infinitive. According to @labelle_anticausativizing_2013 [235] the construction is "formally causative, but semantically passive". Similar to the être passive, the subject of the passive sentence is the experiencer of the action. As with être, the direct object of the active sentence becomes the subject of the passive sentence. *Language example: active sentence with two (animate) arguments, se faire passive sentence, evtl. etre passive sentence*

Unlike the être passive, however, the diathesis is not restricted to the direct object: The PSF has "greater syntactic flexibility" (@raineri_analyse_2010 [93-95] and can also be formed with ditransitive verbs, where the indirect object becomes the subject: *Language example: active ditransitive sentence, se faire passiv.* *les intéressés peuvent se faire expliquer tout ce qui touche aux retraites*

<!-- Another possibility are possessor constructions where the subject of the passive is only part of the object in the active sentence (Le coiffeur coupe les cheveux de Pierre. -> Pierre s'est fait couper les cheveux). However, the passive status of these constructions is controversial, as discussed below/ in the next section. Some could also be understood as causatives if the subject both initiates and undergoes the action. (Pierre made the hairdresser cut his hair) (add source)) -->

The peculiarity of the construction has led scholars to wonder: "how can a passive-like construction arise from a combination of a causative and a reflexive marker, where no marking of the passive function is apparent?" (@koenig_semantic_1992 [145]).

@labelle_anticausativizing_2013 [23] suggests that the passive meaning is obtained through a syntactic mechanism that involves binding the empty NP-object position of the embedded verb to the subject through the reflexive particle "se".

<!-- In the se faire passive, the embedded verb (i.e., the infinitive after faire) has an empty NP-object position since only the subject is realised. The reflexive clitic "se" binds the empty object position of the embedded verb. Therefore, "se" connects the (missing) object-NP and the surface subject (coindexation). In contrast to the être passive, the object does not become the new subject but is connected to the subject by "se". The subject is interpreted as affected because it is linked to the missing object of the embedded verb (which would be affected in the active sentence but is absent on the surface in se faire). For this to work, the embedded verb needs an agent as a subject (because otherwise, the missing object would not be affected). -->

Sentences including this construction can have either/both a passive and/or a causative sense:

*Language examples: passive se faire; causative se faire (same construction, different meaning)* @koenig_semantic_1992\[143\]

Crucially, passive se faire and causative se faire are structurally identical, differing only in their semantics. As @labelle_anticausativizing_2013\[236\] explains, the ambiguity is resolved by "the lexical items involved". In the causative reading, "se" is understood as reflexive and "faire" as causative. In contrast, if interpreted as a passive, the meanings of "se" and "faire" are not transparent. Thus, the subject referent does not necessarily cause the event and is not necessarily responsible (@koenig_semantic_1992\[143\]: "any notion of causation is absent"). Sometimes both readings are possible and the ambiguity cannot be resolved (@koenig_semantic_1992\[144\]), for example when cognition and perception verbs are involved (@bellec_construction_2014\[206, 207\]).

<!-- @kupferman_construction_1995 reasons that the PSF is a real, fully adequate passive construction, equivalent to the être passive. Others have different opinions... -->

### Debate of the construction's passive-status (vs. causative) {#sec-debate}

The above-described stance taken by @bellec_construction_2014, @labelle_anticausativizing_2013, @koenig_semantic_1992 and @kupferman_construction_1995 that the PSF is a full-fledged passive construction is by no means undisputed. The opposing view - that the construction is a type of causative - is also widely represented in the literature.

It is most prominently presented by @tasmowski-de_ryck_causatif_1987 and @gaatone_desagreable_1983, and was more recently presented by @sveberg_se_2012. In their view, passive meaning emerges from verb semantics and pragmatics, not from the PSF construction itself. As @raineri_analyse_2010 [89] describes, there is also the possibility of a multitude of readings (@kokutani_sur_2005) and of a continuum between causative and passive (Hobak Haff 2000), later accepted by @raineri_analyse_2010 [154-166].

<!-- As @labelle_anticausativizing_2013 [15] explains: -->

@gaatone_desagreable_1983 as well as @tasmowski-de_ryck_causatif_1987 centre their argumentation that the se faire passive is a weak reading of the causative around the notion that the subject of the se faire construction "is always, even unconsciously, responsible for the occurrence of the embedded event."

On the other hand, @kupferman_construction_1995 sees the causative and passive readings as homonyms with fundamentally different semantics and therefore assigns an inherent passive status to the se faire construction. This agrees with @bellec_construction_2014, according to whom the se faire passive has been grammaticalised into a semi-auxiliary with non-transparent components and there is thus no reason for a causative reading.

@raineri_analyse_2010 \[356-357\] concludes that despite its productivity as a passive construction, the PSF is problematic because it neither fits into two homonyms nor functions as a single weak causative. @bellec_construction_2014 [207] defines the following three criteria for a causative reading, which distinguish the causative from the passive interpretation:

-   In the causative, the subject referent is agentive and has the role of both causer and patient. The subject actively initiates what happens to them. In the passive, the subject referent has only the role of the patient and is affected by the event.
-   In the causative reading, the particle “se” has a reflexive meaning whereas in the passive reading, “se” is not reflexive.
-   In the causative reading, “faire” has transparent causative meaning. This causative meaning is lost in the passive reading, where “faire” has no inherent meaning. <!-- add: how will we distinguish them / have we distinguished them in our study -->

This debate has been central to the investigation of the se faire + INF constructions and cannot be left aside. However, even if the status of the PSF has been the subject of disagreement and discussion, the literature clearly shows that there is a passive reading associated with this construction. This is independent of whether the passive sense emerges from within the construction or from the components it occurs with, or whether it has an underlying causative meaning.

By comparing PSF and PE constructions, we may be able to shed light on this debate from a corpuslinguistic perspective by examining which factors contribute to the choice between the two passives. <!-- (-> exclude causative readings?). Can our findings contribute to the debate of whether se faire deserves passive-status? If not: no need to spell out the debate in detail. -->

## *être* vs. *se faire*: Competition or complementation? - Defining the variable context

Both se faire + Vinf and être + Vé are available for the active diathesis [@novakova_construction_2009 [108]. However, as @novakova_construction_2009 [108] notes, the constructions differ in semantic nuances: "Elles ont un sens proche mais pas identique." Thus, they are not semantically identical but rather potentially carry different meanings that may be more or less pronounced depending on the context: Previous research suggests that se faire passive is more likely to have a causative undertone while the être passive is more neutral - possibility of a continuum/ passive gradient. However, this is what makes the analysis interesting and our research worthwhile - what are the exact factors that influence the choice between the the two variants?

This introduces certain difficulties: The passive/ causative continuum discussed in @sec-debate makes the definition of the variable context particularly challenging as some constructions may be more passive than others - where to draw the line?

@bohmann_large-scale_2023 [35] faces similar challenges when defining the variable context of the be and get passive: "passive gradient" exist where both GET and BE may be used, but where the choice between the two entails a semantic difference", "definition of the variable context is not just a problem of automation, but of linguistic interpretation" \[36\]. Follow

Since our aim is to investigate the passive alternation of être and se faire, it is important to ensure that we only compare the être passive with real passive se faire constructions, and not with causative constructions (@bohmann_large-scale_2023 [35]: "all extracted forms should be genuine instances of the variable under investigation").

We follow @raineri_analyse_2010 \[165\] in including only se faire + Inf constructions that have a natural paraphrase with the être passive in order to include only constructions that directly contrast with être. Following @bohmann_large-scale_2023 [35] who claim that "only cases where both variants can be used interchangeably should be considered", we aim to exclude constructions that do not allow for a paraphrase with the other.

Therefore, passive constructions of se faire and être are not always interchangeable: As described above, the canonical passive with être requires a direct object in the active equivalent which becomes the new subject, @koenig_semantic_1992 \[144\], @washio_when_1993, @labelle_anticausativizing_2013 [16]. Thus, for reflexive and ditransitive verbs, the être passive is not available and does not compete with se faire. These constructions will be excluded and only transitive verbs will be studied.

-   se faire passive has evolved to be used in situations without a causer (passive reading) @bellec_construction_2014
-   figurative, "besoin d’expressivité and extravagance" instead of "used up" canonical passive @bellec_construction_2014 \[206\]; part of lexicalised expressions @novakova_construction_2009 [117]/ idioms @raineri_analyse_2010 [165]
-   while some scholars speak of semantic differences between the canonical être passive and the se faire passive, @kupferman_construction_1995 sees them as semantically synonymous -\> therefore: competition
-   how much is it used today (according to authors/ background literature) and in which settings
    -   se faire less formal, more figurative and emotional, used to build the passive with indirect object as the agent & to express passivity in dynamic/ ongoing processes, more spoken
    -   etre passive more “neutral”, more formal, more written
-   être: exclude adjectives and some participles that do not form the passive (similar to @bohmann_large-scale_2023 \[36\])/ "instances that allow for gradation" @hundt_be-_2024 [92]

### Maybe: Comparison to similar alternations in other languages

-   English @bohmann_large-scale_2023, East Asian languages @washio_when_1993 and @guan_multifactorial_2024 , Swedish @cuypere_corpus-based_2014
-   se faire passive "typologically marked", comparison to other Romance languages @koenig_semantic_1992 \[145\]; similar constructions exist in other Romance languages but passive reading only allowed in French @labelle_anticausativizing_2013 [236]

# Data and Methods

## Data Collection

### Corpus

In our investigation of spoken French language, we aimed to select a corpus that represents authentic French spoken language. After thorough consideration and exploration of our options, we decided for the French ParlaMint corpus [@11356/1864]. ParlaMint consists of French parliamentary debates from 2017 to 2020 and contains 38,702,395 tokens (32,214,147 words). We chose this corpus because it is well-documented, POS-tagged and contains rich meta data.

We recognize that the genre of parliamentary debates is quite specialised and our results therefore cannot be generalized to all French spoken language. However, the language of parliamentary debates is a fascinating genre in itself (studied before: references) and is expected to reflect a wide range of topics, thus providing a cross-section of language varieties as they occur in French speech.

We extracted the concordances from the French ParlaMint corpus [@11356/1864] using a CQL expression SketchEngine. To extract the data we chose a semi-automatic approach that included manual filtering.

-   We selected an equal number of occurrences of *être* passive and *se faire* passive.
    -   We made a list of the lemmas in the se faire constructions → used these lemmas to filter the être constructions

    -   Exclusion of causative reading of se faire constructions: the same verbs have been used with être passive in the same corpus (find/extract the constructions) → depends on how many constructions we have left
-   We have not taken into account impersonal passives (*example from corpus*)
-   être passive: In their active/passive alternation study @da_cunha_lalternance_2020 only chose passive sentences in the compound past tense "NP a été VP" in order to favour verbal passives over adjectival passives (*La porte est ouverte*). However, we chose to include the present tense/all tenses. This decision was based on previous literature suggesting that the present tense favors *se faire* passive over *être* passive, making tense an important factor in our study.
-   In line with @da_cunha_lalternance_2020 we chose complément d'agents introduced by the preposition *par*, avoiding additional variation caused by *de*.

```{=html}
<!-- problem: data are extremely unbalanced (ratio in ParlaMint être vs. se faire: ) -\> making the ratio 50: 50 would distort the reality significantly but if the ratio is kept the model will almost always predict être
strategy possibilities:
    -   build a corpus where être is presented more than se faire but not as extreme (approximation, e.g. 40% se faire - 60% etre) ("over/undersampling within limits")
    -   assign more weight to se faire constructions by "weights=" variable of glm function
    -   @bohmann_large-scale_2023 kept the imbalance (ca. 3 Mio be passive vs. 25 000 get passive)
    -   evaluate with precision & recall, not just accuracy (F1-score) -->
```

### Semi-automated approach

**se faire passive**

-   extract se faire passive constructions with CQL expression
-   manually: filter false positives (criteria: )
-   extract lemmas → save in list
-   annotate true positives for features

**être passive**

-   extract être passive constructions with CQL expression
-   In Sketch Engine: Import Word List → lemma list extracted from se faire constructions (for comparability of the constructions, making sure that differences between être and se faire are not due to individual differences of the verb) → Filter out only verb lemmas of se faire from concordance lines → download

**determine sample by setting a seed**

-   seed determined by future lottery numbers (8.3.2025?, <https://www.lotto.de/lotto-6aus49/lottozahlen>) → function set.seed( ) in R
-   take a random sample from the concordance lines that is equal in size to the manually filtered amount of se faire constructions
-   final sample: same amount of constructions for être and se faire, same lemmas, seed set
-   Our sample therefore comprises XXX sentences/constructions: XX se faire and XX être passive, ... (short summary)

### Borderline cases/ Manual filtering exclusion criteria

As was to be expected given the extensive debate on PSF's passive status (@sec-debate) and the challenges reported by @bohmann_large-scale_2023 in their investigation of the English be/get passive alternation, we came across a significant number of "unclear" examples when manually filtering se faire constructions.

Our main criterion was to include only constructions that could be paraphrased with the être passive, i.e. that were grammatical and similar in meaning. However, our somewhat "fuzzy" definition [@bohmann_large-scale_2023] posed some challenges: When is something similar in meaning and still acceptable, and when is the meaning too causative? We aimed to exclude purely causative constructions because they do not alternate with the être passive, but we have found many examples of se faire constructions cases where a binary passive/causative distinction is not possible, for example because since the subject is both the undergoer and the causer.

In general, we have chosen a wide approach and have opted to include as many constructions as possible: This means that even if the causative reading seems natural (because the subject triggers the event), but there is a passive understanding because the subject is undergoing an action, it has been included (going against the first criterion by @bellec_construction_2014). Examples of such constructions are "se faire vacciner" *to (go) be/get vaccinated*, "se faire élire" *to be/get elected; to stand for election.*

We are also aware that some verbs can be causative in one context and passive in another. But for practicality, we decided to include or exclude them on the basis of the verb lemma, rather than individually. For traceability, below are some high frequency ambiguous cases and our argumentation/ reasoning for our decision to include/ exclude them.

In addition to erroneously tagged constructions that are not real se faire + Vinf expressions (e.g.), we excluded expressions that do not not allow for a paraphrase with the être passive, either because they are ungrammatical or fundamentally different in meaning. These include:

-   Fixed idiomatic expressions with a figurative meaning: "se faire passer pour" *to impersonate sb.*, "se faire avoir" *to be fooled*, "se faire mousser" *to boast/ to brag, "se faire taper sur les doigts" lit. to get a slap on the fingers, to be scolded* ; example: "Dites donc, vous savez vous faire mousser!" *My, you know how to make a fuss!/ You know how to blow your own trumpet!*

-   Expressions that cannot be converted into a correct active sentence (i.e., the direct object of the active sentence is not the subject of the passive sentence). (ditransitive, intransitive verbs ?), example: "mais se faire expliquer des elements de justice n'est pas la même chose que d'acheter en ligne des produits bio ou une paire de chaussures." *but getting an explanation of legal matters is not the same as buying organic products or a pair of shoes online*

-   Expressions containing verbs classified as "idiomes lexicalisés" (lexical-ized idioms) according to Raineri (2010: 163). These verbs neither carry a true causative nor passive meaning [^1]: entendre, sentir, attendre, désirer (Raineri 2010: 164); example: "Partout le mécontentement populaire se faire entendre sans qu'' on ait besoin de souffler sur les braises." *Everywhere, popular discontent is making itself heard without the need to fan the flames.*

-   (Expressions with a transparent causative meaning - maybe not necessary if covered by the other criteria)

-   (Expressions where an être passive has a very different meaning)

[^1]: These verbs mainly occur with inanimate subjects (effects, consequences, discontent). @raineri_analyse_2010 [163-164] proposes that the causative meaning is not available for these verbs since the subject is not causing the event. There is also no passive meaning since they cannot cannot occur with a complément d’agent and do not have a paraphrase with être.

## Logistic Regression

Variables, Hypotheses, Preference constraints

Several corpus studies on voice choice in French have shown that the choice between active and passive constructions (*être* passive) is not purely syntactic, but that semantic and pragmatic factors also play a crucial role in determining whether a passive construction is favoured.

-   Subjects in passive sentences are preferably animate, definite, and pronominal. In the study by @da_cunha_lalternance_2020, animate, definite, and pronominal arguments that function as the objects in active sentences were more frequently found in the passive (and vice versa). -\> prominence / informational structure as significant factor. (And: constituent length, shorter constituents are preferred in subject position and favour the passive voice)

### Subject responsibility

As indicated above, subject responsibility in the se faire passive is central to the debate about the construction's passive-status (see above; @sveberg_se_2012, @bellec_construction_2014 [212-213]) <!-- whether the passive reading of se faire is a type of the causative reading or its own instance --> <!-- As indicated above, the question whether the referent of the subject in the se faire construction holds any responsibility is central in the debate of the construction's passive-status. However, there is general consensus that
the PSF construction has a preference to mark the subject as an active participant (although according to @raineri_analyse_2010 [98] is is not "universally accepted" as some authors also name counterexamples @kupferman_construction_1995 [59].)
In @raineri_analyse_2010 [95]'s view, the idea that the PSF expresses subject responsibility is the "most common hypothesis" ("l'hypothèse la plus répandue") in the literature. -->

In line with their argument that the passive reading of se faire + INF is a kind of causative, @gaatone_desagreable_1983 and @tasmowski-de_ryck_causatif_1987 state that the subject referent of the se faire passive is always a causer and therefore always responsible to some degree. This is contrasted to the être passive's subject referent, which is see as more neutral. @gaatone_desagreable_1983 [164] distinguishes between an "active patient" in the se faire construction and a "passive patient" in the être passive (@raineri_analyse_2010 [96]).

In contrast, @kupferman_construction_1995 [59] gives PSF examples where the subject cannot be interpreted as responsible, one of which is repeated here:

```{r gloss}
#| echo: false
gloss_non_resp <- as_gloss(
  "elle sanglotait parce que son tout jeune bébé venait de se faire enlever les amygdales.",
translation = "She sobbed because her little baby got operated on for tonsilitis.",
source = "Kupferman (1995: 59)"
)
gloss_non_resp
```

According to him, subject responsibility emerges from the context and not from the PSF, which is inherently passive [98, 99]. Further support for this idea comes from @labelle_anticausativizing_2013 and @washio_when_1993. For @martin_construction_2002, subject dynamism, but not subject intentionality, is a necessary condition in the PSF (@raineri_analyse_2010 [90]).

Thus, although the requirement of subject responsibility in the PSF is not "universally accepted" (@raineri_analyse_2010 [98], "une catégorie hétérogène complexe" (@raineri_analyse_2010 [249]), the analyses show at least some consensus in that the PSF construction prefers to mark the subject as connected to the event (@raineri_analyse_2010), either through affectedness, responsibility or intentionality. This tendency does not exist for the être passive (source).

**Animacy**

Corpus studies have shown that animacy significantly influences voice choice in French. The animacy of the passive subject (not agent) has been shown to increase the likelihood of using the passive voice (*être* passive) over the active voice. @da_cunha_lalternance_2020 (+2022) -\> more explanation as to why

As implied in the debate on subject responsibility, subjects of se faire passive constructions are predominantly active participants in the described action. Since inanimate subjects are rarely affected and even less often responsible, PSF is common with animate subjects. While some scholars claim that only animate subjects are possible with the PSF, others provide counterexamples (@labelle_anticausativizing_2013, @kokutani_sur_2005, @novakova_construction_2009 \[4\], @sveberg_se_2012) (@raineri_analyse_2010 [92]).

However, these examples appear to be exceptions as there is both corpus (@novakova_construction_2009 [111]) and experimental (@koenig_semantic_1992 \[151, 153\]) evidence that subject animacy and PSF are positively correlated: In @novakova_construction_2009 [111]'s corpus study, less than 3% of the se faire sentences had an inanimate subjects. In a survey by @koenig_semantic_1992, French native speakers rated the acceptability of PSF with animate subjects significantly higher than with inanimate subjects. Inanimate subjecta were only allowed by some speakers when the subject "affects a relevant animate entity" (@koenig_semantic_1992 [151, 153]. The preponderance of animate subjects for the se faire passive seems to be one of the few aspects on which scholars agree (@raineri_analyse_2010 [92]).

We tag animacy as binary "inanimate"/ "animate" although we recognize that animacy has been prominently considered as a continuum in the literature.

**Hypothesis**: Both *se faire* and *être* passive favour animate subjects over inanimate subjects but subject animacy is a stronger predictor of the PSF than the être passive.

**Interaction**: @raineri_analyse_2010 [94] suspects that the PSF's animacy (almost-) constraint restricts verb selection and provides an explanation for the construction's incompatibility with certain verbs (i.e. not all verbs can occur with animate subjects).

<!-- The preference of the se faire passive of animate over inanimate subjects is parallel to the English get passive (@hundt_be-_2024), although @raineri_analyse_2010 [92] finds that the get passive allows more inanimate subjects than the se faire passive. According to @bohmann_large-scale_2023 [43], the get passive has become more likely with inanimate subjects over time which may signify a lessening of subject responsibility. @raineri_analyse_2010 brings up the possibility that a similar process of grammaticalisation may be happening in French. (not sure if relevant)-->

**Agent PP** (par-phrase, Complément d'agent)

The par-phrase introduces an external agent into the passive construction: It denotes the causer/ instigator of the action and is often derived through diathesis from the subject of the active equivalent.

We expect the presence or absence of the par-phrase to influence passive choice. The par-phrase introduces an external agent who is likely be the driving force behind the action described: @kupferman_construction_1995 postulates an "agentivity requirement" for agents denoted by the par-phrase, attributing a strong responsibility to the component. @hundt_be-_2024 and @bohmann_large-scale_2023 found that in English, the Agent PP reduces subject responsibility by introducing an agentive/ responsible entity.

**Hypothesis**: We hypothesize that the presence of an agent PP ("Complément d'agent") prefers the canonical être passive over the se faire passive (because it assigns responsibility to an external agent that is not the subject which is unusual for PSF, where the subject is usually somewhat agentive.)

### Aspectuality

Regarding the factor of aspectuality, @keenan_passive_2007 [340] cites one of Greenberg's universals: "If a language has two or more basic passives they are likely to differ semantically with respect to the aspect ranges they cover." Assuming that the PSF is a "basic passive", this suggests that the être and se faire passives will demonstrate different aspect preferences. Aspectuality is reflected in the verb, either in semantics ("Aktionsart") or grammatical marking (Tense, Aspect, Mode) [^2] (source missing!).

[^2]: The general consensus in the literature is that aspect marks progressivity grammatically and Aktionsart semantically/ lexically. However, this clear distinction between aspect and Aktionart has been doubted, and some theories see Aktionsart und Aspect as two instantiations of Aspectuality (source missing!).

-   **Verb semantics (Aktionsart), Clusters** (@bohmann_large-scale_2023 [51-52]: 21 levels):

    Multiple hypotheses are found in the literature concerning the compatibility of certain verbs with the PSF. For example, @kupferman_construction_1995 \[67\] and @raineri_analyse_2010 [90] state that se faire passive works well with dynamic verbs but not with stative verbs. (but exceptions provided by @novakova_construction_2009); Martin 2002, @raineri_analyse_2010 [100-103].

-   volitional verbs fit well with PSF because they involve an action that leads to a mental state change while non-volitional verbs do not

-   Telicity seems to be another important distinction: According to @bellec_construction_2014 \[209\], the être passive is likely to describe a result or mark an endpoint of an action, while the se faire passive is used for actions in progress and verbs that are in their nature ongoing (e.g. perception verbs, @bellec_construction_2014 [209]). This could be reflected in the telicity of the verb, with telic verbs preferring the être passive and atelic verbs favouring the se faire passive.

-   lemma clustering method, not too many different levels because not each level will have enough data points

-   **TAM (Tense, Aspect, Mode)**

    -   **Tense:** Descriptions of future events are more likely to focus on the start and process of an event/ action than on its endpoint (@bellec_construction_2014 [210]) Therefore, it seems likely that future tense will favour the se faire passive over the être passive. Regarding past tense, the être passive will probably be more popular than se faire for constructions in the passé composé (passé composé also used with se faire if speaker wants to emphasize a process, @bellec_construction_2014). In contrast, the se faire passive may be preferred if the action is described in the imparfait (only a guess).
    -   **Aspect** tagging @hundt_be-_2024 : simple, perfect, progressive; @bohmann_large-scale_2023 [39]: "Form of auxiliary verb: present, preterite, perfect, infinitive, -ing"
        -   @raineri_analyse_2010 [365] sees the "type of process denoted by the verb" as a "necessary condition", inchoative/ durative aspect: "get passive and se faire passive focus on onset of the process that affects the subject referent, whereas be and etre passive present the process in its entirety, without offering an internal perspective."
    -   **Mode?** actuality, possibility, necessity) -
        -   Modal verbs (auxiliary): such as "devoir", "savoir", "pouvoir", "falloir", "vouloir" very frequent with se faire passive @novakova_construction_2009 \[9\] (in @raineri_analyse_2010 [97], reason: they denote "control" to the subject IF POSSIBLE

-   **(Periphrase constructions**: According to @spang-hanssen_quelques_1967 \[141\] and @novakova_construction_2009, periphrases that denote progressivity such as “être en train de + INF”, “venir de + INF” and “aller + INF” occur more naturally with the se faire passive than with the être passive. This is due to the dynamicity that the se faire passive entails (Quelle) compared to the more stative être passive. (also: @bellec_construction_2014; @keenan_passive_2007 [341], @raineri_analyse_2010 [91])

### Adversativity/ Non-neutrality

The factor of "Adversativity/ Non-neutrality" is reflected in the semantics of the verb and captures the (negative) consequences of the described action for the subject. It has been found to be relevant to the choice between the be and get passive (@hundt_be-_2024 [89, 96-97]; @bohmann_large-scale_2023). @spang-hanssen_quelques_1967 and @raineri_analyse_2010 draw parallels between the English get passive and the French PSF construction.

@gaatone_desagreable_1983 argues that the se faire passive has a strong "unpleasant sense" (sens désagreable") because it occurs mainly with verbs denoting unpleasant events [^3]. Although @kupferman_construction_1995 \[77\] rejects the "unpleasant sense" as a characteristic of se faire, he agrees that it is more commonly used in unpleasant contexts ("has tended to specialize in unpleasant roles" (compared to être, which is more neutral). This is supported by corpus data: @novakova_construction_2009 [110] found a clear preference for PSF to be realised with unpleasant actions, with 30% of verbs in PSF denoting unpleasant events and only 5% denoting pleasant events. Also: @raineri_analyse_2010 [247].

[^3]: According to @gaatone_desagreable_1983, unpleasant events contradict the notion of subject responsibility because subjects would not willingly bring the unpleasant event upon themselves (argument for "primacy of causative reading"). According to him, the reason this is the subject's involvement in se faire constructions: if the subject experiences a pleasant event and this is expressed by se faire -\> mostly the subject has initiated the action -\> causative and not passive reading (language example?)/ in contrast, it is less likely that the subject initiates an unpleasant action.

In their corpus study on the English be - get passive alternation, @hundt_be-_2024 label it as binary on the verb as adversative, non-adversative/neutral. For practicality and clear statistical modelling, We follow them in our annotation: "Adversative" if event negatively affects the subject, "non-adversative" if event is neutral or positive.

<!-- -   \@bohmann_large-scale_2023 [38, 43] Sentiment values of verb participles extracted from SentiWordNet: positivity, negativity, objectivity scores; reversed if sentence is negated -->

As @bohmann_large-scale_2023 notes, there are two hypotheses related to emotions of the verb: The adversativity and the non-neutrality hypothesis. Both expect the get passive to favour verbs with negative scores; but only the non-neutrality hypothesis predicts get passive to correlate positively with positive emotion scores as well. Since positive emotion verbs have been shown to be quite rare with PSF (@novakova_construction_2009 [110]) we have chosen to not test them/ put them together with "neutral".

**Hypothesis:** Se faire passive be preferred over the être passive for verbs that denote "unpleasant" actions (as reflected by the adversativity rating).

**Interaction:** Affecteness of the subject may be reflected by an interaction of subject animacy and verb adversativity: Negative consequences of actions, expressed by adversative verbs, affect animate subjects more than inanimate subjects.

### (Maybe: Register)

Text genre, formal vs. informal

-   se faire passive more likely in informal settings, être passive more likely in formal settings

### Morphosyntactic features

Negation, subject person, subject number, subject gender

-   **Person**

-   **Gender:** We want to test whether grammatical gender influences the alternation between *être* and *se faire* passives. Since previous findings indicate that a feminine **agent** increases the likelihood of *être* passive (passif court, which means the agent was omitted) over active voice, it is worth exploring whether this effect extends to the choice between *être* and *se faire* constructions. \@da_cunha_lalternance_2020 (+2022) found that in written French a feminine agent increased the likelihood of *être* passive (passif court model), with no interaction with animacy (whether the argument was human or non-human) (or definiteness). This could suggest that gender itself plays a direct role in voice choice. However, given the small sample size of the study, it is also possible that an interaction effect exists but was not statistically detectable in this study (check 2022 study). We want to test whether gender also plays a role independently for our alternation. Also: we expect (almost) no inanimate subjects in *se faire* passive contructions.\
    \<-\> prominence/subject responsibility? (similar to animacy factor for *se faire* passive? grammatical gender though, not social gender but only expected to refer to animate feminine vs. masculine subjects for se faire??)

    -   levels of the gender factor:

    -   Hypothesis: A feminine **agent** favours the *être* passive /disfavours the *se faire* passive??

-   **Number:** Surprisingly, number also had an effect in @da_cunha_lalternance_2020: plural arguments favoured the passive over active in all models. (+ Relativsätze & Priming als signifikante Faktoren) @da_cunha_lalternance_2020

@bohmann_large-scale_2023 [39]: get passive disfavoured in complex constructions (e.g. negation); parallel in French?

-   **Negation:**

## Memory based learning

-   non-parametric: no assumptions about underlying distribution of data
-   no problem with sparse observation of features/ low-frequency events @theijssen2013choosing [243, 244] because it "does not rely on any kind of generative model"
-   only mandatory annotation is class label @theijssen2013choosing [243] (passive choice)
-   "simple-theory neutral features"/ or some linguistic features (animacy, verb, pp-phrase...)
-   compares test instances with labeled training data (stored in the memory)
-   MBL very susceptible to majority classes -\> unbalanced data is a big problem (poor performance for minority classes possible)
-   k-Nearest-Neighbor (kNN):
    -   kNN algorithm is trained on data to predict the most likely passive choice
    -   nearest neighbours out of all instances are determined -\> observation receives class of majority of nearest neighbours
    -   leave-one-out-setting @theijssen2013choosing [244]: train on all but one, test on remaining instance, iterate for each instance
    -   hyperparameters setting (e.g. number of neighbors)? -\> help of wrapper
    -   Tutorial: <https://www.datacamp.com/tutorial/k-nearest-neighbors-knn-classification-with-r-tutorial>

1.  choose the value of K, which is the number of nearest neighbors that will be used to make the prediction.
2.  Calculate the distance between that point and all the points in the training set.
3.  Select the K nearest neighbors based on the distances calculated.
4.  Assign the label of the majority class to the new data point.
5.  Repeat steps 2 to 4 for all the data points in the test set.
6.  Evaluate the accuracy of the algorithm.

-   in R: "class" or "caret" package
-   set seed for reproducibility
-   split set manually or by "caTools" library
-   Preprocessing/ Feature scaling if feautres have different scales, to ensure proportions between features and prevent that features with largest scale dominate distance calculations
-   caret: Model Tuning: find optimal value of "K"
-   caret: train best performing model
-   Model evaluation

# Results

# Discussion

# Conclusion/Outlook

Future studies: modality (spoken/written) as a factor (compare modalities), analyses on other modalities/varieties/registers

# References
